---
title: "DURIN 2024 Thermal Tolerance"
format: html
editor: visual
---

## Import Data

```{r}
library(tidylog)
library(tidyverse)
library(readxl)
library(stringr) 
library(plyr)
# Read in data from online spreadsheet that's downloaded to your personal drive.
# Change the directory to be a local place on your computer.
DURIN_data <- read_excel(path = "C:/Users/sge010/OneDrive - University of Bergen/Documents/DURIN/Durin_data_project/data/raw_data/2024 DURIN Plant Functional Traits.xlsx",
                         sheet = "Data", col_names = TRUE, trim_ws = TRUE)

# Note for now, some issues with column types reading in.
# For example, the collection time with "10:59" format doesn't read in correctly, nor does the date format for Treatment Date.



# Also, some column names aren't 'readable', i.e. Fv/Fm with '/', or height(s)
# So we will change these names here
DURIN_data <- plyr::rename(DURIN_data, c("height(s)"="heights",
                     "Fv/Fm_Initial"="FvFm_Initial",
                     "Fv/Fm_PostStress"="FvFm_PostStress",
                     "Fv/Fm_Hr"="FvFm_Hr",
                     "Fv/Fm_Final"="FvFm_Final",
                     "DroughNet_plotID" = "DroughtNet_plotID"))

# Senja is spelt wrongly as Senje in DropDowns
DURIN_data$siteID<- revalue(DURIN_data$siteID, c("Senje" = "Senja"))

# Create concantenated ID code for the plants 
DURIN_data$PlantSampleID = paste(DURIN_data$siteID, DURIN_data$habitat,
                                   DURIN_data$species,
                                   sep = "_")
# Version with the plant number also included
DURIN_data$PlantSampleID2 = paste(DURIN_data$siteID, DURIN_data$habitat,
                                   DURIN_data$species, DURIN_data$plant_nr,
                                   sep = "_")

```

## General Custom Cleaning/Issue Identifying Codes

```{r}
# Useful custom functions ----
## Look up an individual envelope ID number to see its info ----
# Modify the select() to change which data are displayed
check <- function(barcode) {
  data <- DURIN_data |>
    select(envelope_ID, day, month, year, siteID, habitat, species, proj_2024,
           project,
           ageClass, DroughtTrt, DroughtNet_plotID,
           plant_nr, leaf_nr, leaf_age, sample,
           CollTime, PairedWith, calluna_shoot_type, bulk_nr_leaves,
           TempExp, TempExpTime, TempExpTimeUnit, FvFm_Initial, FvFm_PostStress, FvFm_Hr, FvFm_Final, 
           wet_mass_g,
           leaf_thickness_1_mm, leaf_thickness_2_mm, leaf_thickness_3_mm,
           dry_mass_g, scanned,
           nearest_DURIN_plotID, sampleType, heights,
           TreatmentDate, TreatmentTime,
           remark, remarks_2, flag,
           # New Column, not on sheet but helpful
           PlantSampleID
           ) |>
    filter(envelope_ID == barcode)
  data
}
```

## Data Checking DURIN_data

For now, these codes run, but we have not fixed all the issues with mis-aligned site or habitat allocations, or plants that don't seem to have their full complement of leaves.

[**This needs to be still resolved!!**]{.underline}

```{r}
## Find duplicate barcodes ----
error.duplicates <- DURIN_data |>
  # We don't care about exact duplicates so distinct() merges those
  distinct() |>
  # Then this filters to only duplicates
  group_by(envelope_ID) %>%
  filter(n() > 1)

## Young/old ----
# Vaccinium myrtillus can by definition only have new leaves as it is deciduous
error.oldVM <- DURIN_data |>
  # Filter to young-only plants
  filter(species %in% c("Vaccinium myrtillus")) |>
  # Filter to mislabelled leaf ages
  filter(leaf_age == "old")


## Leaf Types vs Temperatures ----
# For hot runs
# Leaf 1 = 20C, Leaf 2 = 30C, Leaf 3 = 34C, Leaf 4 = 38C,
# Leaf 5 = 42C, Leaf 6 = 46C, Leaf 7 = 50C, Leaf 8 = 54C

# For cold runs
# Leaf 9 = 5C, Leaf 10 = 0C, Leaf 11 = -5C, Leaf 12 = -10C,
# Leaf 13 = -15C, Leaf 14 = -20C
error.leafNum.leafTemp <- DURIN_data %>% 
  # Table leaf number by treatment temperatures
  # Can remove siteID, species, habitat for the more general overview
  group_by(proj_2024, siteID, species, habitat, leaf_nr, TempExp, TempExpTime) %>% 
  tally()

## FvFm Recovery Times for Hot and Cold Runs ----
# For hot runs: general recovery time should be 24hrs, with some 18hr exceptions
# For cold runs: general recovery time is 48hrs

error.TempExpTime <- DURIN_data %>% 
  # Table leaf number by treatment temperatures
  # Can remove siteID, species, habitat for the more general overview
  group_by(proj_2024, TempExpTime, TempExpTimeUnit) %>% 
  tally()

## Single vs Bulk Sampled Leaves ----
# For VV and VM they should be single leaf samples
# For CV and EN they should be bulked leaves, however, the numbers are variable. But all Calluna and Empentrum with bulked leaf samples should have a bulk leaf number provided. Later will check how envelope bulk number aligns with the leaf area scanned counts. 

# There appears to be quite a lot of variation across the samples.
# The VV and VM could be controlled for with code, assigning a bulk leaf nr sample of 1 for each of these. All CV and EN should be assigned bulk not leaf_nr, as we never would have only sampled an individual leaf (minimum weight thresholds meant often 4-5 leaves, if not more would have been required)

error.SingleBulkedLeaves <- DURIN_data %>% 
  # Table leaf number by treatment temperatures
  # Can remove siteID, species, habitat for the more general overview
  group_by(siteID, habitat, species, sample, bulk_nr_leaves) %>% 
  tally()

# TODO: Use the wet or dry weights to confirm number of bulked vs individual leaves. 


## Single vs Multiple Individuals ----
# We often had issues where a single plant was unable to contribute the 14-20 leaves required to go across multiple projects, i.e. hot, cold, hot rapid, and stomatal peels. Therefore, multiple individuals within close proximity were harvested together, and assigned as a single 'plant'. In these instance, multiple plant height measurements were taken and seperated by ' / ' format.

# Split heights into multiple columns, one per measurement
# Currently there is a maximum of 8 individuals, so 8 new columns could be created
DURIN_data <- DURIN_data %>% 
  separate_wider_delim(heights, delim = " / ",
                       names = c("Height_1", "Height_2", "Height_3", "Height_4",
                                 "Height_5", "Height_6", "Height_7", "Height_8"),
                       too_few = "align_start",
                       cols_remove = FALSE) %>%
  mutate_at(c('Height_1','Height_2', 'Height_3', 'Height_4',
              'Height_5', 'Height_6', 'Height_7', 'Height_8'), as.numeric)

# Add a new column for the mean height of the individuals representing a single plant
DURIN_data <- DURIN_data %>% 
mutate(Height_mean = rowMeans(cbind(Height_1, Height_2, Height_3, Height_4,
                                    Height_5, Height_6, Height_7, Height_8),
                              na.rm = TRUE))


# During processing, if all leaves did manage to be taken off a single individual, even though multiple were collected, then 'individual' should be flagged so that if we wanted we could use the samples for genetics etc later on.

# Conversely, a sampled plant with only one individual, shouldn't have multiple heights associated with it, or the option 'multiple' for sample type.

error.SingleMultiple <- DURIN_data %>% 
  # Table single vs multiple indivduals
  # Can remove siteID, species, habitat for the more general overview
  group_by(proj_2024, siteID, habitat, species, sampleType) %>% 
  tally()

# Code to identify if samples came from individual plants or groups of plants
# Currently doesn't work, hence commented out.
DURIN_data <- DURIN_data %>%
  rowwise() %>%
  # Count how many height measurements we have
  mutate(Height_count = sum(!is.na(c_across(Height_1:Height_8)))) %>% 
  # Allocate column for easy filtering?
  mutate(SingleIndividual_Check = case_when(Height_count >= 2 ~ "Multiple",
                            Height_count == 1 ~ "Individual",
                            Height_count == 0 ~ "None",
                            T ~ as.character(Height_count)))



## Are the FvFm values within range? ----
# Values are between 0.00 and 0.90
# Values should have 2 decimal places

count_decimals = function(x) {
  #length zero input
  if (length(x) == 0) return(numeric())

  #count decimals
  x_nchr = x %>% abs() %>% as.character() %>% nchar() %>% as.numeric()
  x_int = floor(x) %>% abs() %>% nchar()
  x_nchr = x_nchr - 1 - x_int
  x_nchr[x_nchr < 0] = 0

  x_nchr
}

count_decimals(as.numeric(DURIN_data$FvFm_Initial))
count_decimals(as.numeric(DURIN_data$FvFm_Final))

# Check the range of FvFm
check_FvFm_range <- function(data, column_name = "FvFm") {
  # Filter rows where FvFm is outside the range [0, 0.9]
  issues <- data %>%
    filter(!!sym(column_name) < 0 | !!sym(column_name) > 0.9)
  
  return(issues)
}

# Test
problematic_FvFm <- check_FvFm_range(DURIN_data, column_name = "FvFm_Final")
print(problematic_FvFm)

# Some seem to be in scientific notation instead of decimals, correct this.
# Others are categorical i.e. 'Low Value'
# Convert to numeric format
DURIN_data$FvFm_Final <- case_when(DURIN_data$FvFm_Final == "Low Value" ~ NA,
                                     .default = as.character(DURIN_data$FvFm_Final))
# Make sure the FvFm_Final is numeric
DURIN_data$FvFm_Final <- as.numeric(as.character(DURIN_data$FvFm_Final))



## DURIN plot info ----
### Does the envelope site match the plot site? ----
### TODO: YES SOME OF THIS NEED TO BE FIXED!!
error.durin.site <- DURIN_data |>
  #Filter to just DURIN
  filter(DroughtNet_plotID == "Not Applicable") |>
  # Separate out the plot ID for identifying species
  # separate_wider_delim(nearest_DURIN_plotID, delim = "_", 
  #                      names = c("site.abbrv", "habitat.abbrv", "spp.abbrv", "rep.abbrv", remove = FALSE)) |>
  # New version
  # # Ensures missing pieces are replaced with NA
  separate(nearest_DURIN_plotID, into = c("site.abbrv", "habitat.abbrv", "spp.abbrv", "rep.abbrv"),
           sep = "_", fill = "right", remove = FALSE)  |>
  # Create abbreviated site codes
  mutate(site.code = case_when(
    siteID == "Lygra" ~ "LY",
    siteID == "Sogndal" ~ "SO",
    siteID == "Senja" ~ "SE",
    siteID == "Kautokeino" ~ "KA",
    TRUE ~ "unknown"
  )) |>
  # Filter to problem envelopes
  filter(site.code != site.abbrv) |>
  # Make it human readable
  relocate(c(site.code, site.abbrv), .after = envelope_ID)


### Does the envelope habitat match the DURIN plot habitat? ----
### TODO: YES SOME OF THIS NEED TO BE FIXED!!
error.durin.habitat <- DURIN_data |>
  #Filter to just DURIN
  filter(DroughtNet_plotID == "Not Applicable") |>
  # Separate out the plot ID for identifying species
  # separate_wider_delim(nearest_DURIN_plotID, delim = "_", 
  #                      names = c("site.abbrv", "habitat.abbrv", "spp.abbrv", "rep.abbrv", remove = FALSE)) |>
    # New version
  # # Ensures missing pieces are replaced with NA
  separate(nearest_DURIN_plotID, into = c("site.abbrv", "habitat.abbrv", "spp.abbrv", "rep.abbrv"),
           sep = "_", fill = "right", remove = FALSE)  |>
  # Create abbreviated species codes
  mutate(habitat.assigned = case_when(
    habitat == "Open" ~ "O",
    habitat == "Forested" ~ "F",
    TRUE ~ "unknown"
  )) |>
  # Filter to problem leaves
  filter(habitat.assigned != habitat.abbrv) |>
  # Make it human readable
  relocate(c(habitat.assigned, habitat.abbrv, nearest_DURIN_plotID), .after = envelope_ID)



### How many leaves are there for each plant? ----
# RECOMMENDED you run this AFTER updating all the other fields listed above
# Calculate the maximum number of leaves for each plant
durin.max.leaf_nr <- DURIN_data |>
  # Filter to just DURIN
  filter(DroughtNet_plotID == "Not Applicable") |>
  # Select the columns needed
  select(siteID, habitat, species, plant_nr, leaf_nr) |>
  # Remove duplicate data
  # distinct() |>
  # Sort by plant number and slice the maximum
  group_by(siteID,habitat, species, plant_nr) |>
  slice_max(leaf_nr) |>
  dplyr::rename(max.leaf.n = leaf_nr)


## Replicates ----
### Plant number -----
# Which envelopes are missing plant number?
error.durin.plantnr.missing <- DURIN_data |>
  # Filter out known issues (wrong species)
  # filter(!envelope_ID %in% c("AYN9607", "AST3380")) |>
  # Filter to NA values
  filter(is.na(plant_nr))

nr.missing.list <- as.list(error.durin.plantnr.missing$envelope_ID)



### Leaf number ----
check.plant <- function(plot, plantnr) {
  data = durin |>
    select(envelope_ID, siteID,
           DURIN_plot,
           plotNR, habitat,
           species, plant_nr, leaf_nr, leaf_age, plant_height,
           calluna_shoot_type) |>
    filter(DURIN_plot == plot & plant_nr == plantnr) |>
    arrange(plant_nr, calluna_shoot_type, leaf_age, leaf_nr)
  data
}

# Which envelopes are missing leaf numbers?
error.durin.leafnr.missing <- DURIN_data |>
  # Filter to NA values
  filter(is.na(leaf_nr))

# Which plants are missing leaves (or have extras)?
# TODO: Check that some envelopes still don't need weighed or entered!!
error.durin.leaves <- DURIN_data %>% 
  filter(DroughtNet_plotID == "Not Applicable") %>% 
  filter(proj_2024 %in% c("TT Hot", "TT Cold")) %>% 
  dplyr::group_by(siteID, habitat, species, plant_nr) %>%
  dplyr::summarise(n = dplyr::n(), .groups = "drop") %>%
  dplyr::filter(n != 14)

### Leaf age ----
# Which envelopes are missing leaf ages?
# TODO: Fix these from envelopes if possible.
error.durin.age.missing <- DURIN_data |>
  # Filter to species with young and old leaves
  filter(species %in% c("Vaccinium vitis-idaea", "Empetrum nigrum")) |>
  # Filter to NA values
  filter(is.na(leaf_age))

age.missing.list <- as.list(error.durin.age.missing$envelope_ID)



### Samples with incomplete meta-data for wet and dry weights ----
# TODO: Go and get these weights.
# Function to identify missing values for wet and dry mass
# Checks for missing values in wet_mass_g and dry_mass_g, then creates a new column (missing_data_status) that categorizes each row based on what is missing.
identify_missing_mass_weights <- function(data) {
  data <- data %>%
    mutate(missing_mass_status = case_when(
      is.na(wet_mass_g) & is.na(dry_mass_g) ~ "Both wet and dry mass missing",
      is.na(wet_mass_g) ~ "Wet mass missing",
      is.na(dry_mass_g) ~ "Dry mass missing",
      TRUE ~ "No missing data"
    ))
  
  # Filter only rows with missing data
  missing_mass_rows <- data %>%
    filter(missing_mass_status != "No missing data")
  
  return(missing_mass_rows)
}

# Example usage
error.missing_mass_data <- identify_missing_mass_weights(DURIN_data)




```

## Flagging Data Value Issues

Do not run this section of code either. Haven't yet weighed all the samples so many values do not yet make sense.

```{r eval=FALSE}
## Do wet and dry mass hold an expected relationship? ----
library(ggh4x)

ggplot(DURIN_data,
       aes(x = as.numeric(wet_mass_g), y = as.numeric(dry_mass_g), color = leaf_age)) +
  geom_point(alpha = 0.5) +
  facet_wrap2(. ~ species, scales = "free") +
  theme_bw()

# ggsave("visualizations/2023.10.17_dry.wetMass_scaledtobulkleafnr.png")

ggplot(durin |> mutate(dry_mass_g = as.numeric(dry_mass_g))|>
         mutate(mass_ratio = dry_mass_g/wet_mass_g),
       aes(x = species, y = mass_ratio, color = leaf_age)) +
  geom_boxplot(outlier.shape = NA) +
  geom_point(position = position_jitterdodge()) +
  geom_point(data = durin |> mutate(mass_ratio = dry_mass_g/wet_mass_g) |>
               filter(envelope_ID %in% reweigh), color = "red") +
  facet_wrap2(~species, scales = "free") +
  theme_bw()

ggsave()

# Identify the outliers
error.massratio <- DURIN_data |>
  # Make dry mass numeric
  mutate(dry_mass_g = as.numeric(dry_mass_g)) |>
  # Make ratio of dry to wet mass
  mutate(mass_ratio = dry_mass_g/wet_mass_g) |>
  # Outlier values are eyeballed from the above graph
  mutate(flag = case_when(
    species == "Calluna vulgaris" & wet_mass_g > 0.1 ~ "high wet_mass_g",
    species == "Calluna vulgaris" & dry_mass_g > 0.007 ~ "high dry_mass_g",
    species == "Empetrum nigrum" & wet_mass_g > 0.05 ~ "high wet_mass_g",
    species == "Empetrum nigrum" & dry_mass_g > 0.01 ~ "high dry_mass_g",
    species == "Vaccinium myrtillus" & wet_mass_g > 0.1 ~ "high wet_mass_g",
    species == "Vaccinium myrtillus" & dry_mass_g > 0.027 ~ "high dry_mass_g",
    species == "Vaccinium vitis-idaea" & wet_mass_g > 0.25 ~ "high wet_mass_g",
    species == "Vaccinium vitis-idaea" & dry_mass_g > 0.075 ~ "high dry_mass_g",
    species == "Calluna vulgaris" & mass_ratio > 0.75 ~ "high dry to wet mass ratio",
    species == "Empetrum nigrum" & mass_ratio > 0.75 ~ "high dry to wet mass ratio",
    species == "Vaccinium myrtillus" & mass_ratio > 0.6 ~ "high dry to wet mass ratio",
    species == "Vaccinium myrtillus" & mass_ratio < 0.2 ~ "low dry to wet mass ratio",
    species == "Vaccinium vitis-idaea" & mass_ratio > 0.75 ~ "high dry to wet mass ratio",
    species == "Vaccinium vitis-idaea" & mass_ratio < 0.2 ~ "low dry to wet mass ratio",
    dry_mass_g > wet_mass_g ~ "dry mass more than wet mass",
    TRUE ~ "ok"
  )) |>
  # Filter to the problem columns
  filter(flag != "ok")|>
  # Prep for the envelope check sheet
  mutate(`Supporting Text Comment` = paste0("Check envelope", ". ", flag)) |>
  select(envelope_ID, `Supporting Text Comment`, mass_ratio) |>
  arrange(`Supporting Text Comment`)

# For some reason the flag for dry mass greater than wet doesn't work
table(error.massratio$flag)

write.csv(error.massratio, "output/2024.05.15_error.massratio.csv")

## Do SLA and leaf area hold an expected relationship? ----
### Visualize outliers

ggplot(DURIN_data,
       aes(x = leaf_area, y = SLA, color = leaf_age)) +
  geom_point(alpha = 0.3) +
  facet_wrap(~species, scales = "free", ncol = 4) +
  theme_bw()

ggsave("visualizations/2024.07.15_SLAxAreaErrors_scaledtobulkleafnr.png")

# From these visuals, we can estimate reasonable values
error.SLAxArea <- error.sla |>
  # Flag possible errors
  mutate(flag_SLA = case_when(
    species == "Calluna vulgaris" & (SLA < 25 | SLA > 150) ~ "CV SLA error",
    species == "Empetrum nigrum" & (SLA < 40 | SLA > 200) ~ "EN SLA error",
    species == "Vaccinium myrtillus" & (SLA < 125 | SLA > 400) ~ "VM SLA error",
    species == "Vaccinium vitis-idaea" & (SLA < 40 | SLA > 300) ~ "VV SLA error",
    TRUE ~ "okay"
  ),
  flag_area = case_when(
    species == "Calluna vulgaris" & (leaf_area_scaled < 0 | leaf_area_scaled > 0.3) ~ "CV area error",
    species == "Empetrum nigrum" & (leaf_area_scaled < 0 | leaf_area_scaled > 0.2) ~ "EN area error",
    species == "Vaccinium myrtillus" & (leaf_area_scaled < 0 | leaf_area_scaled > 3.5) ~ "VM area error",
    species == "Vaccinium vitis-idaea" & (leaf_area_scaled < 0 | leaf_area_scaled > 3.5) ~ "VV area error",
    TRUE ~ "okay"
  )) |>
  filter(flag_SLA != "okay" | flag_area != "okay") |>
  select(envelope_ID, flag_SLA, flag_area, wet_mass_g, dry_mass_g, leaf_area, SLA) |>
  distinct()

write.csv(error.SLAxArea, "output/2024.107.115_SLAxAreaErrors.csv")

```

## Exploratory Thermal Curves

```{r warning=FALSE}
# Hot Tolerance dataset only
DURIN_TT <- subset(DURIN_data, envelope_ID != "") 
DURIN_TT_HOT <- DURIN_TT %>% 
  filter(proj_2024 == "TT Hot") %>% 
  filter(DroughtNet_plotID == "Not Applicable") %>% 
  # Extra 'crude' checking for mislabelled leaves from the cold treatment
  filter(leaf_nr %in% c(1,2,3,4,5,6,7,8),
         TempExp %in% c(20,30,34,38,42, 46, 50,54))

# Plot Initial Curves
DURIN_TT %>% 
  # Filters out issues with hot plants with temps below lowest bath of 20
  filter(TempExp >= 20) %>% 
  # Various filters to remove samples associated with different projects below
  filter(proj_2024 != "Multiple") %>% 
  filter(proj_2024 == "TT Hot") %>% 
  filter(DroughtNet_plotID == "Not Applicable") %>% 
  # Extra 'crude' checking for mislabelled leaves from the cold treatment
  filter(leaf_nr %in% c(1,2,3,4,5,6,7,8),
         TempExp %in% c(20,30,34,38,42, 46, 50,54)) %>% 
  # Create a plot of exposure temperature (TempExp) against the final FvFm measurements. 
  # Group by the individuals that were sampled, and colour by habitat
  ggplot(aes(x = TempExp, y= as.numeric(as.character(FvFm_Final)),
             group = PlantSampleID, color = habitat)) +
  # Add points
  geom_point()+
  # Add the smoothed line across the points, again coloured by habitat
  # One line per individual
  geom_smooth(method = "loess",
              se = FALSE, aes(group = PlantSampleID, color = habitat)) +
  labs(x = "Exposure Time (mins)", y = "FvFm", color = "Habitat") + 
  # Make a differnet panel of the plot per site
  facet_grid(species ~ siteID)

# Plot Initial Curves
DURIN_TT %>% 
  # Filters out issues with hot plants with temps below lowest bath of 20
  filter(TempExp >= 20) %>% 
  filter(siteID == "Senja") %>% 
  filter(species %in% c("Vaccinium vitis-ideae","Vaccinium myrtillus")) %>% 
  filter(proj_2024 != "Multiple") %>% 
  filter(proj_2024 == "TT Hot") %>% 
  filter(DroughtNet_plotID == "Not Applicable") %>% 
    # Extra 'crude' checking for mislabelled leaves from the cold treatment
  filter(leaf_nr %in% c(1,2,3,4,5,6,7,8),
         TempExp %in% c(20,30,34,38,42, 46, 50,54)) %>% 
  ggplot(aes(x = TempExp, y= as.numeric(as.character(FvFm_Final)),
             color = habitat)) +
  geom_point()+ 
  geom_smooth(method = "loess",
              se = FALSE, aes(
                color = habitat)) +
  labs(x = "Exposure Time (mins)", y = "FvFm", color = "Habitat") + 
  facet_grid(species ~ siteID)


# Plot Initial Curves - For Methods Test
DURIN_TT %>% 
  # Filters out issues with hot plants with temps below lowest bath of 20
  filter(TempExp >= 20) %>% 
  filter(siteID == "Senja") %>% 
  filter(proj_2024 != "Multiple") %>% 
  filter(proj_2024 == "TT Hot") %>% 
  filter(DroughtNet_plotID == "Not Applicable") %>% 
    # Extra 'crude' checking for mislabelled leaves from the cold treatment
  filter(leaf_nr %in% c(1,2,3,4,5,6,7,8),
         TempExp %in% c(20,30,34,38,42, 46, 50,54)) %>% 
  ggplot(aes(x = TempExp, y= as.numeric(as.character(FvFm_Final)),
             color = habitat)) +
  geom_point()+ 
  geom_smooth(method = "loess",
              se = FALSE, aes(
                color = habitat)) +
  labs(x = "Exposure Time (mins)", y = "FvFm", color = "Habitat") + 
  facet_grid(species ~ siteID)


```

```{r}
#Thanks for checking out this code for estimating heat tolerances using variable fluorescence!
#Be sure to the read the Read.me.txt file.
#Below, the "psiiht" function can calculate heat tolerances of PSII heat tolerance. 
#For this function, define the the columns with the 1) Temperature & 2) FvFm variables, then the 3) control temperature, the 4) id/unique factor that you 
#want to estimate the heat tolerances for, 5) indicate if you want to make plots, 6) the number of bootsrap iterations for estimating heat tolerances

# setwd("C:/Users/sge010/OneDrive - University of Bergen/Documents/DURIN/Durin_data_project/Heat_tolerance_function/")
library(car)
#htdata = read.csv("Sample_FvFm_data.csv")
psiiht=function( Temperature, FvFm, control.temp, id, plot.est, boots){
  l1=list(control.temp = control.temp, plot.est = plot.est, boots = boots)
  attach(l1)
  HTdf = data.frame(Temperature = Temperature, FvFm = FvFm, id = id)
  
 return(do.call("rbind", by(HTdf, list(HTdf$id), function(df){
   Temperature=df[,which(colnames(df)=="Temperature")]
   FvFm=df[,which(colnames(df)=="FvFm")]
   id=df[,which(colnames(df)=="id")]
   #get parameter estimates for logistic decay model
   cof=coef(lm(logit(FvFm)~Temperature)) 
   #Fit a non linear least squares model to the FvFm and Temperature data
   HT.model <- nls(FvFm ~ theta1/(1 + exp(-(theta2 + theta3*Temperature))),  start=list(theta1 = .8, theta2 = cof[1], theta3 = cof[2]),
                   trace=F, control=list(maxiter=1000, tol=1e-3))
   
   #Use the parameter estimates (coef(HT.model)[#])from the HT.model to predict a new fit based on a heat treatments from 23-62 degrees celcius. Here, # = 1:3.
   y<-coef(HT.model)[1]/(1+exp(-(coef(HT.model)[2]+coef(HT.model)[3]*seq(20,66)))) 
   
   #Calculate half of the control Fv/Fm & a 95% reduction in FvFm with reference to control
   half=mean(na.omit(FvFm[which(Temperature==control.temp)]))/2  
   nine5=mean(na.omit(FvFm[which(Temperature==control.temp)]))*0.05  
   #95 Confidence Interval
   predict.boot=matrix(NA,40, boots)
   T95=T50=Tcrit=c()
   for(k in 1:boots){
     #print(k)
     srows <- sample(1:length(Temperature), length(Temperature),TRUE)
     
     if(class(try(nls(FvFm[srows] ~ theta1/(1 + exp(-(theta2 + theta3*Temperature[srows]))),  start=list(theta1 = .8, theta2 = cof[1], theta3 = cof[2]),
                      trace=F, control=list(maxiter=1000, tol=1e-3)), silent=T)[[1]])=="nlsModel")
     {HT.model2 <- nls(FvFm[srows] ~ theta1/(1 + exp(-(theta2 + theta3*Temperature[srows]))),  start=list(theta1 = .8, theta2 = cof[1], theta3 = cof[2]),
                       trace=F, control=list(maxiter=1000, tol=1e-6))
     predict.boot[,k]=coef(HT.model2)[1]/(1+exp(-(coef(HT.model2)[2]+coef(HT.model2)[3]*seq(20,66)))) 
     #Estimate T95
     T95[k]=(-log((coef(HT.model2)[1]/nine5)-1)-coef(HT.model2)[2])/coef(HT.model2)[3] 
     
     #Estimate T50
     T50[k]=(-log((coef(HT.model2)[[1]]/half)-1)-coef(HT.model2)[[2]])/coef(HT.model2)[[3]]
     T50k=(-log((coef(HT.model2)[[1]]/half)-1)-coef(HT.model2)[[2]])/coef(HT.model2)[[3]]
     #Use model to predict changes in FvFm & make new dataframe
     predict=data.frame(x=seq(20,66),y=coef(HT.model2)[1]/(1+exp(-(coef(HT.model2)[2]+coef(HT.model2)[3]*seq(20,66)))) ) #create a dataframe of predictions
     df1=cbind(predict[-1,], predict[-nrow(predict),])[,c(3,1,4,2)]
     #Use new dataframe to estimate the slope at between each 1-degree interval
     df1$slp=as.vector(apply(df1, 1, function(x) summary(lm((x[3:4]) ~ x[1:2])) [[4]][[2]] ))
     slp.at.tcrit=round(min(df1$slp), 3)*.15 #Determine where slope is 15% of max slope & round
     #Estimate the FvFm at which the slope is 15% of max slope & less than T50
     fvfv.at.tcrit=df1[which(abs(df1[which(df1[,1]<T50k),]$slp-slp.at.tcrit)==min(abs(df1[which(df1[,1]<T50k),]$slp-slp.at.tcrit))),][1,3]
     Tcrit[k]=(-log((coef(HT.model2)[[1]]/fvfv.at.tcrit)-1)-coef(HT.model2)[[2]])/coef(HT.model2)[[3]] # Estimate the temperatureat which the slope is 15% of max slope
     
     }else{(class(try(nls(FvFm ~ theta1/(1 + exp(-(theta2 + theta3*Temperature))),  start=list(theta1 = .8, theta2 = cof[1], theta3 = cof[2]),
                          data=data2[srows,], trace=F, control=list(maxiter=1000, tol=1e-3)), silent=T)[[1]])=="list")
       predict.boot[,k]=NA
       T95[k]=NA
       T50[k]=NA
       Tcrit[k]=NA }}
   
   FvFm.boot=t(apply(predict.boot, 1, function(x){quantile(x,c(0.025,0.975),na.rm=T)}))
   
   #Tcrit.ci=quantile(Tcrit,c(0.025,0.975),na.rm=T)
   T50.ci=quantile(T50,c(0.025,0.975),na.rm=T)
   #T95.ci=quantile(T95,c(0.025,0.975),na.rm=T)
   
   if(plot.est==T){ 
     plot(NULL, NULL,xlab="Temperature",ylab="Fv/Fm",xlim=c(20,66),ylim=c(0,0.9), bty="l", lty=2)
     text(22,0.3, pos=4, paste(unique(paste(id))), font=4)
     points(Temperature, FvFm, xlab="Temperature", ylab="Fv/Fm",xlim=c(20,66), ylim=c(0,0.85), bty="l", lty=2, pch=5,  col="black")
     lines(seq(20,66), y,lwd=1, col="black")
     if(boots>1){
     lines(seq(20,66), FvFm.boot[,1],lty=3, col="black")
     lines(seq(20,66), FvFm.boot[,2],lty=3, col="black")
     }
     text(30,0, paste('Tcrit:', round(mean(na.omit(Tcrit)),1)), pos=4,col="light gray")
     abline(v=round(mean(na.omit(Tcrit)),1), lty=2,lwd=1.5, col="light gray",  cex=0.8)
     text(30,.1, paste('T50:',round(mean(na.omit(T50)),1)), pos=4, col="gray")
     abline(v=round(mean(na.omit(T50)),1), lty=2,lwd=1.5, col="gray",cex=0.8)
     text(30,.2, paste('T95:',round(mean(na.omit(T95)),1)),pos=4, col="black")
     abline(v=round(mean(na.omit(T95)),1), lty=2,lwd=1.5, col="black", cex=0.8)
   }
   return(data.frame(id=(unique(id)), 
                     #Tcrit.lci=round(Tcrit.ci[[1]],1),
                     Tcrit.mn=round(mean(na.omit(Tcrit)),1),  
                     #Tcrit.uci=round(Tcrit.ci[[2]],1),
                     
                     #T50.lci=round(T50.ci[[1]],1),
                     T50.mn=round(mean(na.omit(T50)),1),  
                     #T50.uci=round(T50.ci[[2]],1),
                     
                     #T95.lci=round(T95.ci[[1]],1),
                     T95.mn=round(mean(na.omit(T95)),1)  
                     #T95.uci=round(T95.ci[[2]],1)))
   ))
 })))
  detach(l1)
}

# Run Function
# Hot data only for now

T50_estimates_DURIN <- psiiht(Temperature = DURIN_TT_HOT$TempExp,
       FvFm = DURIN_TT_HOT$FvFm_Final,
       control.temp = 20,
       id = DURIN_TT_HOT$PlantSampleID,
       plot.est = T,
       boots = 3)


# Test on smaller subset
DURIN_TT_HOT_Lygra_O_VM <- DURIN_TT_HOT %>% 
  filter(siteID == "Lygra" & habitat == "Open" & species == "Vaccinium myrtillus")

T50_estimates_DURIN_LY_VM <- psiiht(Temperature = DURIN_TT_HOT_Lygra_O_VM$TempExp,
       FvFm = DURIN_TT_HOT_Lygra_O_VM$FvFm_Final,
       control.temp = 30,
       id = DURIN_TT_HOT_Lygra_O_VM$PlantSampleID2,
       plot.est = T,
       boots = 3)

```

```         
```

```{r}
# Version for the methods testing in Senja

# Adapted from Feeley and Perez 2020 Paper.
#Thanks for checking out this code for estimating heat tolerances using variable fluorescence!
#Be sure to the read the Read.me.txt file.
#Below, the "psiiht" function can calculate heat tolerances of PSII heat tolerance. 
#For this function, define the the columns with the 1) Temperature & 2) FvFm variables, then the 3) control temperature, the 4) id/unique factor that you 
#want to estimate the heat tolerances for, 5) indicate if you want to make plots, 6) the number of bootsrap iterations for estimating heat tolerances

# Function to estimate T50 of FvFm
# Note 'HTdf' is the template dataset used in the function by the original authors.

psiiht.2 <- function(Temperature, FvFm, control.temp, id, site, plot.est, boots) {
  l1 <- list(control.temp = control.temp, plot.est = plot.est, boots = boots)
  HTdf <- data.frame(Temperature = Temperature, FvFm = FvFm, id = id, site = site)
  
  return(do.call("rbind", by(HTdf, list(HTdf$id, HTdf$site), function(df) {
    Temperature <- df$Temperature
    FvFm <- df$FvFm
    id <- df$id
    site <- df$site
    
    # Get parameters for the logistics model
    custom_logit <- function(p) {
      p[p < 0.025] <- 0.025
      p[p > 0.975] <- 0.975
      return(log(p / (1 - p)))
    }
    logit_FvFm <- custom_logit(FvFm)
    cof <- coef(lm(logit_FvFm ~ Temperature))
    
    # Example: Adjusting initial values
    start_params <- list(
        theta1 = max(FvFm), 
        theta2 = cof[1], 
        theta3 = cof[2]
    )
    
    model_fit <- tryCatch({
      nls(FvFm ~ theta1 / (1 + exp(-(theta2 + theta3 * Temperature))),
          start = start_params, algorithm = "port",
          trace = FALSE, 
          lower = c(theta1 = 0, theta2 = -Inf, theta3 = -Inf), 
          upper = c(theta1 = 1, theta2 = Inf, theta3 = Inf),
          control = list(maxiter = 5000, tol = 1e-6))
    }, error = function(e) {
      message("Error while fitting model : ", e$message)
      return(NULL)
    })
    
   if (is.null(model_fit)) {
      # Check if id, siteID, etc are empty before including them in the data.frame
      if(length(unique(id)) > 0 && length(unique(site)) > 0) {
        return(data.frame(id = unique(id), site = unique(site), Tcrit.mn = NA, T50.mn = NA, T95.mn = NA))
      } else {
        return(NULL)  # Return NULL if any of the elements are empty
      }
   }
    
    print(coef(model_fit))
    # TODO: Make sure to check the sequential values depending on the experimental set-up.
    # We start at 20, rather than 23 as per the original. 
    y <- coef(model_fit)[1] / (1 + exp(-(coef(model_fit)[2] + coef(model_fit)[3] * seq(20, 66))))
    
    # Calculation of T50 and T95 values
    half <- mean(na.omit(FvFm[which(Temperature == control.temp)])) / 2  
    nine5 <- mean(na.omit(FvFm[which(Temperature == control.temp)])) * 0.05  
    
    predict.boot <- matrix(NA, 47, boots)
    T95 <- T50 <- Tcrit <- c()
    
    for(k in 1:boots) {
    srows <- sample(1:length(Temperature), length(Temperature), TRUE)
      boot_fit <- tryCatch({
        nls(FvFm[srows] ~ theta1 / (1 + exp(-(theta2 + theta3 * Temperature[srows]))),  
            start = start_params, algorithm = "port",
            trace = FALSE, 
            lower = c(theta1 = 0, theta2 = -Inf, theta3 = -Inf), 
            upper = c(theta1 = 1, theta2 = Inf, theta3 = Inf),
            control = list(maxiter = 1000, tol = 1e-3))
      }, error = function(e) {
        return(NULL)
      })
      
      if (!is.null(boot_fit)) {
        predict.boot[, k] <- coef(boot_fit)[1] / (1 + exp(-(coef(boot_fit)[2] + coef(boot_fit)[3] * seq(20, 66))))
        
        T95[k] <- (-log((coef(boot_fit)[1] / nine5) - 1) - coef(boot_fit)[2]) / coef(boot_fit)[3]
        T50[k] <- (-log((coef(boot_fit)[1] / half) - 1) - coef(boot_fit)[2]) / coef(boot_fit)[3]
        
        predict <- data.frame(x = seq(20, 66), 
                              y = coef(boot_fit)[1] / (1 + exp(-(coef(boot_fit)[2] + coef(boot_fit)[3] * seq(20, 66)))))
        df1 <- cbind(predict[-1, ], predict[-nrow(predict), ])[, c(3, 1, 4, 2)]
        df1$slp <- as.vector(apply(df1, 1, function(x) summary(lm((x[3:4]) ~ x[1:2]))[[4]][[2]]))
        slp.at.tcrit <- round(min(df1$slp), 3) * .15 
        fvfv.at.tcrit <- df1[which(abs(df1[which(df1[,1] < T50[k]),]$slp - slp.at.tcrit) == min(abs(df1[which(df1[,1] < T50[k]),]$slp - slp.at.tcrit))),][1, 3]
        Tcrit[k] <- (-log((coef(boot_fit)[1] / fvfv.at.tcrit) - 1) - coef(boot_fit)[2]) / coef(boot_fit)[3]
      } else {
        predict.boot[, k] <- NA
        T95[k] <- NA
        T50[k] <- NA
        Tcrit[k] <- NA
      }
    }
    
          FvFm.boot <- t(apply(predict.boot, 1, function(x) {quantile(x, c(0.025, 0.975), na.rm = TRUE)}))
    T50.ci <- quantile(T50, c(0.025, 0.975), na.rm = TRUE)
          FvFm.boot <- t(apply(predict.boot, 1, function(x) {quantile(x, c(0.025, 0.975), na.rm = TRUE)}))
    T50.ci <- quantile(T50, c(0.025, 0.975), na.rm = TRUE)
    
if(plot.est == TRUE) {
  plot(NULL, NULL, xlab = "Temperature", ylab = "Fv/Fm", xlim = c(20, 66), ylim = c(0, 0.9), bty = "l", lty = 2)
  
  # Add text to the top of the chart
  mtext(paste(id, site, sep = " "), side = 3, line = 1, cex = 1, font = 4)
  
  # Plot the points
  points(Temperature, FvFm, xlab = "Temperature", ylab = "Fv/Fm", xlim = c(20, 66), ylim = c(0, 0.85), bty = "l", lty = 2, pch = 5, col = "black")
  
  # Draw the curve if the data is available
  if(!is.null(y)) {
    lines(seq(20, 66), y, lwd = 1, col = "black")
  }
  
  # Add the curves of the bootstraps if they are available
  if(boots > 1) {
    if(!is.null(FvFm.boot)) {
      lines(seq(20, 66), FvFm.boot[, 1], lty = 3, col = "black")
      lines(seq(20, 66), FvFm.boot[, 2], lty = 3, col = "black")
    }
  }
  
  # Show the values of Tcrit, T50, and T95 if they are not NA
  if(!is.na(mean(na.omit(Tcrit)))) {
    text(30, 0, paste(round(mean(na.omit(Tcrit)), 1)), pos = 4, col = "light gray")
    abline(v = round(mean(na.omit(Tcrit)), 1), lty = 2, lwd = 1.5, col = "light gray", cex = 0.8)
  } else {
    text(30, 0, "NA", pos = 4, col = "light gray")
  }

  if(!is.na(mean(na.omit(T50)))) {
    text(30, .1, paste(round(mean(na.omit(T50)), 1)), pos = 4, col = "gray")
    abline(v = round(mean(na.omit(T50)), 1), lty = 2, lwd = 1.5, col = "gray", cex = 0.8)
  } else {
    text(30, .1, "NA", pos = 4, col = "gray")
  }

  if(!is.na(mean(na.omit(T95)))) {
    text(30, .2, paste(round(mean(na.omit(T95)), 1)), pos = 4, col = "black")
    abline(v = round(mean(na.omit(T95)), 1), lty = 2, lwd = 1.5, col = "black", cex = 0.8)
  } else {
    text(30, .2, "NA", pos = 4, col = "black")
  }
}
    
        
    if(length(unique(id)) > 0 && length(unique(site)) > 0) {
      return(data.frame(id = unique(id), 
                        site = unique(site),
                        Tcrit.mn = round(na.omit(Tcrit), 1),  
                        T50.mn = round(na.omit(T50), 1),  
                        T95.mn = round(na.omit(T95), 1)))
    } else {
      return(NULL)
    }
  })))
}


# Overall dataset
T50_estimates_DURIN.2 <- psiiht.2(Temperature = DURIN_TT_HOT$TempExp,
                                  site = DURIN_TT_HOT$siteID,
       FvFm = DURIN_TT_HOT$FvFm_Final,
       control.temp = 20,
       id = DURIN_TT_HOT$PlantSampleID,
       plot.est = T,
       boots = 1000)


# Overall dataset
T50_estimates_DURIN.2a <- psiiht.2(Temperature = DURIN_TT_HOT$TempExp,
                                  site = DURIN_TT_HOT$siteID,
       FvFm = DURIN_TT_HOT$FvFm_Final,
       control.temp = 30,
       id = DURIN_TT_HOT$PlantSampleID,
       plot.est = T,
       boots = 1000)




```

```{r}

## Overview with a boxplot
# Seperate by concatenated ID

T50_estimates_DURIN.2 <- T50_estimates_DURIN.2 %>% 
  separate(id, "_",
           into = c("siteID", "habitat", "species"),
           remove = FALSE)

# Define colour palettes
Habitat <- c("Forested" = "#083508", "Open" = "#589758")
Species <- c("Calluna vulgaris" = "#DF697E",
             "Empetrum nigrum" = "#404040",
             "Vaccinium myrtillus" = "#323284",
             "Vaccinium vitis-idaea" = "#D93137")

(HOT_boxplot_Methods <- ggplot(T50_estimates_DURIN.2, aes(x = siteID, y = T50.mn,
                                            fill = habitat)) +
                 geom_boxplot() +
                 facet_grid(. ~ species) +
                 scale_fill_manual(values = Habitat) +  
                 labs(x = "Habitat",
                      y = "T50 of Fv/Fm (ºC)",
                      fill = "Species") + 
                 theme_bw())

# # Version with the legend at the bottom or the plot
# (time_boxplot_Methods_base <- ggplot(Temp_Comb, aes(x = TempExpTime, y = T50.mn,
#                                             fill = TempExpTime)) +
#                  geom_boxplot() +
#                  facet_grid(.~id) +
#                  scale_fill_manual(values = c("#ea9999", "#e06666",
#                                               "#cc0000","#990000", "#660000")) +  
#                  labs(x = "Exposure Time (minutes)",
#                       y = "T50 of Fv/Fm (ºC)",
#                       fill = "Exposure Time (mins)") + 
#                  theme_bw() +
#                  theme(legend.position = "bottom"))
```

# Function to check the correlation between leaf number and Temperature Exposure

Leaf 1-8 should correspond to Temperature Exposures of 20, 30, 34, 38, 42, 46, 50, and 54 respectively

```{r function_check_leaf_temperature_alignment}
library(dplyr)

check_leaf_temperature_mapping <- function(df) {
  # Define correct leaf-temp mapping
  expected_mapping <- data.frame(
    leaf_nr = 1:8,
    expected_temp = c(20, 30, 34, 38, 42, 46, 50, 54)
  )

  df <- df %>%
    group_by(siteID, habitat, species, plant_nr) %>%
    
    # Join expected temperature mapping to ensure correct assignment
    left_join(expected_mapping, by = "leaf_nr") %>%
    
    # Check if recorded TempExp matches expected temperature for each leaf
    mutate(LeafTempAligned = ifelse(TempExp == expected_temp, "Aligned", "Misaligned")) %>%
    
    ungroup()

  return(df)
}


# Create Example Data with Misaligned Values
df_misaligned <- data.frame(
  siteID = rep(1:2, each = 24),
  habitat = rep(c("Open", "Forested"), each = 12, times = 2),
  species = rep(c("Species_A", "Species_B"), each = 6, times = 4),
  plant_nr = rep(1:3, times = 8),
  leaf_nr = rep(1:8, times = 3),
  TempExp = c(
    # Correct plant
    20, 30, 34, 38, 42, 46, 50, 54,
    # Misaligned (wrong temp for leaf 3 and missing leaf 6)
    20, 30, 38, 34, 42, NA, 50, 54,
    # Misaligned (missing leaf 2, wrong temp for leaf 4)
    20, NA, 34, 42, 38, 46, 50, 54
  ),
  FvFm = runif(24, min = 0, max = 0.9)
)

# Apply function to detect misalignments
df_checked <- check_leaf_temperature_mapping(df_misaligned)

# Print flagged rows
print(df_checked %>% filter(LeafTempAligned == "Misaligned"))

#################################
# Check on our dataframe
#################################
df_checked_DURIN_TT_HOT <- check_leaf_temperature_mapping(DURIN_TT_HOT)

# Print flagged rows
print(df_checked_DURIN_TT_HOT %>% 
      filter(LeafTempAligned == "Misaligned"))
# Or with only selected columns
print(subset(df_checked_DURIN_TT_HOT, LeafTempAligned == "Misaligned",
             select = c(envelope_ID, siteID, habitat, species, plant_nr, leaf_nr, leaf_age, sample,
                TempExp, FvFm_Final, remark, PlantSampleID2, expected_temp, LeafTempAligned)))


```

# Functions to correct FvFm estimates for calculating T50

In some cases the FvFm values go to zero, before getting higher again.

In some instances these may be biological (as discussed in the Canberra thermal tolerance workshop), or it could be a feature of using seperate leaves (and possibly from different plants) so not all values may be consistent.

In any case, these rebounds after the zero FvFm mean the function to calculate T50 cannot work properly as it can't create a logistics curve to calculate the 50% intercept point.

Here we try to a function to test what happens if we change all subsequent FvFm after the first instance of a zero into zeros as well.

```{r}
library(dplyr)

# filter_FvFm_by_leaf_nr <- function(df) {
#   df <- df %>%
#     arrange(siteID, habitat, species, plant_nr, leaf_nr) %>%  # Ensure correct ordering per plant
#   
#     group_by(siteID, habitat, species, plant_nr) %>%  # Group at the correct level
# 
#     # Identify first leaf_nr within each plant where FvFm_Final == 0
#     mutate(first_zero_leaf = case_when(
#       any(FvFm_Final == 0, na.rm = TRUE) ~ min(leaf_nr[FvFm_Final == 0], na.rm = TRUE),
#       TRUE ~ NA_real_  # If no zero exists, assign NA
#     )) %>%
# 
#     # Debugging output to confirm first_zero_leaf varies per plant
#     mutate(Debug_FirstZero = first_zero_leaf) %>%
# 
#     # Apply correction ONLY after the first zero is detected per unique plant
#     mutate(Filtered_FvFm = case_when(
#       is.na(first_zero_leaf) ~ FvFm_Final,  # If no zero exists, keep original values
#       leaf_nr <= first_zero_leaf ~ FvFm_Final,  # Preserve values before first zero
#       TRUE ~ 0  # Only zero values strictly occurring after failure
#     )) %>%
# 
#     ungroup() %>%
#     # Remove helper columns *only if they exist*
#     select_if(~ !any(names(.) %in% c("first_zero_leaf")))  
#   return(df)
# }


#################
# Loop version - works! (or maybe not....)
# filter_FvFm_by_leaf_nr_loop <- function(df) {
#   # Ensure correct ordering first of key variables
#   df <- df %>% arrange(siteID, habitat, species, plant_nr, leaf_nr)
# 
#   # Split dataset into individual plant ID's
#   df_list <- df %>%
#     group_split(siteID, habitat, species, plant_nr) %>%
#     lapply(function(plant_df) {
#       # Find first leaf where FvFm_Final == 0 within the specific plant
#       first_zero_leaf <- ifelse(any(plant_df$FvFm_Final == 0, na.rm = TRUE),
#                                 min(plant_df$leaf_nr[plant_df$FvFm_Final == 0], na.rm = TRUE), NA)
#       
#       # Apply correction ONLY after the first zero leaf per unique plant
#       plant_df$Filtered_FvFm <- ifelse(!is.na(first_zero_leaf) & plant_df$leaf_nr > first_zero_leaf, 0, plant_df$FvFm_Final)
#       
#       return(plant_df)
#     })
#   
#   # Recombine the list into a full dataset
#   df_corrected <- bind_rows(df_list)
# 
#   return(df_corrected)
# }


# New version as for some reason now the abvoe one doesn't work.
# Solution: Change df into a tibble before group_split to avoid the warning. 

filter_FvFm_by_leaf_nr_loop <- function(df) {
  # Ensure correct ordering first of key variables
  df <- df %>% arrange(siteID, habitat, species, plant_nr, leaf_nr)

  # Convert to standard tibble (fixes group_split warning)
  df <- as_tibble(df)

  # Split dataset into individual plant ID's
  df_list <- df %>%
    group_split(siteID, habitat, species, plant_nr) %>%
    lapply(function(plant_df) {
      # Find first leaf where FvFm_Final == 0 within the specific plant
      first_zero_leaf <- ifelse(any(plant_df$FvFm_Final == 0, na.rm = TRUE),
                                min(plant_df$leaf_nr[plant_df$FvFm_Final == 0], na.rm = TRUE), NA)
      
      # Apply correction ONLY after the first zero leaf per unique plant
      plant_df <- plant_df %>%
        mutate(Filtered_FvFm = ifelse(!is.na(first_zero_leaf) & leaf_nr > first_zero_leaf, 0, FvFm_Final),
               Zero_Adjustment_Status = ifelse(Filtered_FvFm == 0 & FvFm_Final != 0, "Zero-Adjusted", "Original"))

      return(plant_df)
    })
  
  # Recombine the list into a full dataset
  df_corrected <- bind_rows(df_list)

  return(df_corrected)
}




# Generate template dataset to test the function on
# df_non_logistic <- expand.grid(
#   siteID = 1:4,                      # 4 Sites
#   habitat = c("Open", "Forested"),    # Both habitats at each site
#   species = c("Species_A", "Species_B", "Species_C", "Species_D"),  # 4 species at each site/habitat
#   plant_nr = 1:3,                     # At least 3 plants per species/habitat/site
#   leaf_nr = 1:8                       # 8 leaves per plant (matching temperatures)
# ) %>%
#   arrange(siteID, habitat, species, plant_nr, leaf_nr) %>%
#   mutate(TempExp = rep(c(20, 30, 34, 38, 42, 46, 50, 54), times = nrow(.) / 8),
#          FvFm_Final = c(
#            # ✅ Expected monotonic decline
#            0.85, 0.80, 0.72, 0.50, 0.30, 0.15, 0, 0,
#            # ❌ Non-logistic: Drops too early, then incorrectly recovers
#            0.88, 0.84, 0.60, 0, 0.10, 0.25, 0.38, 0.20,
#            # ❌ Erratic fluctuations at higher temperatures
#            0.87, 0.82, 0.75, 0.60, 0.15, 0, 0.30, 0.22
#          ))


df_non_logistic <- expand.grid(
  siteID = 1:4,                      # 4 Sites
  habitat = c("Open", "Forested"),    # Both habitats at each site
  species = c("Species_A", "Species_B", "Species_C", "Species_D"),  # 4 species at each site/habitat
  plant_nr = 1:3,                     # At least 3 plants per species/habitat/site
  leaf_nr = 1:8                       # 8 leaves per plant (matching temperatures)
) %>%
  arrange(siteID, habitat, species, plant_nr, leaf_nr) %>%
  mutate(TempExp = rep(c(20, 30, 34, 38, 42, 46, 50, 54), times = nrow(.) / 8),
         FvFm_Final = rep(c(
           0.85, 0.80, 0.72, 0.50, 0.30, 0.15, 0, 0,   # Expected decline
           0.88, 0.84, 0.60, 0, 0.10, 0.25, 0.38, 0.20, # Non-logistic response
           0.87, 0.82, 0.75, 0.60, 0.15, 0, 0.30, 0.22  # Erratic fluctuations
         ), length.out = nrow(.)))  # Ensure it repeats correctly


# print(df_non_logistic)

#filtered_df <- filter_FvFm_by_leaf_nr(df_non_logistic)
filtered_df_loop <- filter_FvFm_by_leaf_nr_loop(df_non_logistic)

# Print results
#print(filtered_df)
print(filtered_df_loop)



# Visual Comparison Function: Pre vs. Post Filtering
# Now, let's add a function that overlays the original and filtered FvFm values to verify how the filtering is applied per plant.

# library(ggplot2)
# library(tidyr)

plot_FvFm_per_species <- function(df) {
  # Turn the dataset into long format, and rename the differenet metrics for FvFm based on the zero-adjusted code above
  df_long <- df %>%
    pivot_longer(cols = c("FvFm_Final", "Filtered_FvFm"), 
                 names_to = "DataType", values_to = "FvFm_Value")
 # plot FvFm value types per plant number
  ggplot(df_long, aes(x = TempExp, y = FvFm_Value, color = DataType, group = interaction(DataType, plant_nr))) +
    geom_point(size = 3, alpha = 0.8) +
    geom_line(size = 1.2) +
    facet_wrap(~ species, scales = "free") +  # Now grouped by species instead of plant_nr
    labs(title = "FvFm Comparison by Species",
         x = "Temperature Exposure (°C)", 
         y = "FvFm Value", 
         color = "Data Type") +
    theme_minimal(base_size = 14) +
    theme(legend.position = "bottom")
}

# Use on the filtered dataset
plot_FvFm_per_species(filtered_df_loop)


# Try this on the hot dataset
DURIN_TT_HOT_Filt_loop <- filter_FvFm_by_leaf_nr_loop(DURIN_TT_HOT)
check_FvFm_filter <- DURIN_TT_HOT_Filt_loop %>% select(PlantSampleID2, leaf_nr, FvFm_Final, Filtered_FvFm, Zero_Adjustment_Status)
plot_FvFm_per_species(DURIN_TT_HOT_Filt_loop)




#############################################
#### An interactive version of the plots to check FvFm status post-zero adjustements
# library(ggplot2)
# library(tidyr)
library(plotly)
library(htmlwidgets)

# Convert dataset for proper visualization
df_long <- filtered_df_loop %>%
  pivot_longer(cols = c("FvFm_Final", "Filtered_FvFm"), 
               names_to = "DataType", values_to = "FvFm_Value")

# Create ggplot visualization
interactive_plot <- ggplot(df_long, aes(x = TempExp, y = FvFm_Value, color = DataType, group = interaction(DataType, plant_nr))) +
  geom_point(size = 3, alpha = 0.8) +
  geom_line(size = 1.2) +
  facet_wrap(~ siteID + habitat + species, scales = "free") +
  labs(title = "Interactive FvFm Comparison",
       x = "Temperature Exposure (°C)", 
       y = "FvFm Value", 
       color = "Data Type") +
  theme_minimal(base_size = 14)

# Convert ggplot to interactive Plotly visualization
interactive_plotly <- ggplotly(interactive_plot)

# Save plot as an interactive HTML file
saveWidget(interactive_plotly, "FvFm_Comparison.html")


### Test on our dataset
# Convert dataset for proper visualization
df_long_test <- DURIN_TT_HOT_Filt_loop %>%
  pivot_longer(cols = c("FvFm_Final", "Filtered_FvFm"), 
               names_to = "DataType", values_to = "FvFm_Value")

# Create ggplot visualization
interactive_plot_DURIN <- ggplot(df_long_test, aes(x = TempExp, y = FvFm_Value,
                                             color = DataType, group = interaction(DataType, plant_nr))) +
  geom_point(size = 3, alpha = 0.8) +
  geom_line(size = 1.2) +
  facet_wrap(~ siteID + habitat + species, scales = "free") +
  labs(title = "Interactive FvFm Comparison",
       x = "Temperature Exposure (°C)", 
       y = "FvFm Value", 
       color = "Data Type") +
  theme_minimal(base_size = 14)

# Convert ggplot to interactive Plotly visualization
interactive_plotly_DURIN <- ggplotly(interactive_plot_DURIN)

# Save plot as an interactive HTML file
saveWidget(interactive_plotly_DURIN, "FvFm_Comparison_DURIN.html")




# Plot Initial Curves to check against the data-set when unadjusted.
DURIN_TT_HOT_Filt_loop %>% 
  # Filters out issues with hot plants with temps below lowest bath of 20
  filter(TempExp >= 20) %>% 
  # Various filters to remove samples associated with different projects below
  filter(proj_2024 != "Multiple") %>% 
  filter(proj_2024 == "TT Hot") %>% 
  filter(DroughtNet_plotID == "Not Applicable") %>% 
  # Create a plot of exposure temperature (TempExp) against the final FvFm measurements. 
  # In their Filtered from from the above code, i.e. values that went to zero FvFm should stay at zero..
  # Group by the individuals that were sampled, and colour by habitat
  ggplot(aes(x = TempExp, y= as.numeric(as.character(Filtered_FvFm)),
             group = PlantSampleID, color = habitat)) +
  # Add points
  geom_point()+
  # Add the smoothed line across the points, again coloured by habitat
  # One line per individual
  geom_smooth(method = "loess",
              se = FALSE, aes(group = PlantSampleID, color = habitat)) +
  # Make a differnet panel of the plot per site
  facet_grid(species ~ siteID)


# Plot Initial Curves
DURIN_TT %>% 
  # Filters out issues with hot plants with temps below lowest bath of 20
  filter(TempExp >= 20) %>% 
  # Various filters to remove samples associated with different projects below
  filter(proj_2024 != "Multiple") %>% 
  filter(proj_2024 == "TT Hot") %>% 
  filter(DroughtNet_plotID == "Not Applicable") %>% 
  # Create a plot of exposure temperature (TempExp) against the final FvFm measurements. 
  # Group by the individuals that were sampled, and colour by habitat
  ggplot(aes(x = TempExp, y= as.numeric(as.character(FvFm_Final)),
             group = PlantSampleID, color = habitat)) +
  # Add points
  geom_point()+
  # Add the smoothed line across the points, again coloured by habitat
  # One line per individual
  geom_smooth(method = "loess",
              se = FALSE, aes(group = PlantSampleID, color = habitat)) +
  # Make a differnet panel of the plot per site
  facet_grid(species ~ siteID)



# Function to plot the FvFm per plant
# NOTE- THis is the Filtered FvFm not the FvFm_Final which is the original
plot_FvFm_per_plant <- function(df) {
  df %>%
    ggplot(aes(x = TempExp, y = Filtered_FvFm, color = factor(plant_nr))) +
    geom_point(size = 3, alpha = 0.7) +
    geom_line(aes(group = plant_nr), linetype = "dashed", alpha = 0.5) +
    facet_wrap(~ species + habitat + siteID, scales = "free") +
    labs(title = "Thermal Response of FvFm per Plant",
         x = "Temperature Exposure (°C)",
         y = "FvFm Value (zero-adjusted)",
         color = "Plant Number") +
    theme_minimal() +
    theme(legend.position = "bottom")
}

# Apply function to visualize dataset
plot_FvFm_per_plant(DURIN_TT_HOT_Filt_loop)



# NOTE- Useing Filtered_FvFm not the original FvFm_Final
library(ggplot2)
library(dplyr)


# 2️⃣ Plot by Species, faceting by Site & Habitat
plot_FvFm_species <- function(df) {
  df %>%
    ggplot(aes(x = TempExp, y = Filtered_FvFm, color = factor(plant_nr))) +
    geom_point(size = 2, alpha = 0.7) +
    geom_line(aes(group = plant_nr), linetype = "dashed", alpha = 0.5) +
    facet_wrap(~ siteID + habitat + species, scales = "free") +
    labs(title = "Thermal Response by Species",
         x = "Temperature Exposure (°C)", y = "FvFm Value (zero-adjusted)") +
    theme_minimal() + theme(legend.position = "bottom")
}


plot_FvFm_species(DURIN_TT_HOT_Filt_loop)


# Plotly version
plot_FvFm_species_plotly <- function(df) {
  p <- ggplot(df, aes(x = TempExp, y = Filtered_FvFm, color = factor(plant_nr))) +
    geom_point(size = 2, alpha = 0.7) +
    geom_line(aes(group = plant_nr), linetype = "dashed", alpha = 0.5) +
    facet_wrap(~ siteID + habitat + species, scales = "free") +
    labs(title = "Thermal Response by Species",
         x = "Temperature Exposure (°C)", y = "FvFm Value (zero-adjusted)") +
    theme_minimal() + theme(legend.position = "bottom")
  
  # Convert ggplot to interactive plotly
  return(ggplotly(p))
}

# Saave to interactive plot that we can export to html
FvFm_species_plant_plot_interactive <- plot_FvFm_species_plotly(DURIN_TT_HOT_Filt_loop)

# Save plot as an interactive HTML file
saveWidget(FvFm_species_plant_plot_interactive, "DURIN_FvFm_Comparison_Species_Plant.html")


```

# Identify cases of FvFm declining then rebounding (not at zero)

```{r}
library(dplyr)

# Step 1: Example Dataset with Known Rebounds
# This dataset contains both normal trends and rebound cases, ensuring the function can correctly detect anomalies.
library(dplyr)

set.seed(42)  # Ensure reproducibility

# Generate a balanced dataset
df_example_rebounds <- expand.grid(
  siteID = 1:4,  
  habitat = c("Open", "Forested"),
  species = c("Species_A", "Species_B", "Species_C", "Species_D"),
  plant_nr = 1:3,
  leaf_nr = 1:8
) %>%
  arrange(siteID, habitat, species, plant_nr, leaf_nr) %>%
  mutate(TempExp = rep(c(20, 30, 34, 38, 42, 46, 50, 54), times = nrow(.) / 8),
         FvFm_Final = runif(nrow(.), min = 0.5, max = 0.9),  # Generate diverse values
         Filtered_FvFm = FvFm_Final)  # Initially set Filtered_FvFm to match original

# Introduce controlled zero values for testing correction logic
df_example_rebounds$FvFm_Final[c(40, 85, 200, 500)] <- 0  # Test filtering behavior

# Introduce rebound cases for anomaly detection
df_example_rebounds$FvFm_Final[c(45, 90, 220, 550)] <- df_example_rebounds$FvFm_Final[c(44, 89, 219, 549)] + 0.2  

print(df_example_rebounds)


library(dplyr)
# Alternative formulation of example dataset
# set.seed(42)  # Ensure reproducibility
# 
# df_example_rebounds <- expand.grid(
#   siteID = 1:4,  
#   habitat = c("Open", "Forested"),
#   species = c("Species_A", "Species_B", "Species_C", "Species_D"),
#   plant_nr = 1:3,
#   leaf_nr = 1:8
# ) %>%
#   arrange(siteID, habitat, species, plant_nr, leaf_nr) %>%
#   mutate(TempExp = rep(c(20, 30, 34, 38, 42, 46, 50, 54), times = nrow(.) / 8),
#          FvFm_Final = runif(nrow(.), min = 0, max = 0.9))  # Generate random thermal responses
# 
# # Introduce some rebound cases manually for testing
# df_example_rebounds$FvFm_Final[c(10, 25, 40, 80)] <- df_example_rebounds$FvFm_Final[c(9, 24, 39, 79)] + 0.2



# FvFm Rebound Function
# This function flags cases where FvFm rises after previously declining, identifying potential measurement inconsistencies.

flag_FvFm_rebounds <- function(df) {
  df <- df %>%
    arrange(siteID, habitat, species, plant_nr, TempExp) %>%
    group_by(siteID, habitat, species, plant_nr) %>%
    mutate(Rebound_FvFm_Flag = case_when(
      FvFm_Final > lag(FvFm_Final) & lag(FvFm_Final) < lag(lag(FvFm_Final)) ~ "Rebound_FvFm",
      TRUE ~ "Normal"
    )) %>%
    ungroup()
  
  return(df)
}

df_flagged_rebounds <- flag_FvFm_rebounds(df_example_rebounds)


# Step 3: Visualizing FvFm Rebound Events
# This plot highlights rebound events with special markers, making them easy to review.

plot_FvFm_rebounds <- function(df) {
  ggplot(df, aes(x = TempExp, y = FvFm_Final, color = Rebound_FvFm_Flag, group = plant_nr)) +
    geom_point(size = 3, alpha = 0.8) +
    geom_line(size = 1.2, aes(group = plant_nr), color = "gray") +  # Background trend lines
    facet_wrap(~ siteID + habitat + species + plant_nr, scales = "free") +
    labs(title = "Detection of FvFm Rebounds",
         x = "Temperature Exposure (°C)", 
         y = "FvFm Value", 
         color = "Rebound Event") +
    scale_color_manual(values = c("Normal" = "black", "Rebound_FvFm" = "red")) +  # Rebounds in red
    theme_minimal(base_size = 14) +
    theme(legend.position = "bottom")
}

plot_FvFm_rebounds(df_flagged_rebounds)



# Version adding a secondary flag column to highlight all samples that are, or occur post 'rebound'


flag_post_rebound_samples <- function(df) {
  df <- df %>%
    arrange(siteID, habitat, species, plant_nr, leaf_nr) %>%
    group_by(siteID, habitat, species, plant_nr) %>%

    # Identify first leaf in each plant sequence where a rebound is flagged
    mutate(first_rebound_leaf = ifelse(any(Rebound_FvFm_Flag == "Rebound_FvFm"), 
                                       min(leaf_nr[Rebound_FvFm_Flag == "Rebound_FvFm"], na.rm = TRUE), NA)) %>%

    # Apply labels based on rebound status
    mutate(Rebound_FvFm_Flag_plant = case_when(
      is.na(first_rebound_leaf) ~ "Normal",  # No rebound detected
      leaf_nr >= first_rebound_leaf ~ "Rebound FvFm and Onwards",  # Any leaf after the first rebound
      TRUE ~ "Normal"
    )) %>%

    ungroup() %>%
    select(-first_rebound_leaf)  # Remove helper column

  return(df)
}

# Ensure `Rebound_Flag` is created first
df_flagged_rebounds <- flag_FvFm_rebounds(df_example_rebounds)

# Apply post-rebound labeling
df_flagged_rebound_post <- flag_post_rebound_samples(df_flagged_rebounds)






# Plotly version
library(ggplot2)
library(plotly)
library(tidyr)
library(htmlwidgets)

# Reshape the dataset for plotting
df_long_rebounds <- df_flagged_rebounds %>%
  pivot_longer(cols = c("FvFm_Final", "Filtered_FvFm"), 
               names_to = "DataType", values_to = "FvFm_Value")

# Create ggplot visualization
interactive_plot_rebounds <- ggplot(df_long_rebounds, aes(x = TempExp, y = FvFm_Value, color = Rebound_FvFm_Flag, group = plant_nr)) +
  geom_point(size = 3, alpha = 0.8) +
  geom_line(size = 1.2, aes(group = plant_nr), color = "gray") +
  facet_wrap(~ siteID + habitat + species, scales = "free") +
  labs(title = "Interactive Detection of FvFm Rebounds",
       x = "Temperature Exposure (°C)", 
       y = "FvFm Value", 
       color = "Rebound Event") +
  scale_color_manual(values = c("Normal" = "black", "Rebound_FvFm" = "red")) +
  theme_minimal(base_size = 14) +
  theme(legend.position = "bottom")

# Convert ggplot to interactive Plotly visualization
interactive_plotly_rebounds <- ggplotly(interactive_plot_rebounds)

# Save plot as an interactive HTML file
saveWidget(interactive_plotly_rebounds, "FvFm_Rebound_Comparison.html")




# Table summary
generate_rebound_summary <- function(df) {
  df %>%
    filter(Rebound_FvFm_Flag == "Rebound_FvFm") %>%
    group_by(siteID, habitat, species, plant_nr) %>%
    summarise(
      Total_Rebounds = n(),
      Temperature_Range = paste(min(TempExp), "-", max(TempExp), "°C"),
      First_Rebound_Leaf = min(leaf_nr[Rebound_FvFm_Flag == "Rebound_FvFm"])
    ) %>%
    ungroup()
}

rebound_summary <- generate_rebound_summary(df_flagged_rebound_post)
print(rebound_summary)




#################################
# On the DURIN data
# Ensure `Rebound_Flag` is created first
DURIN_flagged_rebounds_DURIN <- flag_FvFm_rebounds(DURIN_TT_HOT_Filt_loop)

# Apply post-rebound labeling
DURIN_flagged_rebound_post <- flag_post_rebound_samples(DURIN_flagged_rebounds_DURIN)
# See plots
# Reshape the dataset for plotting
DURIN_long_rebounds <- DURIN_flagged_rebound_post %>%
  pivot_longer(cols = c("FvFm_Final", "Filtered_FvFm"), 
               names_to = "DataType", values_to = "FvFm_Value")

# Create ggplot visualization
DURIN_interactive_plot_rebounds <- ggplot(DURIN_long_rebounds, aes(x = TempExp, y = FvFm_Value, color = Rebound_FvFm_Flag, group = plant_nr)) +
  geom_point(size = 3, alpha = 0.8) +
  geom_line(size = 1.2, aes(group = plant_nr), color = "gray") +
  facet_wrap(~ siteID + habitat + species, scales = "free") +
  labs(title = "Interactive Detection of FvFm Rebounds",
       x = "Temperature Exposure (°C)", 
       y = "FvFm Value", 
       color = "Rebound Event") +
  scale_color_manual(values = c("Normal" = "black", "Rebound_FvFm" = "red")) +
  theme_minimal(base_size = 14) +
  theme(legend.position = "bottom")

# Convert ggplot to interactive Plotly visualization
DURIN_interactive_plot_rebounds <- ggplotly(DURIN_interactive_plot_rebounds)

# Save plot as an interactive HTML file
saveWidget(DURIN_interactive_plot_rebounds, "DURIN_FvFm_Rebound_Comparison.html")


```

```{r}
# Overall dataset
T50_estimates_DURIN.2a <- psiiht.2(Temperature = DURIN_TT_HOT_Filt_loop$TempExp,
                                  site = DURIN_TT_HOT_Filt_loop$siteID,
       FvFm = DURIN_TT_HOT_Filt_loop$Filtered_FvFm,
       control.temp = 20,
       id = DURIN_TT_HOT_Filt_loop$PlantSampleID,
       plot.est = T,
       boots = 50)


(HOT_boxplot_Methods.filt <- ggplot(T50_estimates_DURIN.2a, aes(x = siteID, y = T50.mn,
                                            fill = habitat)) +
                 geom_boxplot() +
                 facet_grid(. ~ species) +
                 scale_fill_manual(values = Habitat) +  
                 labs(x = "Habitat",
                      y = "T50 of Fv/Fm (ºC)",
                      fill = "Species") + 
                 theme_bw())

```

```{r}
library(ggplot2)
# An alternative version that is designed to store values even if function crashes
psiiht.2_safe <- function(Temperature, FvFm, control.temp, id, site, plot.est, boots, save_plots = TRUE, plot_dir = "T50_plots") {
  l1 <- list(control.temp = control.temp, plot.est = plot.est, boots = boots)
  HTdf <- data.frame(Temperature = Temperature, FvFm = FvFm, id = id, site = site)

  # Create results dataframe
  results_df <- data.frame(id = character(), site = character(), Tcrit.mn = numeric(), 
                           T50.mn = numeric(), T95.mn = numeric(), failure_reason = character(), 
                           stringsAsFactors = FALSE)

  # Create directory for saving plots
  if (save_plots && !dir.exists(plot_dir)) dir.create(plot_dir)

  by(HTdf, list(HTdf$id, HTdf$site), function(df) {
    Temperature <- df$Temperature
    FvFm <- df$FvFm
    id <- df$id
    site <- df$site

    # Failure tracking
    failure_reason <- NA

    # Check for missing values
    if (any(is.na(Temperature)) | any(is.na(FvFm))) {
      failure_reason <- "Missing values in Temperature or FvFm"
    }

    # Ensure enough samples exist
    if (nrow(df) < 3) {
      failure_reason <- "Not enough samples for logistic model"
    }

    # Detect erratic distribution (e.g., increasing instead of declining)
    if (!all(diff(FvFm) <= 0)) {
      failure_reason <- "FvFm values fluctuate instead of decreasing"
    }

    # Get logistic model parameters
    custom_logit <- function(p) {
      p[p < 0.025] <- 0.025
      p[p > 0.975] <- 0.975
      return(log(p / (1 - p)))
    }
    
    logit_FvFm <- custom_logit(FvFm)
    cof <- tryCatch(coef(lm(logit_FvFm ~ Temperature)), error = function(e) NULL)

    # Handle coefficient issues
    if (is.null(cof)) {
      failure_reason <- "Failed to compute logistic regression coefficients"
    }

    # Adjust initial values
    start_params <- list(
      theta1 = max(FvFm), 
      theta2 = cof[1], 
      theta3 = cof[2]
    )

    model_fit <- tryCatch({
      nls(FvFm ~ theta1 / (1 + exp(-(theta2 + theta3 * Temperature))),
          start = start_params, algorithm = "port",
          trace = FALSE, 
          lower = c(theta1 = 0, theta2 = -Inf, theta3 = -Inf), 
          upper = c(theta1 = 1, theta2 = Inf, theta3 = Inf),
          control = list(maxiter = 5000, tol = 1e-6))
    }, error = function(e) {
      failure_reason <- paste("NLS model fitting failed:", e$message)
      return(NULL)
    })

    # Store results even if model fails
    if (!is.null(model_fit)) {
      T50_estimate <- (-log((coef(model_fit)[1] / (max(FvFm) / 2)) - 1) - coef(model_fit)[2]) / coef(model_fit)[3]
      results_df <<- rbind(results_df, data.frame(id = unique(id), site = unique(site), 
                                                  Tcrit.mn = round(T50_estimate, 1), 
                                                  T50.mn = round(T50_estimate, 1), 
                                                  T95.mn = round(T50_estimate, 1), 
                                                  failure_reason = NA))
    } else {
      results_df <<- rbind(results_df, data.frame(id = unique(id), site = unique(site), 
                                                  Tcrit.mn = NA, T50.mn = NA, T95.mn = NA, 
                                                  failure_reason = failure_reason))
    }

    # Plot generation
    if (plot.est) {
      p <- ggplot(df, aes(x = Temperature, y = FvFm)) +
        geom_point(color = "black") +
        geom_smooth(method = "nls", formula = y ~ theta1 / (1 + exp(-(theta2 + theta3 * x))),
                    method.args = list(start = start_params), se = FALSE, color = "blue") +
        labs(title = paste("Thermal Response Curve:", unique(id)),
             x = "Temperature (°C)", y = "FvFm") +
        theme_minimal()

      # Save plot if enabled
      if (save_plots) {
        ggsave(filename = paste0(plot_dir, "/", unique(id), "_T50_plot.png"), plot = p, width = 6, height = 4)
      }

      print(p)  # Show plot during execution
    }
  })

  return(results_df)  # Final results table with errors logged
}

# Run the function with diagnostics and saved plots
T50_estimates_DURIN.2_safe <- psiiht.2_safe(Temperature = DURIN_TT_HOT_Filt_loop$TempExp,
                                       site = DURIN_TT_HOT_Filt_loop$siteID,
                                       FvFm = DURIN_TT_HOT_Filt_loop$FvFm_Final,
                                       control.temp = 30,
                                       id = DURIN_TT_HOT_Filt_loop$PlantSampleID,
                                       plot.est = TRUE,
                                       boots = 1000,
                                       save_plots = TRUE)

# Print saved results table
print(T50_estimates_DURIN.2_safe)


T50_estimates_DURIN.2_safe <- T50_estimates_DURIN.2_safe %>% 
  separate(id, "_",
           into = c("siteID", "habitat", "species"),
           remove = FALSE)


(T50_plot <- ggplot(T50_estimates_DURIN.2_safe, aes(x = siteID, y = T50.mn,
                                            fill = habitat)) +
                 geom_boxplot() +
                 facet_grid(. ~ species) +
                 scale_fill_manual(values = Habitat) +  
                 labs(x = "Habitat",
                      y = "T50 of Fv/Fm (ºC)",
                      fill = "Species") + 
                 theme_bw())

```

```{r}
psiiht.3 <- function(Temperature, FvFm, control.temp, habitat, siteID, species, plot.est, boots) {
  # Create dataframe with relevant variables
  HTdf <- data.frame(Temperature = Temperature, FvFm = FvFm, habitat = habitat, siteID = siteID, species = species)

  return(do.call("rbind", by(HTdf, list( HTdf$siteID, HTdf$habitat, HTdf$species), function(df) {
    Temperature <- df$Temperature
    FvFm <- df$FvFm
    habitat <- df$habitat
    siteID <- df$siteID
    species <- df$species

    # Logistic model parameters
    custom_logit <- function(p) {
      p[p < 0.025] <- 0.025
      p[p > 0.975] <- 0.975
      return(log(p / (1 - p)))
    }
    
    logit_FvFm <- custom_logit(FvFm)
    cof <- coef(lm(logit_FvFm ~ Temperature))

    # Initial model parameters
    start_params <- list(
        theta1 = max(FvFm), 
        theta2 = cof[1], 
        theta3 = cof[2]
    )

    model_fit <- tryCatch({
      nls(FvFm ~ theta1 / (1 + exp(-(theta2 + theta3 * Temperature))),
          start = start_params, algorithm = "port",
          trace = FALSE, 
          lower = c(theta1 = 0, theta2 = -Inf, theta3 = -Inf), 
          upper = c(theta1 = 1, theta2 = Inf, theta3 = Inf),
          control = list(maxiter = 5000, tol = 1e-6))
    }, error = function(e) {
      message("Error while fitting model: ", e$message)
      return(NULL)
    })

    if (is.null(model_fit)) {
      return(data.frame(siteID = unique(siteID),
                        habitat = unique(habitat),
                        species = unique(species),
                        T50.mn = NA, T95.mn = NA, Tcrit.mn = NA))
    }

    # Generate predictions using the fitted model
    y <- coef(model_fit)[1] / (1 + exp(-(coef(model_fit)[2] + coef(model_fit)[3] * seq(20, 66))))

    # T50 and T95 calculations
    half <- mean(na.omit(FvFm[which(Temperature == control.temp)])) / 2  
    nine5 <- mean(na.omit(FvFm[which(Temperature == control.temp)])) * 0.05  

    predict.boot <- matrix(NA, 47, boots)
    T95 <- T50 <- Tcrit <- numeric(boots)

    for (k in 1:boots) {
      srows <- sample(1:length(Temperature), length(Temperature), TRUE)
      
      boot_fit <- tryCatch({
        nls(FvFm[srows] ~ theta1 / (1 + exp(-(theta2 + theta3 * Temperature[srows]))),
            start = start_params, algorithm = "port",
            trace = FALSE, 
            lower = c(theta1 = 0, theta2 = -Inf, theta3 = -Inf), 
            upper = c(theta1 = 1, theta2 = Inf, theta3 = Inf),
            control = list(maxiter = 1000, tol = 1e-3))
      }, error = function(e) {
        return(NULL)
      })

      if (!is.null(boot_fit)) {
        predict.boot[, k] <- coef(boot_fit)[1] / (1 + exp(-(coef(boot_fit)[2] + coef(boot_fit)[3] * seq(20, 66))))
        T95[k] <- (-log((coef(boot_fit)[1] / nine5) - 1) - coef(boot_fit)[2]) / coef(boot_fit)[3]
        T50[k] <- (-log((coef(boot_fit)[1] / half) - 1) - coef(boot_fit)[2]) / coef(boot_fit)[3]
      } else {
        T95[k] <- NA
        T50[k] <- NA
        Tcrit[k] <- NA
      }
    }

    # Confidence intervals
    T50.ci <- quantile(T50, c(0.025, 0.975), na.rm = TRUE)
    T95.ci <- quantile(T95, c(0.025, 0.975), na.rm = TRUE)

    # Plot results if requested
    if (plot.est == TRUE) {
      plot(NULL, NULL, xlab = "Temperature", ylab = "Fv/Fm", xlim = c(20, 66), ylim = c(0, 0.9), bty = "l", lty = 2)
      mtext(paste(habitat, siteID, species, sep = " "), side = 3, line = 1, cex = 1, font = 4)
      points(Temperature, FvFm, pch = 5, col = "black")

      if (!is.null(y)) {
        lines(seq(20, 66), y, lwd = 1, col = "black")
      }

      if (boots > 1) {
        FvFm.boot <- t(apply(predict.boot, 1, function(x) {quantile(x, c(0.025, 0.975), na.rm = TRUE)}))
        lines(seq(20, 66), FvFm.boot[, 1], lty = 3, col = "black")
        lines(seq(20, 66), FvFm.boot[, 2], lty = 3, col = "black")
      }

      if (!is.na(mean(na.omit(T50)))) {
        text(30, .1, paste(round(mean(na.omit(T50)), 1)), pos = 4, col = "gray")
        abline(v = round(mean(na.omit(T50)), 1), lty = 2, lwd = 1.5, col = "gray")
      }
    }

    # Return results
    return(data.frame( siteID = unique(siteID),
                       habitat = unique(habitat),
                       species = unique(species),
                      T50.mn = round(mean(na.omit(T50)), 1),  
                      T95.mn = round(mean(na.omit(T95)), 1),  
                      Tcrit.mn = round(mean(na.omit(Tcrit)), 1)))
  })))
}




# Assess data
T50_estimates_DURIN.3 <- psiiht.3(Temperature = DURIN_TT_HOT$TempExp,
                                  siteID = DURIN_TT_HOT$siteID,
                                  habitat = DURIN_TT_HOT$habitat,
                                  species = DURIN_TT_HOT$species,
       FvFm = DURIN_TT_HOT$FvFm_Final,
       control.temp = 30,
       #id = DURIN_TT_HOT$PlantSampleID,
       plot.est = T,
       boots = 5)
```
