---
title: "DURIN TOMST Microclimate Data"
format: html
editor: visual
author: "Sonya R. Geange"
date: today
---

## Importing and filtering data

Installing and loading all the packages required to run the code

```{r install packages}

#install.packages("tidyverse")
#install.packages("lubridate")
#install.packages("dataDownloader")
#install.packages("purrrlyr")
#install.packages("readxl")
#install.packages("fs")
#install.packages("purrr")
#install.packages("readr")
#install.packages("ggplot2")
#install.packages("lme4")
#install.packages("stargazer")
#install.packages("myClim")
```

Load all the packages needed

```{r load_packages}

library(tidyverse)
library(lubridate)
# library(dataDownloader)
library(purrrlyr)
library(readxl)
library(readr)
library(tidyverse)
library(fs)
library(purrr)
library(readr)
library(ggplot2)
library(lme4)
library(stargazer)
library(emmeans)
library(see)
library(svglite)
library(officer)
library(lmerTest)
library(officer)
library(myClim)
library(cowplot)
library(ggpubr)
```

### Importing metadata from the TOMST loggers files

```{r data import}
metadata <- read_delim("C:/Users/sge010/OneDrive - University of Bergen/Documents/DURIN/Durin_data_project/data/raw_data/TOMST_Logger/DURIN_TOSMT_2024_metadata.csv", delim = ";", escape_double = FALSE, trim_ws = TRUE, col_names=TRUE, col_types=c("c","c","n","c","c","c","c","c","n","n","D","D"))

# Seems some extra rows and columns are brought in, remove these
metadata <- metadata[1:190, 1:13]

# Find the "loggerID" and "plotID" columns
loggerIDColumn <- which(names(metadata) == "loggerID")
plotIDColumn <- which(names(metadata) == "plotID")

# Sort data in ascending order according to the "loggerID" column
metadata <- metadata |> 
  arrange(metadata[, loggerIDColumn])

# List files in folder
folder_path <- "C:/Users/sge010/OneDrive - University of Bergen/Documents/DURIN/Durin_data_project/data/raw_data/TOMST_Logger/TOMST_microclimate_loggers/Loggers_All/"
#SO/Sogndal_31.05.2024"      


# List files in folder
file_list <- list.files(folder_path)

# Create a data frame to store file paths, localities and data formats
path_data <- data.frame(
  path = file.path(folder_path, file_list),
  
  loggerID  = as.integer(substr(file_list, 6, 13)), # Extract digits from file name
  data_format = "TOMST" # Set data format as "TOMST"
  )

# Create a data frame to store only useful data
active_metadata <- data.frame(loggerID = path_data$loggerID) |> 
   left_join(metadata, by = join_by(loggerID)) 
path_data <- path_data |> 
   mutate(locality_id = as.character(active_metadata$loggerID)) |> 
   mutate(serial_number = as.character(active_metadata$loggerID)) |> 
            select(path,locality_id,data_format,serial_number) # put column in the right order


localities_data <- data.frame(
    locality_id = as.character(active_metadata$loggerID),
    plot_id = active_metadata$plotID,
    site = active_metadata$site,
    habitat = active_metadata$land_type )
  
# Write data to a csv file
#write_csv(path_data, "chemin_vers_votre_dossier/Path.xlsx", row.names = FALSE)
#write_csv(localities_data, "chemin_vers_votre_dossier/Localities.xlsx", row.names = FALSE)

```

### Importing all the data

```{r}
tms <- mc_read_data(files_table = path_data,
                      localities_table = localities_data,
                      silent = TRUE, clean = TRUE)

#tms.verif <- tms |> 
#  mc_reshape_long()

#> Warning in mc_prep_clean(tms.m, silent = T): MyClim object is already cleaned.
#> Repeated cleaning overwrite cleaning informations.
tms.info1 <- mc_info_clean(tms) # call cleaning log

tms <- mc_join(tms)
tms.info2 <- mc_info_clean(tms)

```

### Cropping with myClim package and vector

```{r}

# Start and end are list 
#start <-  as.POSIXct(as.Date(active_metadata$date_in, format = "%d/%m/%Y"), tz = "UTC")
start_t <- as.POSIXct(as.Date("01/08/2023", format = "%d/%m/%Y"), tz = "UTC")
# we choose this date to have the same data for all TMS loggers
end_t <- as.POSIXct(as.Date("15/06/2024", format = "%d/%m/%Y"), tz = "UTC")
#end <- as.POSIXct(as.Date(active_metadata$date_collect, format = "%d/%m/%Y"), tz = "UTC")
localities = as.vector(names(tms$localities))


tms <- mc_prep_crop(tms, start = start_t, end = NULL, localities)
tms.info <- mc_info_clean(tms)

# Seems to be some issue with 'steps', locality 95221022 has 60, whilst others have 900. Remove?
# Also, earlier we had some instances of loggers with old and new, so remove those too for now
# 95221030 Problematic interval: 2024-05-28 UTC--2024-06-05 09:15:00 UTC
# 95221057 Problematic interval: 2024-05-28 UTC--2024-06-05 09:15:00 UTC
# 95221153 Problematic interval: 2024-05-28 UTC--2024-06-05 09:15:00 UTC
tms.filt <- mc_filter(tms, localities = "95221022", reverse = T)
  tms.filt <- mc_filter(tms.filt, localities = "95221030", reverse = T)
  tms.filt <- mc_filter(tms.filt, localities =  "95221057", reverse = T)
  tms.filt <- mc_filter(tms.filt, localities = "95221153", reverse = T)
```

## Data extraction

## ALL

```{r calculation for all the sites}

## calculate virtual sensor VWC from raw TMS moisture signal
tms.all.calc <- mc_calc_vwc(tms.filt, soiltype = "universal") #reseach on soiltype and TOMST VERSION

## virtual sensor with growing and freezing degree days
tms.all.calc <- mc_calc_gdd(tms.all.calc, sensor = "TMS_T3",)
tms.all.calc <- mc_calc_fdd(tms.all.calc, sensor = "TMS_T3")

## virtual sensor to estimate snow presence from 2 cm air temperature 
tms.all.calc <- mc_calc_snow(tms.all.calc, sensor = "TMS_T2")

## summary data.frame of snow estimation
tms.snow <- tms.all.calc |> 
  mc_calc_snow_agg() 
tms.snow$locality_id = as.numeric(tms.snow$locality_id)
tms.snow <- active_metadata |> 
  select("locality_id" = loggerID , "siteID" = site,"habitat" = land_type) |> 
  filter(habitat %in% c("Open", "Forested")) |> 
  left_join(tms.snow) |> 
  group_by(siteID, habitat) |> 
  summarise(mean_snow_days = mean(snow_days),
            first_day = mean.Date(first_day, na.rm = TRUE),
            last_day = mean.Date(last_day, na.rm = TRUE))



 
temp_env_all  <- tms.filt |> 
  mc_env_temp(period = "week", min_coverage = 0.9) |> 
  filter(value >= -40 & value <= 50)

temp_env_all_daily  <- tms.filt |> 
  mc_env_temp(period = "day", min_coverage = 0.9) |> 
  filter(value >= -50 & value <= 50)

moist_env_all <- mc_env_moist(tms.all.calc, period = "week", min_coverage = 0.9)
snow_all <- tms.all.calc |> 
    mc_reshape_long() |> 
  filter(sensor_name == "snow")

```

Plot comparison

```{r}
active_metadata$site <- factor(active_metadata$site, levels = c("Senja", "Kautokeino", "Lygra","Sogndal"))

temp.soil.mean.all <- temp_env_all |> 
  filter(sensor_name == "T.soil_8_cm.mean") |> #T.air_15_cm.mean
    select("loggerID" = locality_id, sensor_name, datetime, "temp_mean_soil" = value) 
temp.soil.mean.all$loggerID = as.numeric(temp.soil.mean.all$loggerID)


temp.soil.mean.all <- active_metadata |> 
  select(loggerID, site, "habitat" = land_type) |> 
  left_join(temp.soil.mean.all) |> 
  filter(habitat %in% c("Open", "Forested")) |> 
  group_by(site, datetime) |>
  #group_by(site, datetime) |>
  summarise(mean.temp.soil = mean(temp_mean_soil, na.rm = TRUE)) |>  
  ungroup()

temp.soil.mean.all.hab <- temp_env_all |> 
  filter(sensor_name == "T.soil_8_cm.mean") |> #T.air_15_cm.mean
    select("loggerID" = locality_id, sensor_name, datetime, "temp_mean_soil" = value) 
temp.soil.mean.all.hab$loggerID = as.numeric(temp.soil.mean.all.hab$loggerID)

temp.soil.mean.all.hab <- active_metadata |> 
  select(loggerID, site, "habitat" = land_type) |> 
  left_join(temp.soil.mean.all.hab) |> 
  filter(habitat %in% c("Open", "Forested")) |> 
  group_by(site, habitat, datetime) |>
  #group_by(site, datetime) |>
  summarise(mean.temp.soil.hab = mean(temp_mean_soil, na.rm = TRUE)) |>  
  ungroup()
  # mutate(mean.temp.hab = replace(mean.temp.hab, is.nan(mean.temp.hab), -10)) 



temp.mean.all <- temp_env_all |> 
  filter(sensor_name == "T.air_15_cm.mean") |> 
    select("loggerID" = locality_id, sensor_name, datetime, "temp_mean_air" = value) 
temp.mean.all$loggerID = as.numeric(temp.mean.all$loggerID)

temp.mean.all <- active_metadata |> 
  select(loggerID, site, "habitat" = land_type) |> 
  left_join(temp.mean.all) |> 
  filter(habitat %in% c("Open", "Forested")) |> 
  group_by(site, datetime) |>
  #group_by(site, datetime) |>
  summarise(mean.temp.air = mean(temp_mean_air, na.rm = TRUE)) |>  
  ungroup()

temp.mean.all.hab <- temp_env_all |> 
  filter(sensor_name == "T.air_15_cm.mean") |> #T.air_15_cm.mean
    select("loggerID" = locality_id, sensor_name, datetime, "temp_mean_air" = value) 
temp.mean.all.hab$loggerID = as.numeric(temp.mean.all.hab$loggerID)

temp.mean.all.hab <- active_metadata |> 
  select(loggerID, site, "habitat" = land_type) |> 
  left_join(temp.mean.all.hab) |> 
  filter(habitat %in% c("Open", "Forested")) |> 
  group_by(site, habitat, datetime) |>
  #group_by(site, datetime) |>
  summarise(mean.temp.air.hab = mean(temp_mean_air, na.rm = TRUE)) |>  
  ungroup()
  # mutate(mean.temp.hab = replace(mean.temp.hab, is.nan(mean.temp.hab), -10)) 

temp.min.all.hab <- temp_env_all |> 
  filter(sensor_name == "T.air_15_cm.min5p") |> #T.air_15_cm.mean
    select("loggerID" = locality_id, sensor_name, datetime, "temp_min_air" = value) 
temp.min.all.hab$loggerID = as.numeric(temp.min.all.hab$loggerID)

temp.min.all.hab <- active_metadata |> 
  select(loggerID, site, "habitat" = land_type) |> 
  left_join(temp.min.all.hab) |> 
  filter(habitat %in% c("Open", "Forested")) |> 
  group_by(site, habitat, datetime) |>
  #group_by(site, datetime) |>
  summarise(min.temp.air.hab = mean(temp_min_air, na.rm = TRUE)) |>  
  ungroup()

temp.max.all.hab <- temp_env_all |> 
  filter(sensor_name == "T.air_15_cm.max95p") |> #T.air_15_cm.mean
    select("loggerID" = locality_id, sensor_name, datetime, "temp_max_air" = value) 
temp.max.all.hab$loggerID = as.numeric(temp.max.all.hab$loggerID)

temp.max.all.hab <- active_metadata |> 
  select(loggerID, site, "habitat" = land_type) |> 
  left_join(temp.max.all.hab) |> 
  filter(habitat %in% c("Open", "Forested")) |> 
  group_by(site, habitat, datetime) |>
  #group_by(site, datetime) |>
  summarise(max.temp.air.hab = mean(temp_max_air, na.rm = TRUE)) |>  
  ungroup()

moist.mean.all <- moist_env_all |> 
  filter(sensor_name == "VWC.soil_0_15_cm.mean") |> 
  select("loggerID" = locality_id, sensor_name, datetime, "moist_mean" = value) 
moist.mean.all$loggerID = as.numeric(moist.mean.all$loggerID) 
moist.mean.all <- moist.mean.all|> 
  left_join(temp.soil.mean.all) |> 
  mutate(moist_mean = if_else(mean.temp.soil < 1, NA, moist_mean))

moist.mean.all <- active_metadata |> 
  select(loggerID, site, "habitat" = land_type) |> 
  left_join(moist.mean.all) |> 
  filter(habitat %in% c("Open", "Forested")) |> 
  group_by(site, datetime) |> 
  #group_by(site, datetime) |> 
  summarise(mean.moist = mean(moist_mean, na.rm = TRUE)) |> 
  ungroup()# |> 


moist.mean.all.hab <- moist_env_all |> 
  filter(sensor_name == "VWC.soil_0_15_cm.mean") |> 
  select("loggerID" = locality_id, sensor_name, datetime, "moist_mean" = value) 
moist.mean.all.hab$loggerID = as.numeric(moist.mean.all.hab$loggerID)
moist.mean.all.hab <- moist.mean.all.hab|> 
  left_join(temp.soil.mean.all.hab) |> 
  mutate(moist_mean = if_else(mean.temp.soil.hab < 1, NA, moist_mean))

moist.mean.all.hab <- active_metadata |> 
  select(loggerID, site, "habitat" = land_type) |> 
  left_join(moist.mean.all.hab) |> 
  filter(habitat %in% c("Open", "Forested")) |> 
  group_by(site, habitat, datetime) |> 
  #group_by(site, datetime) |> 
  summarise(mean.moist.hab = mean(moist_mean, na.rm = TRUE)) |> 
  ungroup()

moist.min.all.hab <- moist_env_all |> 
  filter(sensor_name == "VWC.soil_0_15_cm.5p") |> 
  select("loggerID" = locality_id, sensor_name, datetime, "moist_min" = value) 
moist.min.all.hab$loggerID = as.numeric(moist.min.all.hab$loggerID)
moist.min.all.hab <- moist.min.all.hab|> 
  left_join(temp.soil.mean.all.hab) |> 
  mutate(moist_min = if_else(mean.temp.soil.hab < 1, NA, moist_min))

moist.min.all.hab <- active_metadata |> 
  select(loggerID, site, "habitat" = land_type) |> 
  left_join(moist.min.all.hab) |> 
  filter(habitat %in% c("Open", "Forested")) |> 
  group_by(site, habitat, datetime) |> 
  #group_by(site, datetime) |> 
  summarise(min.moist.hab = mean(moist_min, na.rm = TRUE)) |> 
  ungroup()

moist.max.all.hab <- moist_env_all |> 
  filter(sensor_name == "VWC.soil_0_15_cm.95p") |> 
  select("loggerID" = locality_id, sensor_name, datetime, "moist_max" = value) 
moist.max.all.hab$loggerID = as.numeric(moist.max.all.hab$loggerID)
moist.max.all.hab <- moist.max.all.hab|> 
  left_join(temp.soil.mean.all.hab) |> 
  mutate(moist_max = if_else(mean.temp.soil.hab < 1, NA, moist_max))

moist.max.all.hab <- active_metadata |> 
  select(loggerID, site, "habitat" = land_type) |> 
  left_join(moist.max.all.hab) |> 
  filter(habitat %in% c("Open", "Forested")) |> 
  group_by(site, habitat, datetime) |> 
  #group_by(site, datetime) |> 
  summarise(max.moist.hab = mean(moist_max, na.rm = TRUE)) |> 
  ungroup()


# Plot the data
f_line_macro_1 <-ggplot(moist.mean.all, aes(x = datetime, y = mean.moist, color = site)) +
  geom_line()+
  labs(#title = "Average Soil Moisture Over Time",
       x = "Date",
       y = "Weekly Average Soil Volumetric Water Content",
       color = "Site") +
  theme(
    axis.title.x = element_text(size = 14), 
    axis.title.y = element_text(size = 14), 
    axis.text.x = element_text(size = 12), 
    axis.text.y = element_text(size = 12), 
    plot.subtitle = element_text(size = 14),
    plot.caption = element_text(size = 14),
    legend.text = element_text(size = 12)
  ) +
   theme(legend.position = "bottom") +
  theme_classic()

print(f_line_macro_1)

f_line_macro_2 <-ggplot(temp.mean.all, aes(x = datetime, y = mean.temp.air, color = site)) +
  geom_line()+
  labs(#title = "Average air temperature Over Time",
       x = "Date",
       y = "Weekly Average Air (15cm) Temperature (°C)",
       color = "Site")+
  theme(
    axis.title.x = element_text(size = 14), 
    axis.title.y = element_text(size = 14), 
    axis.text.x = element_text(size = 12), 
    axis.text.y = element_text(size = 12), 
    plot.subtitle = element_text(size = 14),
    plot.caption = element_text(size = 14),
    legend.text = element_text(size = 12)
  ) +
   theme(legend.position = "bottom") +
  theme_classic()

print(f_line_macro_2)

 plot_grid(f_line_macro_1, f_line_macro_2, ncol = 1)
 
 f_line_micro_1 <-ggplot(moist.mean.all.hab, aes(x = datetime, y = mean.moist.hab, color = habitat)) +
  geom_line() +
   geom_line(data = moist.min.all.hab, aes(x = datetime, y = min.moist.hab), linetype = "dashed") +
  geom_line(data = moist.max.all.hab, aes(x = datetime, y = max.moist.hab), linetype = "dashed") +
  facet_wrap(~ site) +
  labs(#title = "Average Soil Moisture Over Time",
       x = "Date",
       y = "Weekly Average Soil Solumetric Water Sontent",
       color = "Habitat")+
   theme(legend.position = "bottom")+
  theme(
    axis.title.x = element_text(size = 14), 
    axis.title.y = element_text(size = 14), 
    axis.text.x = element_text(size = 12), 
    axis.text.y = element_text(size = 12), 
    plot.subtitle = element_text(size = 14),
    plot.caption = element_text(size = 14),
    legend.text = element_text(size = 12)
  ) +
   theme(legend.position = "bottom") +
  theme_classic()

print(f_line_micro_1)

f_line_micro_2 <-ggplot(temp.mean.all.hab, aes(x = datetime, y = mean.temp.air.hab, color = habitat)) +
  geom_line() +
   geom_line(data = temp.min.all.hab, aes(x = datetime, y = min.temp.air.hab), linetype = "dashed") +
  geom_line(data = temp.max.all.hab, aes(x = datetime, y = max.temp.air.hab), linetype = "dashed") +
  facet_wrap(~ site) +
  labs(#title = "Average air temperature Over Time",
       x = "Date",
       y = "Weekly Average Air (15cm) Temperature (°C)",
       color = "Habitat")+
   theme(legend.position = "bottom")+
  theme(
    axis.title.x = element_text(size = 14), 
    axis.title.y = element_text(size = 14), 
    axis.text.x = element_text(size = 12), 
    axis.text.y = element_text(size = 12), 
    plot.subtitle = element_text(size = 14),
    plot.caption = element_text(size = 14),
    legend.text = element_text(size = 12)
  ) +
  theme_classic()

print(f_line_micro_2)

 plot_grid(f_line_micro_1, f_line_micro_2, ncol = 1)

snow.mean.all <- snow_all |> 
  select("loggerID" = locality_id, sensor_name, datetime, "snow" = value) 
snow.mean.all$loggerID = as.numeric(snow.mean.all$loggerID)

snow.mean.all <- active_metadata |> 
  select(loggerID, site, "habitat" = land_type) |> 
  left_join(snow.mean.all) |> 
  filter(habitat %in% c("Open", "Forested")) |> 
  group_by(site, datetime) |> 
  summarise(snow.hab = mean(snow, na.rm = TRUE)) |> 
  ungroup() 


f_line_micro_3 <-ggplot(snow.mean.all, aes(x = datetime, y = snow.hab))+
  geom_line() +
  facet_wrap(~ site) +
  labs(#title = "Average air temperature Over Time",
       x = "Date",
       y = "snow cover")+
  theme(
    axis.title.x = element_text(size = 14), 
    axis.title.y = element_text(size = 14), 
    axis.text.x = element_text(size = 12), 
    axis.text.y = element_text(size = 12), 
    plot.subtitle = element_text(size = 14),
    plot.caption = element_text(size = 14),
    legend.text = element_text(size = 12)
  ) +
  theme_classic()

print(f_line_micro_3)


GDD5.mean.all <- temp_env_all |> 
  filter(sensor_name == "T.air_2_cm.GDD5") |>  # T.soil_8_cm.GDD5
    select("loggerID" = locality_id, sensor_name, datetime, "GDD5_mean" = value) 
GDD5.mean.all$loggerID = as.numeric(GDD5.mean.all$loggerID)

GDD5.mean.all <- active_metadata |> 
  select(loggerID, site, "habitat" = land_type) |> 
  left_join(GDD5.mean.all) |> 
  filter(habitat %in% c("Open", "Forested")) |> 
  group_by(site, datetime) |> 
  summarise(GDD5.hab = mean(GDD5_mean, na.rm = TRUE)) |> 
  ungroup() |> 
  mutate(GDD5_day = ifelse(GDD5.hab > 0, 1, 0))

GDD5.month.all <- GDD5.mean.all |> 
  mutate(month_year = format(datetime, "%Y-%m")) |> 
  group_by(site, month_year) |> 
  summarise(GDD5_day_month = sum(GDD5_day, na.rm = TRUE))

FDD0.mean.all <- temp_env_all |> 
  filter(sensor_name == "T.soil_8_cm.FDD0") |> 
    select("loggerID" = locality_id, sensor_name, datetime, "FDD0_mean" = value) 
FDD0.mean.all$loggerID = as.numeric(FDD0.mean.all$loggerID)

FDD0.mean.all <- active_metadata |> 
  select(loggerID, site, "habitat" = land_type) |> 
  left_join(FDD0.mean.all) |> 
  filter(habitat %in% c("Open", "Forested")) |> 
  group_by(site, datetime) |> 
  summarise(FDD0.hab = mean(FDD0_mean, na.rm = TRUE)) |> 
  ungroup() |> 
  mutate(FDD0_day = ifelse(FDD0.hab > 0, 1, 0)) 

FDD0.month.all <- FDD0.mean.all |> 
  mutate(month_year = format(datetime, "%Y-%m")) |> 
  group_by(site, month_year) |> 
  summarise(FDD0_day_month = sum(FDD0_day, na.rm = TRUE))

DD.mean.all <- GDD5.mean.all |> 
  left_join(FDD0.mean.all, by = join_by(site, datetime)) |> 
  mutate(GDD5_day = ifelse(FDD0_day == 0, 1, 0)) |> 
  mutate(month_year = format(datetime, "%Y-%m")) |> 
  group_by(site, month_year) |> 
  summarise(FDD0_day_month = sum(FDD0_day, na.rm = TRUE),
            GDD5_day_month = sum(GDD5_day, na.rm = TRUE))

snow.month.all <- snow.mean.all |> 
  mutate(month_year_day = format(datetime, "%Y-%m-%d")) |> 
  group_by(site, month_year_day) |> 
  summarise(snow.day = sum(snow.hab, na.rm = TRUE)) |> 
  ungroup() |> 
  mutate(snow.day = ifelse(snow.day < 0.5, 0, 1)) |> 
  mutate(month_year = as.character(substr(month_year_day, 1, 7))) |> 
  group_by(site, month_year) |> 
  summarise(snow.day.month = sum(snow.day, na.rm = TRUE)) |> 
  ungroup()

snow.year <- snow.month.all |> 
  group_by(site) |> 
  summarise(snow.year = sum(snow.day.month, na.rm = TRUE))
 
f_line_micro_4 <- ggplot() +
  geom_line(data = GDD5.mean.all, aes(x = datetime, y = GDD5.hab, color = "Growing Degree Day (>5°C)")) +
  geom_line(data = FDD0.mean.all, aes(x = datetime, y = FDD0.hab, color = "Freezing Degree Day (<0°C)")) +
  facet_wrap(~ site) +
  labs(
    #title = "Average air temperature Over Time",
    x = "Date",
    y = "Sum of temperature (°C)",
    color = " "#Sum of temperature (°C)"
  ) +
  theme(legend.position = "bottom")+
  scale_color_manual(
    values = c("Growing Degree Day (>5°C)" = "green", "Freezing Degree Day (<0°C)" = "blue")
  )+
  theme(
    axis.title.x = element_text(size = 14), 
    axis.title.y = element_text(size = 14), 
    axis.text.x = element_text(size = 12), 
    axis.text.y = element_text(size = 12), 
    plot.subtitle = element_text(size = 14),
    plot.caption = element_text(size = 14),
    legend.text = element_text(size = 12)
  ) +
  theme_classic()

print(f_line_micro_4)


f_histo_micro <- ggplot() +
   geom_bar(data = DD.mean.all, 
           aes(x = month_year, y = FDD0_day_month, fill = "Freezing Day (<0°C)"),
           stat = "identity") +
  geom_bar(data = DD.mean.all, 
           aes(x = month_year, y = GDD5_day_month, fill = "Growing Day (>5°C)"),
           stat = "identity") +
 
  facet_wrap(~ site) +
  labs(
    x = "Date",
    y = "Number of day",
    fill = " Day Type"
  ) +
  theme(
    legend.position = "bottom",
    axis.text.x = element_text(angle = 45, hjust = 1),
    position = "dodge"
  )+
  scale_fill_manual(
    values = c("Growing Day (>5°C)" = "green", "Freezing Day (<0°C)" = "blue")
  ) +
  theme(
    axis.title.x = element_text(size = 14), 
    axis.title.y = element_text(size = 14), 
    axis.text.x = element_text(size = 12), 
    axis.text.y = element_text(size = 12), 
    plot.subtitle = element_text(size = 14),
    plot.caption = element_text(size = 14),
    legend.text = element_text(size = 12)
  ) +
  theme_classic()



print(f_histo_micro)

# Add a type column to both dataframes
GDD5.mean.all$type <- "Growing Degree Day (>5°C)"
FDD0.mean.all$type <- "Freezing Degree Day (<0°C)"

# Combine both datasets without creating a new combined dataframe
f_line_micro_5 <- ggplot() +
  geom_line(data = GDD5.mean.all, aes(x = datetime, y = GDD5.hab, color = site)) +
  geom_line(data = FDD0.mean.all, aes(x = datetime, y = FDD0.hab, color = site)) +
  facet_grid(type ~ ., scales = "free", labeller = labeller(type = c(
    "Growing Degree Day (>5°C)" = "Growing Degree Day (>5°C)",
    "Freezing Degree Day (<0°C)" = "Freezing Degree Day (<0°C)"
  ))) +
  labs(
    x = "Date",
    y = "Sum of temperature (°C)",
    color = "Site"
  )+
  theme(
    axis.title.x = element_text(size = 14), 
    axis.title.y = element_text(size = 14), 
    axis.text.x = element_text(size = 12), 
    axis.text.y = element_text(size = 12), 
    plot.subtitle = element_text(size = 14),
    plot.caption = element_text(size = 14),
    legend.text = element_text(size = 12)
  ) +
  theme_classic()

print(f_line_micro_5)

#fig_save <- "D:/Utilisateurs/Touriere/Documents/GEN/4A/SIRD/Rapport/Images/R"

join_macro <- ggarrange(
  f_line_macro_1, f_line_macro_2,
  labels = c("a)", "b)"),
  ncol = 1,
  nrow = 2,
  common.legend = TRUE,
  font.label = list(size = 17, face = "bold")
) 

join_macro

ggsave(filename = paste0(fig_save, "/macro_comp.png"), plot = join_macro, width = 10, height = 12, units = "in", dpi = 300)

join_micro <- ggarrange(
  f_line_micro_1, f_line_micro_2,
  labels = c("a)", "b)"),
  ncol = 1,
  nrow = 2,
  common.legend = TRUE,
  font.label = list(size = 17, face = "bold")
) 

join_micro

#ggsave(filename = paste0(fig_save, "/micro_comp.png"), plot = join_micro, width = 10, height = 12, units = "in", dpi = 300)
```

### Filtering metadata by site: Lygra, Sogndal, Kautokeino, Senja

```{r filtering by site}

site <- list(unique(active_metadata$site))

#filtering to only have data from lygra
ly_metadata <- active_metadata |> 
  filter(site == "Lygra") 

ly_metadata <-list("all_serial" = ly_metadata$loggerID,
                "Open"=filter(ly_metadata,land_type == "Open")$loggerID,
              "Forested"=filter(ly_metadata,land_type == "Forested")$loggerID,
           "DroughtNet"=filter(ly_metadata,land_type == "DroughtNet")$loggerID)
  
#filtering to only have data from sogndal
so_metadata <- active_metadata |> 
  filter(site == "Sogndal") 

so_metadata <-list("all_serial"=so_metadata$loggerID,
                "Open"=filter(so_metadata,land_type == "Open")$loggerID,
              "Forested"=filter(so_metadata,land_type == "Forested")$loggerID)
  
#filtering to only have data from Senja
se_metadata <- active_metadata |> 
  filter(site == "Senja") 

se_metadata <-list("all_serial"=se_metadata$loggerID,
                "Open"=filter(se_metadata,land_type == "Open")$loggerID,
              "Forested"=filter(se_metadata,land_type == "Forested")$loggerID)
 
#filtering to only have data from Kautokeino
ka_metadata <- active_metadata |> 
  filter(site == "Kautokeino") 

ka_metadata <-list("all_serial"=ka_metadata$loggerID,
                "Open"=filter(ka_metadata,land_type == "Open")$loggerID,
              "Forested"=filter(ka_metadata,land_type == "Forested")$loggerID)
```

### Filtering data by habitat : open vs forested (or DroughNet) for each site

```{r Data from lygra}
#filtering data for lygra
if(length(unique(ly_metadata$all_serial))!=0)
{
tms.ly  <- tms |> 
  mc_filter( localities = ly_metadata$all_serial, reverse = F)
}

if(length(unique(ly_metadata$Open))!=0) 
{
tms.ly.open <- tms |> 
  mc_filter( localities = ly_metadata$Open, reverse = F)
}

if(length(unique(ly_metadata$Forested))!=0) 
{
tms.ly.forested <- tms |> 
  mc_filter( localities = ly_metadata$Forested, reverse = F)
}



# if(length(unique(ly_metadata$DroughtNet))!=0)
# {
# tms.ly.DroughtNet <- tms |>
#   mc_filter( localities = ly_metadata$DroughtNet, reverse = F)
# 
# tms.ly.DroughtNet.90 <- tms |>
#   mc_filter( localities = ly_metadata$DroughtNet, reverse = F) |>
#   mc_filter(localities = filter(active_metadata, drt_treatment == "90")$loggerID)
# 
# tms.ly.DroughtNet.0 <- tms |>
#   mc_filter( localities = ly_metadata$DroughtNet, reverse = F) |>
#   mc_filter(localities = filter(active_metadata, drt_treatment == "0")$loggerID)
# }


#filtering data from Sogndal
if(length(unique(so_metadata$all_serial))!=0)
{
tms.so  <- tms |> 
  mc_filter( localities = so_metadata$all_serial, reverse = F)
}
if(length(unique(so_metadata$Open))!=0) 
{
tms.so.open <- tms |> 
  mc_filter( localities = so_metadata$Open, reverse = F)
}
if(length(unique(so_metadata$Forested))!=0) 
{
tms.so.forested <- tms |> 
  mc_filter( localities = so_metadata$Forested, reverse = F)
}

#filtering data for Senja
if(length(unique(se_metadata$all_serial))!=0)
{
tms.se  <- tms |> 
  mc_filter( localities = se_metadata$all_serial, reverse = F)
}

if(length(unique(se_metadata$Open))!=0) 
{
tms.se.open <- tms |> 
  mc_filter( localities = se_metadata$Open, reverse = F)
}
if(length(unique(se_metadata$Forested))!=0) 
{
tms.se.forested <- tms |> 
  mc_filter( localities = se_metadata$Forested, reverse = F)
}

#filtering data for Kautokeino
if(length(unique(ka_metadata$all_serial))!=0)
{
tms.ka  <- tms |> 
  mc_filter( localities = ka_metadata$all_serial, reverse = F)
}

if(length(unique(ka_metadata$Open))!=0) 
{
tms.ka.open <- tms |> 
  mc_filter( localities = ka_metadata$Open, reverse = F)
}
if(length(unique(ka_metadata$Forested))!=0) 
{
tms.ka.forested <- tms |> 
  mc_filter( localities = ka_metadata$Forested, reverse = F)
}
```

## Lygra

```{r calculation for all the data}
   
## calculate virtual sensor VWC from raw TMS moisture signal
tms.ly.calc <- mc_calc_vwc(tms.ly, soiltype = "universal") #reseach on soiltype and TOMST VERSION

## virtual sensor with growing and freezing degree days
tms.ly.calc <- mc_calc_gdd(tms.ly.calc, sensor = "TMS_T3",)
tms.ly.calc <- mc_calc_fdd(tms.ly.calc, sensor = "TMS_T3")

## virtual sensor to estimate snow presence from 2 cm air temperature 
tms.ly.calc <- mc_calc_snow(tms.ly.calc, sensor = "TMS_T2")

## summary data.frame of snow estimation
tms.snow <- mc_calc_snow_agg(tms.ly.calc)
 
temp_env  <- mc_env_temp(tms.ly, period = "month", min_coverage = 0.9)
moist_env <- mc_env_moist(tms.ly.calc, period = "month", min_coverage = 0.9)


```

```{r calculation for the open habitat}
#mean moisture
moist.mean.ly.open <- tms.ly.open |> 
   mc_calc_vwc( soiltype = "universal") |> 
   mc_env_moist(period = "day", min_coverage = 0.9) |> 
  filter(sensor_name == "VWC.soil_0_15_cm.mean") |> 
  pivot_wider(
    id_cols = c(datetime, sensor_name),
    names_from = locality_id,
    values_from = value)

moist.mean.ly.open <-  mutate(moist.mean.ly.open,
                              avg_value = rowMeans(select(moist.mean.ly.open, -datetime, -sensor_name), na.rm = TRUE))

# #min moisture  (not significant)
# moist.min.ly.open <- tms.ly.open |> 
#    mc_calc_vwc( soiltype = "universal") |> 
#    mc_env_moist(period = "day", min_coverage = 0.9) |> 
#   filter(sensor_name == "VWC.soil_0_15_cm.5p") |> 
#   pivot_wider(
#     id_cols = c(datetime, sensor_name),
#     names_from = locality_id,
#     values_from = value)
# 
# moist.min.ly.open <-  mutate(moist.min.ly.open,
#                              avg_value = rowMeans(select(moist.min.ly.open, -datetime, -sensor_name), na.rm = TRUE))


#mean temperature at 15cm
temp.mean.ly.open <- tms.ly.open |> 
   mc_env_temp(period = "day", min_coverage = 0.9) |> 
  filter(sensor_name == "T.air_15_cm.mean") |> 
  pivot_wider(
    id_cols = c(datetime, sensor_name),
    names_from = locality_id,
    values_from = value)

temp.mean.ly.open <-  mutate(temp.mean.ly.open,
                              avg_value = rowMeans(select(temp.mean.ly.open, -datetime, -sensor_name), na.rm = TRUE))

#mean gdd number
gdd.mean.ly.open <- tms.ly.open |> 
   mc_env_temp(period = "month", min_coverage = 0.9) |> 
  filter(sensor_name == "T.soil_8_cm.GDD5") |> 
  pivot_wider(
    id_cols = c(datetime, sensor_name),
    names_from = locality_id,
    values_from = value)
gdd.mean.ly.open <-  mutate(gdd.mean.ly.open,
                              avg_value = rowMeans(select(gdd.mean.ly.open, -datetime, -sensor_name), na.rm = TRUE))

#mean frost days number
fdd.mean.ly.open <- tms.ly.open |> 
   mc_env_temp(period = "month", min_coverage = 0.9) |> 
  filter(sensor_name == "T.soil_8_cm.FDD0") |> 
  pivot_wider(
    id_cols = c(datetime, sensor_name),
    names_from = locality_id,
    values_from = value)
fdd.mean.ly.open <-  mutate(fdd.mean.ly.open,
                              avg_value = rowMeans(select(fdd.mean.ly.open, -datetime, -sensor_name), na.rm = TRUE))

```

```{r calculation for the forested habitat}

#mean moisture
moist.mean.ly.forested <- tms.ly.forested |> 
   mc_calc_vwc( soiltype = "universal") |> 
   mc_env_moist(period = "day", min_coverage = 0.9) |> 
  filter(sensor_name == "VWC.soil_0_15_cm.mean") |> 
  pivot_wider(
    id_cols = c(datetime, sensor_name),
    names_from = locality_id,
    values_from = value)

moist.mean.ly.forested <-  mutate(moist.mean.ly.forested,avg_value = rowMeans(select(moist.mean.ly.forested, -datetime, -sensor_name), na.rm = TRUE))


#mean temperature at 15cm
temp.mean.ly.forested <- tms.ly.forested |> 
   mc_env_temp(period = "day", min_coverage = 0.9) |> 
  filter(sensor_name == "T.air_15_cm.mean") |> 
  pivot_wider(
    id_cols = c(datetime, sensor_name),
    names_from = locality_id,
    values_from = value)

temp.mean.ly.forested <-  mutate(temp.mean.ly.forested,
                              avg_value = rowMeans(select(temp.mean.ly.forested, -datetime, -sensor_name), na.rm = TRUE))
  

#mean gdd number
gdd.mean.ly.forested <- tms.ly.forested |> 
   mc_env_temp(period = "month", min_coverage = 0.9) |> 
  filter(sensor_name == "T.soil_8_cm.GDD5") |> 
  pivot_wider(
    id_cols = c(datetime, sensor_name),
    names_from = locality_id,
    values_from = value)
gdd.mean.ly.forested <-  mutate(gdd.mean.ly.forested,
                              avg_value = rowMeans(select(gdd.mean.ly.forested, -datetime, -sensor_name), na.rm = TRUE))

#mean frost days number
fdd.mean.ly.forested <- tms.ly.forested |> 
   mc_env_temp(period = "month", min_coverage = 0.9) |> 
  filter(sensor_name == "T.soil_8_cm.FDD0") |> 
  pivot_wider(
    id_cols = c(datetime, sensor_name),
    names_from = locality_id,
    values_from = value)
fdd.mean.ly.forested <-  mutate(fdd.mean.ly.forested,
                              avg_value = rowMeans(select(fdd.mean.ly.forested, -datetime, -sensor_name), na.rm = TRUE))

```

```{r calculation for the droughnet plots}

#mean moisture
moist.mean.ly.DN.90 <- tms.ly.DroughtNet.90 |> 
   mc_calc_vwc( soiltype = "universal") |> 
   mc_env_moist(period = "day", min_coverage = 0.9) |> 
  filter(sensor_name == "VWC.soil_0_15_cm.mean") |> 
  pivot_wider(
    id_cols = c(datetime, sensor_name),
    names_from = locality_id,
    values_from = value)

moist.mean.ly.DN.90 <-  mutate(moist.mean.ly.DN.90,avg_value = rowMeans(select(moist.mean.ly.DN.90, -datetime, -sensor_name), na.rm = TRUE))


#mean temperature at 15cm
temp.mean.ly.DN.90 <- tms.ly.DroughtNet.90 |> 
   mc_env_temp(period = "day", min_coverage = 0.9) |> 
  filter(sensor_name == "T.air_15_cm.mean") |> 
  pivot_wider(
    id_cols = c(datetime, sensor_name),
    names_from = locality_id,
    values_from = value)

temp.mean.ly.DN.90 <-  mutate(temp.mean.ly.DN.90,
                              avg_value = rowMeans(select(temp.mean.ly.DN.90, -datetime, -sensor_name), na.rm = TRUE))
  
#mean gdd number
gdd.mean.ly.DN.90 <- tms.ly.DroughtNet.90 |> 
   mc_env_temp(period = "month", min_coverage = 0.9) |> 
  filter(sensor_name == "T.soil_8_cm.GDD5") |> 
  pivot_wider(
    id_cols = c(datetime, sensor_name),
    names_from = locality_id,
    values_from = value)
gdd.mean.ly.DN.90 <-  mutate(gdd.mean.ly.DN.90,
                              avg_value = rowMeans(select(gdd.mean.ly.DN.90, -datetime, -sensor_name), na.rm = TRUE))

#mean frost days number
fdd.mean.ly.DN.90 <- tms.ly.DroughtNet.90 |> 
   mc_env_temp(period = "month", min_coverage = 0.9) |> 
  filter(sensor_name == "T.soil_8_cm.FDD0") |> 
  pivot_wider(
    id_cols = c(datetime, sensor_name),
    names_from = locality_id,
    values_from = value)
fdd.mean.ly.DN.90 <-  mutate(fdd.mean.ly.DN.90,
                              avg_value = rowMeans(select(fdd.mean.ly.DN.90, -datetime, -sensor_name), na.rm = TRUE))

```

## Microclimate analysis for Lygra

### Plot for the comparison

```{r moisture plots}

p.moist.1 <- ggplot() +
  geom_line(data = moist.mean.ly.open, aes(x = datetime, y = avg_value), color = "blue") +

  geom_line(data = moist.mean.ly.forested, aes(x = datetime, y = avg_value), color = "red") +
    # geom_line(data = moist.mean.ly.DN.90, aes(x = datetime, y = avg_value), color = "green") +
  
  labs(title = "Soil Moisture", x = "Time", y = "Soil Moisture Content") +
  theme(legend.position = c(x = 0.99, y = 0.01),
    legend.justification = c(x = "right", y = "bottom")) + 
  theme_classic()

p.moist.2 <- ggplot() +
  geom_line(data = moist.mean.ly.open, aes(x = datetime, y = avg_value, color = "Open habitat")) +
  geom_line(data = moist.mean.ly.forested, aes(x = datetime, y = avg_value, color = "Forested habitat")) +
    # geom_line(data = moist.mean.ly.DN.90, aes(x = datetime, y = avg_value, color = "DroughtNet 90%")) +
  labs(title = "Graphique en ligne du temps vs valeur", x = "Temps", y = "Valeur", color = "Habitat Type") +
  scale_color_manual(values = c("Open habitat" = "blue", "Forested habitat" = "red")) #,"DroughtNet 90%" = "green")) +
  
  theme(plot.title = element_text(hjust = 0.5)) + 
  theme_classic()

p.moist.1
p.moist.2
```

```{r temperature plots}

p.temp.1 <- ggplot() +
  geom_line(data = temp.mean.ly.open, aes(x = datetime, y = avg_value), color = "blue") +

  geom_line(data = temp.mean.ly.forested, aes(x = datetime, y = avg_value), color = "red") +
  # geom_line(data = temp.mean.ly.DN.90, aes(x = datetime, y = avg_value), color = "green") +
  
  labs(title = "Mean Temperature", x = "Time", y = "Temperature") +
  theme(legend.position = c(x = 0.99, y = 0.01),
    legend.justification = c(x = "right", y = "bottom")) + 
  theme_classic()

p.temp.1
```

```{r growing days}
p.gdd.1 <- ggplot() +
  geom_line(data = gdd.mean.ly.open, aes(x = datetime, y = avg_value), color = "blue") +

  geom_line(data = gdd.mean.ly.forested, aes(x = datetime, y = avg_value), color = "red") +
  # geom_line(data = gdd.mean.ly.DN.90, aes(x = datetime, y = avg_value), color = "green") +
  
  labs(title = "Growing Degree Days", x = "Time", y = "Days") +
  theme(legend.position = c(x = 0.99, y = 0.01),
    legend.justification = c(x = "right", y = "bottom")) + 
  theme_classic()

p.gdd.1

```

```{r Freezing degree days}
p.fdd.1 <- ggplot() +
  geom_line(data = fdd.mean.ly.open, aes(x = datetime, y = avg_value), color = "blue") +

  geom_line(data = fdd.mean.ly.forested, aes(x = datetime, y = avg_value), color = "red") +
  # geom_line(data = fdd.mean.ly.DN.90, aes(x = datetime, y = avg_value), color = "green") +
  
  labs(title = "Freezing Days", x = "Time", y = "Temperature") +
  theme(legend.position = c(x = 0.99, y = 0.01),
    legend.justification = c(x = "right", y = "bottom")) + 
  theme_classic()

p.fdd.1
```

```{r combining graph}

# Assembler les graphiques dans une seule image avec cowplot
combined_plot <- plot_grid(p.moist.1, p.temp.1, p.gdd.1, p.fdd.1, ncol = 2, labels = "AUTO")

# Ajouter une légende commune (exemple avec une légende fictive)
legend <- get_legend(p.moist.2 + theme_minimal())

final_plot <- cowplot::plot_grid(
  combined_plot,
  legend,
  nrow = 1,
  rel_widths = c(1, 0.2)
)

# Afficher le résultat
final_plot

```

# Function version of it

```{r}
library(ggplot2)
library(cowplot)

create_plot <- function(data_open, data_forested, title, y_label, colors) {
  p <- ggplot() +
    geom_line(data = data_open, aes(x = datetime, y = avg_value, color = "Open")) +
    geom_line(data = data_forested, aes(x = datetime, y = avg_value, color = "Forested")) +
    scale_color_manual(values = colors) +
    labs(title = title, x = "Time", y = y_label, color = "Site Type") +
    theme_classic() +
    theme(legend.position = "bottom")  # Ensuring the legend is positioned below
  
  return(p)
}

create_four_panel_plot <- function(dataset_list, colors) {
  # Create individual plots
  p.moist <- create_plot(dataset_list$moist_open, dataset_list$moist_forested, "Soil Moisture", "Soil Moisture Content", colors)
  p.temp <- create_plot(dataset_list$temp_open, dataset_list$temp_forested, "Mean Temperature", "Temperature", colors)
  p.gdd <- create_plot(dataset_list$gdd_open, dataset_list$gdd_forested, "Growing Degree Days", "Days", colors)
  p.fdd <- create_plot(dataset_list$fdd_open, dataset_list$fdd_forested, "Freezing Days", "Temperature", colors)

  # Combine the plots into a grid
  combined_plot <- plot_grid(p.moist + theme(legend.position = "none"), 
                             p.temp + theme(legend.position = "none"),
                             p.gdd + theme(legend.position = "none"), 
                             p.fdd + theme(legend.position = "none"), 
                             ncol = 2, labels = "AUTO")

  # Extract the shared legend from one plot (since all plots have the same legend)
  legend <- get_legend(p.moist)

  # Stack the four-panel figure on top of the shared legend
  final_plot <- plot_grid(
    combined_plot,
    legend,
    ncol = 1,  # Placing legend below by stacking plots in one column
    rel_heights = c(1, 0.1)  # Making the legend smaller relative to the plots
  )
  
  return(final_plot)
}

# Define your custom color palette
custom_colors <- c("Forested" = "#083508", "Open" = "#589758")  # Replace these with any colors you prefer


# Example usage:
dataset_list <- list(
  moist_open = moist.mean.ly.open,
  moist_forested = moist.mean.ly.forested,
  temp_open = temp.mean.ly.open,
  temp_forested = temp.mean.ly.forested,
  gdd_open = gdd.mean.ly.open,
  gdd_forested = gdd.mean.ly.forested,
  fdd_open = fdd.mean.ly.open,
  fdd_forested = fdd.mean.ly.forested
)

(final_plot2 <- create_four_panel_plot(dataset_list, custom_colors))

```

## Macroclimate analysis (between sites)

```{r}
# Sogndal Datasets
#mean moisture
moist.mean.so.forested <- tms.so.forested |> 
   mc_calc_vwc( soiltype = "universal") |> 
   mc_env_moist(period = "day", min_coverage = 0.9) |> 
  filter(sensor_name == "VWC.soil_0_15_cm.mean") |> 
  pivot_wider(
    id_cols = c(datetime, sensor_name),
    names_from = locality_id,
    values_from = value)

moist.mean.so.forested <-  mutate(moist.mean.so.forested,avg_value = rowMeans(select(moist.mean.so.forested, -datetime, -sensor_name), na.rm = TRUE))


#mean temperature at 15cm
temp.mean.so.forested <- tms.so.forested |> 
   mc_env_temp(period = "day", min_coverage = 0.9) |> 
  filter(sensor_name == "T.air_15_cm.mean") |> 
  pivot_wider(
    id_cols = c(datetime, sensor_name),
    names_from = locality_id,
    values_from = value)

temp.mean.so.forested <-  mutate(temp.mean.so.forested,
                              avg_value = rowMeans(select(temp.mean.so.forested, -datetime, -sensor_name), na.rm = TRUE))
  

#mean gdd number
gdd.mean.so.forested <- tms.so.forested |> 
   mc_env_temp(period = "month", min_coverage = 0.9) |> 
  filter(sensor_name == "T.soil_8_cm.GDD5") |> 
  pivot_wider(
    id_cols = c(datetime, sensor_name),
    names_from = locality_id,
    values_from = value)
gdd.mean.so.forested <-  mutate(gdd.mean.so.forested,
                              avg_value = rowMeans(select(gdd.mean.so.forested, -datetime, -sensor_name), na.rm = TRUE))

#mean frost days number
fdd.mean.so.forested <- tms.so.forested |> 
   mc_env_temp(period = "month", min_coverage = 0.9) |> 
  filter(sensor_name == "T.soil_8_cm.FDD0") |> 
  pivot_wider(
    id_cols = c(datetime, sensor_name),
    names_from = locality_id,
    values_from = value)
fdd.mean.so.forested <-  mutate(fdd.mean.so.forested,
                              avg_value = rowMeans(select(fdd.mean.so.forested, -datetime, -sensor_name), na.rm = TRUE))


# Senja Datasets
#mean moisture
moist.mean.se.forested <- tms.se.forested |> 
   mc_calc_vwc( soiltype = "universal") |> 
   mc_env_moist(period = "day", min_coverage = 0.9) |> 
  filter(sensor_name == "VWC.soil_0_15_cm.mean") |> 
  pivot_wider(
    id_cols = c(datetime, sensor_name),
    names_from = locality_id,
    values_from = value)

moist.mean.se.forested <-  mutate(moist.mean.se.forested,avg_value = rowMeans(select(moist.mean.se.forested, -datetime, -sensor_name), na.rm = TRUE))


#mean temperature at 15cm
temp.mean.se.forested <- tms.se.forested |> 
   mc_env_temp(period = "day", min_coverage = 0.9) |> 
  filter(sensor_name == "T.air_15_cm.mean") |> 
  pivot_wider(
    id_cols = c(datetime, sensor_name),
    names_from = locality_id,
    values_from = value)

temp.mean.se.forested <-  mutate(temp.mean.se.forested,
                              avg_value = rowMeans(select(temp.mean.se.forested, -datetime, -sensor_name), na.rm = TRUE))
  

#mean gdd number
gdd.mean.se.forested <- tms.se.forested |> 
   mc_env_temp(period = "month", min_coverage = 0.9) |> 
  filter(sensor_name == "T.soil_8_cm.GDD5") |> 
  pivot_wider(
    id_cols = c(datetime, sensor_name),
    names_from = locality_id,
    values_from = value)
gdd.mean.se.forested <-  mutate(gdd.mean.se.forested,
                              avg_value = rowMeans(select(gdd.mean.se.forested, -datetime, -sensor_name), na.rm = TRUE))

#mean frost days number
fdd.mean.se.forested <- tms.se.forested |> 
   mc_env_temp(period = "month", min_coverage = 0.9) |> 
  filter(sensor_name == "T.soil_8_cm.FDD0") |> 
  pivot_wider(
    id_cols = c(datetime, sensor_name),
    names_from = locality_id,
    values_from = value)
fdd.mean.se.forested <-  mutate(fdd.mean.se.forested,
                              avg_value = rowMeans(select(fdd.mean.se.forested, -datetime, -sensor_name), na.rm = TRUE))


# Senja Datasets
#mean moisture
moist.mean.ka.forested <- tms.ka.forested |> 
   mc_calc_vwc( soiltype = "universal") |> 
   mc_env_moist(period = "day", min_coverage = 0.9) |> 
  filter(sensor_name == "VWC.soil_0_15_cm.mean") |> 
  pivot_wider(
    id_cols = c(datetime, sensor_name),
    names_from = locality_id,
    values_from = value)

moist.mean.ka.forested <-  mutate(moist.mean.ka.forested,avg_value = rowMeans(select(moist.mean.ka.forested, -datetime, -sensor_name), na.rm = TRUE))


#mean temperature at 15cm
temp.mean.ka.forested <- tms.ka.forested |> 
   mc_env_temp(period = "day", min_coverage = 0.9) |> 
  filter(sensor_name == "T.air_15_cm.mean") |> 
  pivot_wider(
    id_cols = c(datetime, sensor_name),
    names_from = locality_id,
    values_from = value)

temp.mean.ka.forested <-  mutate(temp.mean.ka.forested,
                              avg_value = rowMeans(select(temp.mean.ka.forested, -datetime, -sensor_name), na.rm = TRUE))
  

#mean gdd number
gdd.mean.ka.forested <- tms.ka.forested |> 
   mc_env_temp(period = "month", min_coverage = 0.9) |> 
  filter(sensor_name == "T.soil_8_cm.GDD5") |> 
  pivot_wider(
    id_cols = c(datetime, sensor_name),
    names_from = locality_id,
    values_from = value)
gdd.mean.ka.forested <-  mutate(gdd.mean.ka.forested,
                              avg_value = rowMeans(select(gdd.mean.ka.forested, -datetime, -sensor_name), na.rm = TRUE))

#mean frost days number
fdd.mean.ka.forested <- tms.ka.forested |> 
   mc_env_temp(period = "month", min_coverage = 0.9) |> 
  filter(sensor_name == "T.soil_8_cm.FDD0") |> 
  pivot_wider(
    id_cols = c(datetime, sensor_name),
    names_from = locality_id,
    values_from = value)
fdd.mean.ka.forested <-  mutate(fdd.mean.ka.forested,
                              avg_value = rowMeans(select(fdd.mean.ka.forested, -datetime, -sensor_name), na.rm = TRUE))

```

```{r}
library(ggplot2)
library(cowplot)

# Function to create a plot for one variable across four sites with open/forested
create_plot <- function(data_list, title, y_label, colors, line_types) {
  p <- ggplot() +
    geom_line(data = data_list[[1]], aes(x = datetime, y = avg_value, color = "Site 1 Open", linetype = "Open")) +
    geom_line(data = data_list[[2]], aes(x = datetime, y = avg_value, color = "Site 1 Forested", linetype = "Forested")) +
    geom_line(data = data_list[[3]], aes(x = datetime, y = avg_value, color = "Site 2 Open", linetype = "Open")) +
    geom_line(data = data_list[[4]], aes(x = datetime, y = avg_value, color = "Site 2 Forested", linetype = "Forested")) +
    geom_line(data = data_list[[5]], aes(x = datetime, y = avg_value, color = "Site 3 Open", linetype = "Open")) +
    geom_line(data = data_list[[6]], aes(x = datetime, y = avg_value, color = "Site 3 Forested", linetype = "Forested")) +
    geom_line(data = data_list[[7]], aes(x = datetime, y = avg_value, color = "Site 4 Open", linetype = "Open")) +
    geom_line(data = data_list[[8]], aes(x = datetime, y = avg_value, color = "Site 4 Forested", linetype = "Forested")) +
    scale_color_manual(values = colors) +
    scale_linetype_manual(values = line_types) +
    labs(title = title, x = "Time", y = y_label, color = "Site", linetype = "Area Type") +
    theme_classic() +
    theme(legend.position = "bottom")
  
  return(p)
}

# Function to generate the four-panel figure
create_four_panel_plot <- function(dataset_list, colors, line_types) {
  p.moist <- create_plot(dataset_list$moist, "Soil Moisture", "Soil Moisture Content", colors, line_types)
  p.temp <- create_plot(dataset_list$temp, "Mean Temperature", "Temperature", colors, line_types)
  p.gdd <- create_plot(dataset_list$gdd, "Growing Degree Days", "Days", colors, line_types)
  p.fdd <- create_plot(dataset_list$fdd, "Freezing Days", "Temperature", colors, line_types)

  # Combine plots into a grid, removing individual legends
  combined_plot <- plot_grid(p.moist + theme(legend.position = "none"), 
                             p.temp + theme(legend.position = "none"),
                             p.gdd + theme(legend.position = "none"), 
                             p.fdd + theme(legend.position = "none"), 
                             ncol = 2, labels = "AUTO")

  # Extract shared legend
  legend <- get_legend(p.moist)

  # Arrange final figure with legend below
  final_plot <- plot_grid(combined_plot, legend, ncol = 1, rel_heights = c(1, 0.15))

  return(final_plot)
}

# Define custom colors per site (distinct colors for clarity)
custom_colors <- c("Site 1 Open" = "#1f78b4", "Site 1 Forested" = "#a6cee3",
                   "Site 2 Open" = "#33a02c", "Site 2 Forested" = "#b2df8a",
                   "Site 3 Open" = "#e31a1c", "Site 3 Forested" = "#fb9a99",
                   "Site 4 Open" = "#ff7f00", "Site 4 Forested" = "#fdbf6f")

# Define custom line types for open vs forested areas
custom_line_types <- c("Open" = "solid", "Forested" = "dashed")

# Example dataset input (each variable now has 8 datasets: 4 sites × 2 conditions)
dataset_list <- list(
  moist = list(moist.mean.ly.open, moist.mean.ly.forested,
               moist.mean., moist.site2.forested,
               moist.site3.open, moist.site3.forested,
               moist.site4.open, moist.site4.forested),
  temp = list(temp.site1.open, temp.site1.forested,
              temp.site2.open, temp.site2.forested,
              temp.site3.open, temp.site3.forested,
              temp.site4.open, temp.site4.forested),
  gdd = list(gdd.site1.open, gdd.site1.forested,
             gdd.site2.open, gdd.site2.forested,
             gdd.site3.open, gdd.site3.forested,
             gdd.site4.open, gdd.site4.forested),
  fdd = list(fdd.site1.open, fdd.site1.forested,
             fdd.site2.open, fdd.site2.forested,
             fdd.site3.open, fdd.site3.forested,
             fdd.site4.open, fdd.site4.forested)
)

# Generate the final plot
final_plot_allSites <- create_four_panel_plot(dataset_list, custom_colors, custom_line_types)


```

```{r}
temp.max.all.daily <- temp_env_all_daily |> 
  filter(sensor_name == "T.air_15_cm.max95p") |> 
    select("loggerID" = locality_id, sensor_name, datetime, "temp_max_air" = value) 
temp.max.all.daily$loggerID = as.numeric(temp.max.all.daily$loggerID)

temp.max.all.daily <- active_metadata |> 
  select(loggerID, site, "habitat" = land_type) |> 
  left_join(temp.max.all.daily) |> 
  filter(habitat %in% c("Open", "Forested")) |> 
  group_by(site, datetime) |>
  #group_by(site, datetime) |>
  summarise(max.temp.air = max(temp_max_air, na.rm = TRUE)) |>  
  ungroup()

temp.max.all.daily.hab <- temp_env_all_daily |> 
  filter(sensor_name == "T.air_15_cm.max95p") |>
    select("loggerID" = locality_id, sensor_name, datetime, "temp_max_air" = value) 
temp.max.all.daily.hab$loggerID = as.numeric(temp.max.all.daily.hab$loggerID)

temp.max.all.daily.hab <- active_metadata |> 
  select(loggerID, site, "habitat" = land_type) |> 
  left_join(temp.max.all.daily.hab) |> 
  filter(habitat %in% c("Open", "Forested")) |> 
  filter(!is.na(datetime)) |> 
  group_by(site, habitat, datetime) |>
  #group_by(site, datetime) |>
  summarise(temp.max.all.hab = max(temp_max_air, na.rm = TRUE)) |>  
  ungroup()

temp.max.all.daily.hab.log <- temp_env_all_daily |> 
  filter(sensor_name == "T.air_15_cm.max95p") |>
    select("loggerID" = locality_id, sensor_name, datetime, "temp_max_air" = value) 
temp.max.all.daily.hab.log$loggerID = as.numeric(temp.max.all.daily.hab.log$loggerID)

temp.max.all.daily.hab.log <- active_metadata |> 
  select(loggerID, site, "habitat" = land_type) |> 
  left_join(temp.max.all.daily.hab.log) |> 
  filter(habitat %in% c("Open", "Forested")) |> 
  filter(!is.na(datetime)) |> 
  group_by(site, habitat, loggerID, datetime) |>
  #group_by(site, datetime) |>
  summarise(temp.max = max(temp_max_air, na.rm = TRUE)) |>  
  ungroup()




library(dplyr)

# Function to count temperature exceedances
count_temp_exceedances <- function(temp_data, temp_column, site_column, habitat_column, date_column) {
  
  # Define thermal limits
  upper_limits <- c(1, 5, 25, 30, 35, 40, 45)
  lower_limits <- c(0, -5, -10, -20, -25)
  
  # Group data by site and habitat, then count exceedances
  results_df <- temp_data %>%
    group_by(!!sym(site_column), !!sym(habitat_column), !!sym(date_column)) %>%
    summarise(across(all_of(temp_column), list(
      Above_0C = ~ sum(. >= 1, na.rm = TRUE),
      Above_5C = ~ sum(. >= 5, na.rm = TRUE),
      Above_25C = ~ sum(. >= 25, na.rm = TRUE),
      Above_30C = ~ sum(. >= 30, na.rm = TRUE),
      Above_35C = ~ sum(. >= 35, na.rm = TRUE),
      Above_40C = ~ sum(. >= 40, na.rm = TRUE),
      Above_45C = ~ sum(. >= 45, na.rm = TRUE),
      Below_0C  = ~ sum(. <= 0, na.rm = TRUE),
      Below_minus5C  = ~ sum(. <= -5, na.rm = TRUE),
      Below_minus10C = ~ sum(. <= -10, na.rm = TRUE),
      Below_minus20C = ~ sum(. <= -20, na.rm = TRUE),
      Below_minus25C = ~ sum(. <= -25, na.rm = TRUE)
    ), .names = "{col}_{fn}"), .groups = "drop") %>%
    rename(Site = !!site_column, Habitat = !!habitat_column, Date = !!date_column)
  
  return(results_df)
}



temp_extremes <- count_temp_exceedances(temp_data = temp.max.all.daily.hab,
                                        temp_column =  "temp.max.all.hab",
                                        date_column = "datetime", 
                                        site_column = "site",
                                        habitat_column = "habitat")

# By loggerID for variance
# Function to count temperature exceedances with loggerID
count_temp_exceedances_logger <- function(temp_data, temp_column, site_column, habitat_column, date_column, loggerID_column) {
  
  # Define thermal limits
  upper_limits <- c(1, 5, 10, 15, 25, 30, 35, 40, 45)
  lower_limits <- c(0, -5, -10, -15, -20, -25)
  
  # Group data by site, habitat, datetime, and loggerID, then count exceedances
  results_df <- temp_data %>%
    group_by(!!sym(site_column), !!sym(habitat_column), !!sym(date_column), !!sym(loggerID_column)) %>%
    summarise(across(all_of(temp_column), list(
      Above_0C = ~ sum(. >= 1, na.rm = TRUE),
      Above_5C = ~ sum(. >= 5, na.rm = TRUE),
      Above_10C = ~ sum(. >= 10, na.rm = TRUE),
      Above_15C = ~ sum(. >= 15, na.rm = TRUE),
      Above_25C = ~ sum(. >= 25, na.rm = TRUE),
      Above_30C = ~ sum(. >= 30, na.rm = TRUE),
      Above_35C = ~ sum(. >= 35, na.rm = TRUE),
      Above_40C = ~ sum(. >= 40, na.rm = TRUE),
      Above_45C = ~ sum(. >= 45, na.rm = TRUE),
      Below_0C  = ~ sum(. <= 0, na.rm = TRUE),
      Below_minus5C  = ~ sum(. <= -5, na.rm = TRUE),
      Below_minus10C = ~ sum(. <= -10, na.rm = TRUE),
      Below_minus15C = ~ sum(. <= -15, na.rm = TRUE),
      Below_minus20C = ~ sum(. <= -20, na.rm = TRUE),
      Below_minus25C = ~ sum(. <= -25, na.rm = TRUE)
    ), .names = "{col}_{fn}"), .groups = "drop") %>%
    rename(Site = !!site_column, Habitat = !!habitat_column, Date = !!date_column, LoggerID = !!loggerID_column)
  
  return(results_df)
}

temp_extremes_logger <- count_temp_exceedances_logger(temp_data = temp.max.all.daily.hab.log,
                                        temp_column =  "temp.max",
                                        date_column = "datetime", 
                                        site_column = "site",
                                        habitat_column = "habitat",
                                        loggerID_column = "loggerID")



library(dplyr)

# Function to sum exceedances within a date range
library(dplyr)

# Function to sum exceedances within a date range
library(dplyr)

# Function to sum exceedances within a date range, grouped by site & habitat
sum_temp_exceedances_by_group <- function(data, start_date, end_date) {
  
  # Ensure Date column is formatted properly
  data <- data %>%
    mutate(Date = as.Date(Date)) %>%
    filter(Date >= as.Date(start_date) & Date <= as.Date(end_date))  # Filter by date range
  
  # Select only temperature exceedance columns
  exceedance_cols <- setdiff(names(data), c("Site", "Habitat", "Date"))
  
  # Summarize exceedances **by site & habitat**
  grouped_exceedances <- data %>%
    group_by(Site, Habitat) %>%
    summarise(across(all_of(exceedance_cols), \(x) sum(x, na.rm = TRUE)), .groups = "drop")
  
  return(grouped_exceedances)
}

# Example usage:
 grouped_totals <- sum_temp_exceedances_by_group(temp_extremes,
                                                 start_date = "2023-08-10",
                                                 end_date = "2024-07-05")
 print(grouped_totals)

 
 
 # By Logger ID
# Function to sum exceedances within a date range, grouped by site, habitat, and loggerID
count_dates_temp_exceedances <- function(temp_data, temp_column, site_column, habitat_column, date_column, loggerID_column) {
  
  # Define thermal thresholds
  thresholds <- c(1, 5, 10, 15, 25, 30, 35, 40, 45, 0, -5, -10, -15, -20, -25)
  
  # Count unique dates where exceedances occur per logger
  results_df <- temp_data %>%
    mutate(Date = as.Date(!!sym(date_column))) %>%
    group_by(!!sym(site_column), !!sym(habitat_column), !!sym(loggerID_column)) %>%
    summarise(across(all_of(temp_column), list(
      Above_0C = ~ sum(. >= 1, na.rm = TRUE),
      Above_5C = ~ sum(. >= 5, na.rm = TRUE),
      Above_10C = ~ sum(. >= 10, na.rm = TRUE),
      Above_15C = ~ sum(. >= 15, na.rm = TRUE),
      Above_25C = ~ sum(. >= 25, na.rm = TRUE),
      Above_30C = ~ sum(. >= 30, na.rm = TRUE),
      Above_35C = ~ sum(. >= 35, na.rm = TRUE),
      Above_40C = ~ sum(. >= 40, na.rm = TRUE),
      Above_45C = ~ sum(. >= 45, na.rm = TRUE),
      Below_0C  = ~ sum(. <= 0, na.rm = TRUE),
      Below_minus5C  = ~ sum(. <= -5, na.rm = TRUE),
      Below_minus10C = ~ sum(. <= -10, na.rm = TRUE),
      Below_minus15C = ~ sum(. <= -15, na.rm = TRUE),
      Below_minus20C = ~ sum(. <= -20, na.rm = TRUE),
      Below_minus25C = ~ sum(. <= -25, na.rm = TRUE)
    ), .names = "{col}_{fn}"), .groups = "drop") %>%
    rename(Site = !!site_column, Habitat = !!habitat_column, LoggerID = !!loggerID_column)

  return(results_df)
}


# Once we have per-logger exceedance counts, we can compute averages, standard deviation, and standard error for each site and habitat.

summarize_exceedances_per_site_habitat <- function(exceedance_data, site_column, habitat_column) {
  
  # Select exceedance count columns dynamically
  exceedance_cols <- setdiff(names(exceedance_data), c(site_column, habitat_column, "LoggerID"))

  # Summarize across loggers for each site and habitat
  summary_df <- exceedance_data %>%
    group_by(!!sym(site_column), !!sym(habitat_column)) %>%
    summarise(across(all_of(exceedance_cols), list(
      Mean = ~ mean(.x, na.rm = TRUE),
      SD = ~ sd(.x, na.rm = TRUE),
      SE = ~ sd(.x, na.rm = TRUE) / sqrt(n())
    ), .names = "{col}_{fn}"), .groups = "drop") %>%
    rename(Site = !!site_column, Habitat = !!habitat_column)

  return(summary_df)
}


# Step 1: Get exceedance count per logger
logger_exceedances <- count_dates_temp_exceedances(temp_data = temp.max.all.daily.hab.log,
                                                   temp_column = "temp.max",
                                                   site_column = "site",
                                                   habitat_column = "habitat",
                                                   loggerID_column = "loggerID",
                                                   date_column = "datetime")

# Step 2: Aggregate statistics across loggers for each site/habitat
site_habitat_summary <- summarize_exceedances_per_site_habitat(logger_exceedances,
                             site_column = "Site",
                            habitat_column = "Habitat")

print(site_habitat_summary)




library(ggplot2)
library(tidyr)  # Helps reshape data
library(stringr)  # For text cleaning

# Define custom colors for sites
Site_Colors <- c("Lygra" = "#1b9e77", "Sogndal" = "#d95f02",
                 "Senja" = "#7570b3", "Kautokeino" = "#aeb370")  # Adjust as needed

  

# Define custom shapes for habitats
Habitat_Shapes <- c("Forested" = 16, "Open" = 17)  # 16 = filled circle, 17 = triangle

# Function to plot exceedances with correct formatting, ordering, and site colors
plot_exceedances <- function(summary_data) {
  
  # Convert data to long format for plotting
  long_data <- summary_data %>%
    pivot_longer(cols = contains("_Mean"), names_to = "Threshold", values_to = "Mean") %>%
    mutate(Threshold = str_replace_all(Threshold, "_Mean", "")) 

  print("After pivot_longer:")
  print(unique(long_data$Threshold))  # Debugging step

  # Extract SD values for error bars
  sd_data <- summary_data %>%
    pivot_longer(cols = contains("_SD"), names_to = "Threshold_SD", values_to = "SD") %>%
    mutate(Threshold_SD = str_replace_all(Threshold_SD, "_SD", ""))

  # Merge mean and SD data
  plot_data <- long_data %>%
    left_join(sd_data, by = c("Site", "Habitat", "Threshold" = "Threshold_SD")) %>%
    mutate(Threshold = str_replace_all(Threshold, "temp.max_", "")) %>%
    mutate(Threshold = str_replace_all(Threshold, "Below_minus", "< -")) %>%
    mutate(Threshold = str_replace_all(Threshold, "Below_0C", "< 0°C")) %>%  # Keep "0" intact like negatives
    mutate(Threshold = str_replace_all(Threshold, "Above_", "> ")) %>%
    mutate(Threshold = str_replace_all(Threshold, "_", " ")) %>%
    mutate(Threshold = str_replace_all(Threshold, "(>|<) (-?\\d+)C$", "\\1 \\2°C"))  # Ensure proper degree formatting

  print("After applying Threshold formatting:")
  print(unique(plot_data$Threshold))  # Debugging step

  # **Explicitly define factor levels for correct ordering**
  ordered_levels <- c("< -25°C", "< -20°C","< -15°C", "< -10°C", "< -5°C", "< 0°C",
                      "> 0°C", "> 5°C", "> 10°C", "> 15°C", "> 25°C", "> 30°C", "> 35°C", "> 40°C", "> 45°C")

  # **TEST: Manually convert to factor before applying ggplot**
  test_levels <- factor(plot_data$Threshold, levels = ordered_levels, ordered = TRUE)
  
  print("Test Levels Before ggplot:")
  print(unique(test_levels))  # Debugging step

  # Convert to factor **inside the function**
  plot_data$Threshold <- factor(plot_data$Threshold, levels = ordered_levels, ordered = TRUE)

  print("Final ordered Thresholds:")
  print(unique(plot_data$Threshold))  # Debugging step before ggplot

  # Create the plot
  ggplot(plot_data, aes(x = Threshold, y = Mean, color = Site, shape = Habitat)) +
    geom_point(size = 3, position = position_dodge(0.5)) +
    geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD), width = 0.2, position = position_dodge(0.5)) +
    scale_color_manual(values = Site_Colors) +
    scale_shape_manual(values = Habitat_Shapes) +
    labs(title = "Mean Temperature Exceedances by Site & Habitat",
         x = "Temperature Thresholds",
         y = "Mean Count ± Standard Deviation",
         color = "Site",
         shape = "Habitat") +
    theme_minimal(base_size = 14) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
}


# Example usage:
plot_exceedances(site_habitat_summary)


library(ggplot2)
library(tidyr)
library(stringr)
library(dplyr)

Site_Colors <- c("Lygra" = "#1b9e77", "Sogndal" = "#d95f02",
                 "Senja" = "#7570b3", "Kautokeino" = "#aeb370")  # Adjust as needed

  
Habitat_Shapes <- c("Forested" = 16, "Open" = 17)

# Function to format & facet temperature bins with separate rows for "Above" & "Below"
plot_exceedances_faceted <- function(summary_data) {
  
  # Convert data to long format for plotting
  long_data <- summary_data %>%
    pivot_longer(cols = contains("_Mean"), names_to = "Threshold", values_to = "Mean") %>%
    mutate(Threshold = str_replace_all(Threshold, "_Mean", ""))

  print("After pivot_longer:")
  print(unique(long_data$Threshold))  # Debugging check

  # Extract SD values
  sd_data <- summary_data %>%
    pivot_longer(cols = contains("_SD"), names_to = "Threshold_SD", values_to = "SD") %>%
    mutate(Threshold_SD = str_replace_all(Threshold_SD, "_SD", ""))

  # Merge mean and SD data
  plot_data <- long_data %>%
    left_join(sd_data, by = c("Site", "Habitat", "Threshold" = "Threshold_SD")) %>%
    mutate(Threshold = str_replace_all(Threshold, "temp.max_", "")) %>%
    mutate(Threshold = str_replace_all(Threshold, "Below_minus", "< -")) %>%
    mutate(Threshold = str_replace_all(Threshold, "Below_0C", "< 0°C")) %>%
    mutate(Threshold = str_replace_all(Threshold, "Above_", "> ")) %>%
    mutate(Threshold = str_replace_all(Threshold, "_", " ")) %>%
    mutate(Threshold = str_replace_all(Threshold, "(>|<) (-?\\d+)C", "\\1 \\2°C")) 

  print("After applying Threshold formatting:")
  print(unique(plot_data$Threshold))  # Debugging check

  # Create a new column to classify Thresholds as "Above" or "Below"
  plot_data <- plot_data %>%
    mutate(Threshold_Type = ifelse(grepl(">", Threshold), "Above", "Below"))

  print("Threshold classification:")
  print(unique(plot_data$Threshold_Type))  # Debugging check

  # Define factor levels to ensure proper ordering
  plot_data$Threshold <- factor(plot_data$Threshold, 
                                levels = c("< -25°C", "< -20°C","< -15°C", "< -10°C", "< -5°C", "< 0°C",
                      "> 0°C", "> 5°C", "> 10°C", "> 15°C", "> 25°C", "> 30°C", "> 35°C", "> 40°C", "> 45°C"),
                                ordered = TRUE)

  print("Final ordered Thresholds:")
  print(unique(plot_data$Threshold))  # Debugging check

  # Convert Site to factor for proper color application
  plot_data$Site <- factor(plot_data$Site)

  # Create faceted plot with separate rows for Above vs. Below thresholds
ggplot(plot_data, aes(x = Habitat, y = Mean, color = Site, shape = Habitat, group = Site)) +
    geom_point(size = 3, position = position_dodge(0.5)) +  # Keeps points distinct
    geom_line(position = position_dodge(0.5), linewidth = 1) +  # Connects Open & Forested per site
    geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD), width = 0.2, position = position_dodge(0.5)) +
    facet_wrap(~ Threshold, scales = "free_y") +  # Independent facets for temperature bins
    scale_color_manual(values = Site_Colors) +
    scale_shape_manual(values = Habitat_Shapes) +
    scale_y_continuous(limits = c(0, NA)) +  # Prevents negative y-values
    labs(title = "Mean Temperature Exceedances by Habitat",
         x = "Habitat Type",
         y = "Mean Count ± SD",
         color = "Site",
         shape = "Habitat") +
    theme_minimal(base_size = 14) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))

}

# Example usage:
plot_exceedances_faceted(site_habitat_summary)


# Function to format & facet temperature bins with separate rows for "Above" & "Below"
plot_exceedances_faceted2 <- function(summary_data) {
  
  # Convert data to long format for plotting
  long_data <- summary_data %>%
    pivot_longer(cols = contains("_Mean"), names_to = "Threshold", values_to = "Mean") %>%
    mutate(Threshold = str_replace_all(Threshold, "_Mean", ""))

  print("After pivot_longer:")
  print(unique(long_data$Threshold))  # Debugging check

  # Extract SD values
  sd_data <- summary_data %>%
    pivot_longer(cols = contains("_SD"), names_to = "Threshold_SD", values_to = "SD") %>%
    mutate(Threshold_SD = str_replace_all(Threshold_SD, "_SD", ""))

  # Merge mean and SD data
  plot_data <- long_data %>%
    left_join(sd_data, by = c("Site", "Habitat", "Threshold" = "Threshold_SD")) %>%
    mutate(Threshold = str_replace_all(Threshold, "temp.max_", "")) %>%
    mutate(Threshold = str_replace_all(Threshold, "Below_minus", "< -")) %>%
    mutate(Threshold = str_replace_all(Threshold, "Below_0C", "< 0°C")) %>%
    mutate(Threshold = str_replace_all(Threshold, "Above_", "> ")) %>%
    mutate(Threshold = str_replace_all(Threshold, "_", " ")) %>%
    mutate(Threshold = str_replace_all(Threshold, "(>|<) (-?\\d+)C", "\\1 \\2°C")) 

  print("After applying Threshold formatting:")
  print(unique(plot_data$Threshold))  # Debugging check

  # Create a new column to classify Thresholds as "Above" or "Below"
  plot_data <- plot_data %>%
    mutate(Threshold_Type = ifelse(grepl(">", Threshold), "Above", "Below"))

  print("Threshold classification:")
  print(unique(plot_data$Threshold_Type))  # Debugging check

  # Define factor levels to ensure proper ordering
  plot_data$Threshold <- factor(plot_data$Threshold, 
                                levels = c("< -25°C", "< -20°C","< -15°C", "< -10°C", "< -5°C", "< 0°C",
                      "> 0°C", "> 5°C", "> 10°C", "> 15°C", "> 25°C", "> 30°C", "> 35°C", "> 40°C", "> 45°C"),
                                ordered = TRUE)

  print("Final ordered Thresholds:")
  print(unique(plot_data$Threshold))  # Debugging check

  # Convert Site to factor for proper color application
  plot_data$Site <- factor(plot_data$Site)

  # Create faceted plot with separate rows for Above vs. Below thresholds
  ggplot(plot_data, aes(x = Habitat, y = Mean, color = Site, shape = Habitat)) +
    geom_point(size = 3, position = position_dodge(0.5)) +
        geom_line(position = position_dodge(0.5), linewidth = 1) +  # Connects Open & Forested per site
    geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD), width = 0.2, position = position_dodge(0.5)) +
    facet_grid(cols = vars(Threshold), scales = "free_y") +  # Rows for Above vs. Below
    scale_color_manual(values = Site_Colors) +
    scale_shape_manual(values = Habitat_Shapes) +
    scale_y_continuous(limits = c(0, NA)) +  # Prevents negative y-values
    labs(title = "Mean Temperature Exceedances by Habitat",
         x = "Habitat Type",
         y = "Mean Count ± SD",
         color = "Site",
         shape = "Habitat") +
    theme_minimal(base_size = 14) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
}

# Example usage:
plot_exceedances_faceted2(site_habitat_summary)

```

```{r}
library(ggplot2)
library(ggridges)

# Function to create a temperature exceedences heatmap
plot_exceedances_heatmap <- function(summary_data) {
  
  # Convert data to long format for plotting
  long_data <- summary_data %>%
    pivot_longer(cols = contains("_Mean"), names_to = "Threshold", values_to = "Mean") %>%
    mutate(Threshold = str_replace_all(Threshold, "_Mean", ""))

  print("After pivot_longer:")
  print(unique(long_data$Threshold))  # Debugging check

  # Extract SD values
  sd_data <- summary_data %>%
    pivot_longer(cols = contains("_SD"), names_to = "Threshold_SD", values_to = "SD") %>%
    mutate(Threshold_SD = str_replace_all(Threshold_SD, "_SD", ""))

  # Merge mean and SD data
  plot_data <- long_data %>%
    left_join(sd_data, by = c("Site", "Habitat", "Threshold" = "Threshold_SD")) %>%
    mutate(Threshold = str_replace_all(Threshold, "temp.max_", "")) %>%
    mutate(Threshold = str_replace_all(Threshold, "Below_minus", "< -")) %>%
    mutate(Threshold = str_replace_all(Threshold, "Below_0C", "< 0°C")) %>%
    mutate(Threshold = str_replace_all(Threshold, "Above_", "> ")) %>%
    mutate(Threshold = str_replace_all(Threshold, "_", " ")) %>%
    mutate(Threshold = str_replace_all(Threshold, "(>|<) (-?\\d+)C", "\\1 \\2°C")) 

  print("After applying Threshold formatting:")
  print(unique(plot_data$Threshold))  # Debugging check

  # Create a new column to classify Thresholds as "Above" or "Below"
  plot_data <- plot_data %>%
    mutate(Threshold_Type = ifelse(grepl(">", Threshold), "Above", "Below"))

  print("Threshold classification:")
  print(unique(plot_data$Threshold_Type))  # Debugging check

  # Define factor levels to ensure proper ordering
  plot_data$Threshold <- factor(plot_data$Threshold, 
                                levels = c("< -25°C", "< -20°C","< -15°C", "< -10°C", "< -5°C", "< 0°C",
                      "> 0°C", "> 5°C", "> 10°C", "> 15°C", "> 25°C", "> 30°C", "> 35°C", "> 40°C", "> 45°C"),
                                ordered = TRUE)

  print("Final ordered Thresholds:")
  print(unique(plot_data$Threshold))  # Debugging check

  # Convert Site to factor for proper color application
  plot_data$Site <- factor(plot_data$Site)


library(scales)  # For color brightness calculations

# Determine text color based on background brightness
plot_data <- plot_data %>%
  # Switch based on exceedance count
    mutate(TextColor = ifelse(Mean > 200, "black", "white"))  %>% 
    mutate(Site_Habitat = paste(Site, Habitat, sep = "\n"))

ggplot(plot_data, aes(x = Site_Habitat, y = Threshold, fill = Mean)) +
    geom_tile(color = "white") +  # Ensures separation between tiles
    geom_text(aes(label = round(Mean, 0), color = TextColor), size = 4) +  # Adjusts text color dynamically
    scale_color_identity() +  # Uses pre-defined text colors
    scale_fill_viridis_c(option = "magma") +  # Heatmap color scheme
    scale_x_discrete(guide = guide_axis(angle = 0)) +  # Keep multi-line labels readable
    facet_wrap(~ Site, nrow = 1, scales = "free_x") +  # Separate facets per site
    theme_minimal() +
    theme(
        strip.text = element_text(size = 12, face = "bold"),  # Increases facet label size
        panel.spacing = unit(2, "lines")  # Adds extra space between site facets
    ) +
    labs(title = "Heatmap of Temperature Exceedances by Site & Habitat",
         x = "Site & Habitat",
         y = "Temperature Threshold",
         fill = "Mean Exceedance Count")
}

# Example usage:
plot_exceedances_heatmap(site_habitat_summary)
```

```{r}
temp.max.all.daily.hab.log <- temp_env_all_daily |> 
  filter(sensor_name == "T.air_15_cm.max95p") |>
    select("loggerID" = locality_id, sensor_name, datetime, "temp_max_air" = value) 
temp.max.all.daily.hab.log$loggerID = as.numeric(temp.max.all.daily.hab.log$loggerID)

temp.max.all.daily.hab.log <- active_metadata |> 
  select(loggerID, site, "habitat" = land_type) |> 
  left_join(temp.max.all.daily.hab.log) |> 
  filter(habitat %in% c("Open", "Forested")) |> 
  filter(!is.na(datetime)) |> 
  group_by(site, habitat, loggerID, datetime) |>
  #group_by(site, datetime) |>
  summarise(temp.max = max(temp_max_air, na.rm = TRUE)) |>  
  ungroup()

# Minimum temperature extremes
temp.min.all.daily.hab.log <- temp_env_all_daily |> 
  filter(sensor_name == "T.air_15_cm.min5p") |>
    select("loggerID" = locality_id, sensor_name, datetime, "temp_min_air" = value) 
temp.min.all.daily.hab.log$loggerID = as.numeric(temp.min.all.daily.hab.log$loggerID)

temp.min.all.daily.hab.log <- active_metadata |> 
  select(loggerID, site, "habitat" = land_type) |> 
  left_join(temp.min.all.daily.hab.log) |> 
  filter(habitat %in% c("Open", "Forested")) |> 
  filter(!is.na(datetime)) |> 
  group_by(site, habitat, loggerID, datetime) |>
  #group_by(site, datetime) |>
  summarise(temp.min = min(temp_min_air, na.rm = TRUE)) |>  
  ungroup()





# Function to count maximum temperature exceedances (Above thresholds)
count_max_temp_exceedances <- function(temp_data, site_column, habitat_column, date_column, loggerID_column) {
  
  # Define upper thermal thresholds
  thresholds <- c(1, 5, 10, 15, 25, 30, 35, 40, 45)

  results_df <- temp_data %>%
    mutate(Date = as.Date(!!sym(date_column))) %>%
    group_by(!!sym(site_column), !!sym(habitat_column), !!sym(loggerID_column)) %>%
    summarise(across("temp.max", list(
      Above_0C = ~ sum(. >= 1, na.rm = TRUE),
      Above_5C = ~ sum(. >= 5, na.rm = TRUE),
      Above_10C = ~ sum(. >= 10, na.rm = TRUE),
      Above_15C = ~ sum(. >= 15, na.rm = TRUE),
      Above_25C = ~ sum(. >= 25, na.rm = TRUE),
      Above_30C = ~ sum(. >= 30, na.rm = TRUE),
      Above_35C = ~ sum(. >= 35, na.rm = TRUE),
      Above_40C = ~ sum(. >= 40, na.rm = TRUE),
      Above_45C = ~ sum(. >= 45, na.rm = TRUE)
    ), .names = "{col}_{fn}"), .groups = "drop")
  
  return(results_df)
}

# Function to count minimum temperature exceedances (Below thresholds)
count_min_temp_exceedances <- function(temp_data, site_column, habitat_column, date_column, loggerID_column) {
  
  # Define lower thermal thresholds
  thresholds <- c(0, -5, -10, -15, -20, -25)

  results_df <- temp_data %>%
    mutate(Date = as.Date(!!sym(date_column))) %>%
    group_by(!!sym(site_column), !!sym(habitat_column), !!sym(loggerID_column)) %>%
    summarise(across("temp.min", list(
      Below_0C  = ~ sum(. <= 0, na.rm = TRUE),
      Below_minus5C  = ~ sum(. <= -5, na.rm = TRUE),
      Below_minus10C = ~ sum(. <= -10, na.rm = TRUE),
      Below_minus15C = ~ sum(. <= -15, na.rm = TRUE),
      Below_minus20C = ~ sum(. <= -20, na.rm = TRUE),
      Below_minus25C = ~ sum(. <= -25, na.rm = TRUE)
    ), .names = "{col}_{fn}"), .groups = "drop")
  
  return(results_df)
}

summarize_exceedances_per_site_habitat <- function(exceedance_data, site_column, habitat_column) {
  
  # exceedance_data <- exceedance_data %>%
  #   mutate(LoggerID = as.factor(LoggerID))  # Forces LoggerID to be categorical
  
  # Select exceedance count columns dynamically (excluding LoggerID)
  exceedance_cols <- setdiff(names(exceedance_data), c(site_column, habitat_column, "loggerID"))

  # **Sum exceedances across loggers**, then compute SD & SE
  summary_df <- exceedance_data %>%
    group_by(!!sym(site_column), !!sym(habitat_column)) %>%
    summarise(across(all_of(exceedance_cols), list(
      Mean = ~ mean(.x, na.rm = TRUE),  # Correct averaging (LoggerID NOT included)
      SD = ~ sd(.x, na.rm = TRUE),  # Standard deviation across loggers
      SE = ~ sd(.x, na.rm = TRUE) / sqrt(n())  # Correct application of sample size
    ), .names = "{col}_{fn}"), .groups = "drop") %>%
    rename(Site = !!site_column, Habitat = !!habitat_column)

  return(summary_df)
}



# Compute summaries for max and min exceedances separately
max_exceedance_summary <- summarize_exceedances_per_site_habitat(max_exceedances, "site", "habitat")
min_exceedance_summary <- summarize_exceedances_per_site_habitat(min_exceedances, "site", "habitat")

# Merge summaries into a single dataset
combined_summary <- full_join(max_exceedance_summary, min_exceedance_summary, by = c("Site", "Habitat"))




# Function to create a temperature exceedences heatmap
plot_exceedances_heatmap_combined <- function(summary_data) {
  
  # Convert data to long format for plotting
  long_data <- summary_data %>%
    pivot_longer(cols = contains("_Mean"), names_to = "Threshold", values_to = "Mean")


  print("After pivot_longer:")
  print(unique(long_data$Threshold))  # Debugging check

  # Extract SD values
  sd_data <- summary_data %>%
    pivot_longer(cols = contains("_SD"), names_to = "Threshold_SD", values_to = "SD") %>%
    mutate(Threshold_SD = str_replace_all(Threshold_SD, "_SD", ""))

  # Merge mean and SD data
  plot_data <- long_data %>%
    left_join(sd_data, by = c("Site", "Habitat", "Threshold" = "Threshold_SD")) %>%
    mutate(Threshold = str_replace_all(Threshold, "temp.max_", "")) %>%
    mutate(Threshold = str_replace_all(Threshold, "temp.min_", "")) %>%
    mutate(Threshold = str_replace_all(Threshold, "Below_minus", "< -")) %>%
    mutate(Threshold = str_replace_all(Threshold, "Below_0C", "< 0°C")) %>%
    mutate(Threshold = str_replace_all(Threshold, "Above_", "> ")) %>%
    mutate(Threshold = str_replace_all(Threshold, "_Mean", "")) %>%
    mutate(Threshold = str_replace_all(Threshold, "_", " ")) %>%
    mutate(Threshold = str_replace_all(Threshold, "(>|<) (-?\\d+)C", "\\1 \\2°C")) 

  print("After applying Threshold formatting:")
  print(unique(plot_data$Threshold))  # Debugging check

  # Create a new column to classify Thresholds as "Above" or "Below"
  plot_data <- plot_data %>%
    mutate(Threshold_Type = ifelse(grepl(">", Threshold), "Above", "Below"))

  print("Threshold classification:")
  print(unique(plot_data$Threshold_Type))  # Debugging check

  # Define factor levels to ensure proper ordering
  plot_data$Threshold <- factor(plot_data$Threshold, 
                                levels = c("< -25°C", "< -20°C","< -15°C", "< -10°C", "< -5°C", "< 0°C",
                      "> 0°C", "> 5°C", "> 10°C", "> 15°C", "> 25°C", "> 30°C", "> 35°C", "> 40°C", "> 45°C"),
                                ordered = TRUE)

  print("Final ordered Thresholds:")
  print(unique(plot_data$Threshold))  # Debugging check

  # Convert Site to factor for proper color application
  plot_data$Site <- factor(plot_data$Site)
  
  print("Check column names:")
  print(names(plot_data))  # Verify column names



library(scales)  # For color brightness calculations

# Determine text color based on background brightness
plot_data <- plot_data %>%
   mutate(Site_Habitat = paste(Site, Habitat, sep = "\n"))  %>% 
   mutate(Site_Habitat = str_replace_all(Site_Habitat, "\\n", " ")) %>%  # Replace line breaks with spaces
      mutate(Site_Habitat = as.character(Site_Habitat))  %>% 
  # Switch based on exceedance count
    mutate(TextColor = ifelse(Mean > 200, "black", "white"))
# print("Check column names:")
# print(names(plot_data))  # Check if 'Site_Habitat' is present
# print("Check unique values of Site_Habitat")  # Check if all values look correct
# print(unique(plot_data$Site_Habitat))  # Check if all values look correct
# print("Check class type of Site_Habitat")  # Check if all values look correct
# print(str(plot_data$Site_Habitat))  # Check if numeric or categorical etc. 
# print("Check for NA's in dataset")
# print(summary(plot_data))  # Look for any `NA` values in Site_Habitat
# print("Print TRUE if NA")
# print(any(is.na(plot_data$Site_Habitat)))  # TRUE if NA values exist



ggplot(plot_data, aes(x = Site_Habitat, y = Threshold, fill = Mean)) +
      #geom_tile(color = "white") +  # Ensures separation between tiles
    geom_tile(color = "white") +
    geom_text(aes(label = round(Mean, 1), color = ifelse(Mean > 200, "black", "white")), size = 4) +  
    scale_color_identity() +  # Uses pre-defined text colors
    scale_fill_viridis_c(option = "magma") +  # Heatmap color scheme
    scale_x_discrete(guide = guide_axis(angle = 0)) +  # Keep multi-line labels readable

   facet_wrap( ~ Site, nrow = 1, scales = "free") +  # Separate facets per site
   # facet_grid(rows = vars(Threshold_Type), scales = "free_y") +  # Adds separation for Above vs. Below within sites

    theme_minimal() +
   theme(legend.position = "bottom") +

    theme(
        strip.text = element_text(size = 12, face = "bold"),  # Increases facet label size
        panel.spacing = unit(2, "lines")  # Adds extra space between site facets
    ) +
    labs(title = "Heatmap of Maximum & Minimum Temperature Exceedances",
         x = "Site & Habitat",
         y = "Temperature Threshold",
         fill = "Number of Days Temperature Thresholds Were Exceeded")

}

# Example usage:
plot_exceedances_heatmap_combined(combined_summary)
```

```{r}
library(ggplot2)
library(tidyr)
library(stringr)
library(dplyr)
library(ggsignif)  # For statistical annotation
library(rstatix)   # For pairwise comparisons

# Define custom colors for sites
Site_Colors <- c("Lygra" = "#1b9e77", "Sogndal" = "#d95f02",
                 "Senja" = "#7570b3", "Kautokeino" = "#aeb370")  # Adjust as needed

# Define custom shapes for habitats
Habitat_Shapes <- c("Forested" = 16, "Open" = 17)

# Function to format and facet temperature bins
plot_exceedances <- function(summary_data) {
  
  # Convert data to long format for plotting
  long_data <- summary_data %>%
    pivot_longer(cols = contains("_Mean"), names_to = "Threshold", values_to = "Mean") %>%
    mutate(Threshold = str_replace_all(Threshold, "_Mean", ""))

  # Extract SD values
  sd_data <- summary_data %>%
    pivot_longer(cols = contains("_SD"), names_to = "Threshold_SD", values_to = "SD") %>%
    mutate(Threshold_SD = str_replace_all(Threshold_SD, "_SD", ""))

  # Merge mean and SD data
  plot_data <- long_data %>%
    left_join(sd_data, by = c("Site", "Habitat", "Threshold" = "Threshold_SD")) %>%
    mutate(Threshold = str_replace_all(Threshold, "temp.max_", "")) %>%
    mutate(Threshold = str_replace_all(Threshold, "Below_minus", "< -")) %>%
    mutate(Threshold = str_replace_all(Threshold, "Below_0C", "< 0°C")) %>%
    mutate(Threshold = str_replace_all(Threshold, "Above_", "> ")) %>%
    mutate(Threshold = str_replace_all(Threshold, "_", " ")) %>%
    mutate(Threshold = str_replace_all(Threshold, "(>|<) (-?\\d+)", "\\1 \\2°C")) 

  # Define factor levels to ensure proper ordering
  plot_data$Threshold <- factor(plot_data$Threshold, 
                                levels = c("< -25°C", "< -20°C", "< -10°C", "< -5°C", "< 0°C",
                                           "> 0°C", "> 5°C", "> 25°C", "> 30°C", "> 35°C", "> 40°C", "> 45°C"),
                                ordered = TRUE)

  # Convert Site to factor to ensure colors apply correctly
  plot_data$Site <- factor(plot_data$Site)

  # Create the faceted plot
  ggplot(plot_data, aes(x = Habitat, y = Mean, color = Site, shape = Habitat)) +
    geom_point(size = 3, position = position_dodge(0.5)) +
    geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD), width = 0.2, position = position_dodge(0.5)) +
    facet_wrap(~ Threshold, scales = "free_y") +  # Each temperature bin gets an independent facet
    scale_color_manual(values = Site_Colors) +
    scale_shape_manual(values = Habitat_Shapes) +
    labs(title = "Mean Temperature Exceedances by Site & Habitat",
         x = "Habitat Type",
         y = "Mean Count ± SD",
         color = "Site",
         shape = "Habitat") +
    theme_minimal(base_size = 14) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
}

# Example usage:
plot_exceedances(site_habitat_summary)

# -------------------------------------------------------------------------------------------------------
# Statistical Analysis: Identifying Significant Differences Between Sites & Habitats
# -------------------------------------------------------------------------------------------------------

# Convert summary_data to long format for analysis
stats_data <- site_habitat_summary %>%
  pivot_longer(cols = contains("_Mean"), names_to = "Threshold", values_to = "Mean") %>%
  mutate(Threshold = str_replace_all(Threshold, "_Mean", "")) %>%
  mutate(Threshold = str_replace_all(Threshold, "temp.max_", "")) %>%
  mutate(Threshold = str_replace_all(Threshold, "Below_minus", "< -")) %>%
  mutate(Threshold = str_replace_all(Threshold, "Below_0C", "< 0°C")) %>%
  mutate(Threshold = str_replace_all(Threshold, "Above_", "> ")) %>%
  mutate(Threshold = str_replace_all(Threshold, "_", " ")) %>%
  mutate(Threshold = str_replace_all(Threshold, "(>|<) (-?\\d+)", "\\1 \\2°C")) 

# ANOVA Test: Checking for effects of Site & Habitat
anova_results <- aov(Mean ~ Site * Habitat, data = stats_data)
summary(anova_results)

# If normality assumptions fail, switch to Kruskal-Wallis Test
kruskal_results <- stats_data %>%
  group_by(Threshold) %>%
  kruskal_test(Mean ~ Site * Habitat)
print(kruskal_results)

# Pairwise comparisons (Tukey's HSD or Dunn's Test for non-parametric data)
pairwise_results <- stats_data %>%
  group_by(Threshold) %>%
  tukey_hsd(Mean ~ Site * Habitat)
print(pairwise_results)

# Annotate plot with significant differences
sig_annotations <- stats_data %>%
  group_by(Threshold) %>%
  pairwise_t_test(Mean ~ Site * Habitat, p.adjust.method = "bonferroni")

# Add significance to plot
ggplot(stats_data, aes(x = Habitat, y = Mean, color = Site, shape = Habitat)) +
  geom_point(size = 3, position = position_dodge(0.5)) +
  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD), width = 0.2, position = position_dodge(0.5)) +
  facet_wrap(~ Threshold, scales = "free_y") +
  scale_color_manual(values = Site_Colors) +
  scale_shape_manual(values = Habitat_Shapes) +
  geom_signif(comparisons = list(c("Forested", "Open")), 
              map_signif_level = TRUE, textsize = 5) +
  labs(title = "Significant Differences in Temperature Exceedances",
       x = "Habitat Type",
       y = "Mean Count ± SD",
       color = "Site",
       shape = "Habitat") +
  theme_minimal(base_size = 14) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```
