---
title: "DURIN Thermal Tolerance Manuscript"
author: "Sonya R. Geange"
format: 
  html:
    self-contained: true
execute:
  echo: false
editor:
  markdown:
    wrap: sentence
---

```{r}
#| label: set-up
#| echo: false
#| message: false

# Load required packages and libraries
library(tidyverse)
library(lubridate)
library(ggplot2)
library(broom)
# library(gt)
library(lme4)
library(lmerTest)
library(broom.mixed)

```

# Title:

# Authors:

# Abstract

# Keywords

chlorophyll fluorescence, thermal load sensitivities,

# Introduction

\[Climate change in high-latitude ecosystems\]

\[Temperature Extremes\]

\[Dwarf-shrubs as a dominant plant functional type\]

# Methods

## Study design (sites, species, experiment)

| Site       | Biography  | Oceanity | GPS | Elevation | MAT (°C) | MAP (mm) |
|------------|------------|----------|-----|-----------|----------|----------|
| Lygra      | Boreal     | Coast    |     |           |          |          |
| Sogndal    | Boreal     | Inland   |     |           |          |          |
| Senja      | Sub-arctic | Coast    |     |           |          |          |
| Kautokeino | Sub-arctic | Inland   |     |           |          |          |

## Field Sampling

At each site and habitat, representative individuals from the four focal species were chosen for leaf-sampling.
Individuals were representative of the community (i.e. average height) and the current season leaves were fully-expanded and un-damagedndefined (Perez-...., 2016).
For *Calluna vulgaris*, strong seasonal signals in developmental status meant that samples taken from Lygra in June 2024, were not as mature as those from sites sampled later in the season.
Therefore, the most mature leaflets were sampled across the sites/habitats.

## Thermal Tolerance Assays and Plant Functional Traits

### Plant Functional Traits

At each site and habitat, we collected leaf samples from each of the four focal species for use in the thermal tolerance assays.
Individuals were chosen to be representative of the local community in terms of size (height, canopy dimensions) and niche occupation (e.g. avoiding moist depressions, or rocky outcrops).
Vegetative height was measured for individuals, before leaves were sample for traits and thermal tolerance.
Leaves were fully-expanded (relative to the season) and not damaged.
Where possible eight leaves were sampled for hot tolerance assays, and paired with six leaves for cold tolerance assays (total of 14 leaves).

In cases where individuals were not sufficiently large enough to provide either eight or sixteen leaves, multiple individuals from close proximity (\<10cm) were sampled collectively.
For the small-leaved species *Calluna vulgaris* and *Empentrum nigrum* the true-leaves were too small to be measured individually for Fv/Fm or plant functional traits, and therefore aggregates of leaves were used (often 5-6 leaves).
For *Calluna vulgaris*, leaves were defined as being xxxx (this was variable due to the fluctuating time within the growing season in which field-work was conducted).
Sampled individuals (leaves or whole plant) were stored in zip-lock bags with damp paper towels to maintain humidity during transport from the field to the lab, and kept in cool conditions (\~ 4 °C).

For each sample (entire leaf, or aggregated leaves), wet mass was measured (0.001g accuracy), leaf area was calculated by scanning the the total single sided surface area using a flatbed scanner and a semi-automated protocol

### Hot tolerance assays

Leaves were exposed to eight temperature treatments: a control temperature of 20 °C, along with 30 °C, 34 °C, 38 °C, 42 °C, 46 °C, 50 °C, and 54 °C.
Temperature treatments consisted of 15l plastic storage tubs filled with 10l of water, which was temperature controlled using Steba SV80 sous vide cookers.
Temperature baths were manually controlled, and temperatures were stablized for 20 minutes before the experiment began.
Storage containers were insulated from ambient light conditions, and LED grow lamps were positioned above the baths to provide light expsoure during the heat treatment periods (Add details).

All leaves were placed in the control bath for 20 minutes prior to leaves being distributed to the different temperature treatments for a further 20 minutes, before returning to the control bath for a 90 minute recovery period, post-stress.
After removal from the post-stress bath, samples were dark-adapted for 24hrs before chlorophyll fluorescence measures were taken.

For Fv/Fm measurements, the dark adapted leaf samples were placed within FluorPen dark adaptation clips for 20 minutes, before Qy (Fv/Fm proxy) was measured using a FluorPen 110D (Photon Systems Instruments).
Samples were then oven-dried at 65 °C for 72hrs, before being weighed for dry mass.

## Data Analysis

```{r}
1 + 1
```

You can add options to executable code like this

```{r}
#| echo: false
2 * 2
```
