---
title: "DURIN 2024 Thermal Tolerance: Methods Test"
format: html
editor: visual
---

## Vaccinium Methods Test for Time Exposure to Heat

This code conducts the cleaning and analysis for the 2024 Senja trials of thermal tolerance based on time exposure for Vaccinium vitis-ideae and Vaccinium myrtillus. Five idnividuals from each species were collected and leaves exposed to the temperatures 20, 30 , 34, 38, 42, 46, 50, and 54 degrees for 5, 10, 15 or 20 minutes respectively.

## Import Data

```{r}
library(tidylog)
library(tidyverse)
library(readxl)
library(stringr) 
library(plyr)
DURIN_data <- read_excel(path = "C:/Users/sge010/OneDrive - University of Bergen/Documents/DURIN/Durin_data_project/data/raw_data/2024 DURIN Plant Functional Traits.xlsx",
                         sheet = "MethodsData", col_names = TRUE, trim_ws = TRUE)

# Note for now, some issues with column types reading in.
# For example, the collection time with "10:59" format doesn't read in correctly, nor does the date format for Treatment Date.



# Also, some column names aren't 'readable', i.e. Fv/Fm with '/', or height(s)
DURIN_data <- rename(DURIN_data, c("height(s)"="heights",
                     "Fv/Fm_Initial"="FvFm_Initial",
                     "Fv/Fm_PostStress"="FvFm_PostStress",
                     "Fv/Fm_Hr"="FvFm_Hr",
                     "Fv/Fm_Final"="FvFm_Final",
                     "DroughNet_plotID" = "DroughtNet_plotID"))

# Senja is spelt wrongly as Senje in DropDowns
DURIN_data$siteID<- revalue(DURIN_data$siteID, c("Senje" = "Senja"))

# Create concantenated ID code for the plants
DURIN_data$PlantSampleID = paste(DURIN_data$siteID, DURIN_data$habitat,
                                   DURIN_data$species, DURIN_data$plant_nr,
                                   sep = "_")
DURIN_data$PlantSampleID2 = paste(DURIN_data$siteID, DURIN_data$habitat,
                                   DURIN_data$species, DURIN_data$plant_nr,
                                   sep = "_")

```

## 

## Exploratory Thermal Curves

```{r warning=FALSE}
# Hot Tolerance
DURIN_TT <- subset(DURIN_data, envelope_ID != "")

# Plot Initial Curves
DURIN_TT %>% 
  # Filters out issues with hot plants with temps below lowest bath of 20
  filter(TempExp >= 20) %>% 
  ggplot(aes(x = TempExp, y = as.numeric(as.character(FvFm_Final)),
             group = PlantSampleID, color = as.factor(TempExpTime))) +
  geom_point()+ 
  geom_smooth(method = "loess",
              se = FALSE, aes(group = PlantSampleID,
                              color = as.factor(TempExpTime)
                              )) +
  labs(x = "Temperature", y = "Fv/Fm", color = "Exposure Time") +
  facet_grid(species ~ siteID) +
  theme_bw()

# Plot Initial Curves
DURIN_TT %>% 
  # Filters out issues with hot plants with temps below the 2nd lowest bath of 30C.
  # We seem to have more variability at 20 than 30 for some reason?
  filter(TempExp >= 25) %>% 
  ggplot(aes(x = TempExp, y= as.numeric(as.character(FvFm_Final)),
             color = as.factor(TempExpTime))) +
  geom_point()+ 
  geom_smooth(method = "loess",
              se = FALSE, aes(
                color = as.factor(TempExpTime))) +
  facet_grid(species ~ siteID) +
  theme_bw() +
  labs(x = "Temperature (Celsius)", y = "FvFm", color = "Exposure Time") ## geom_vline(xintercept = 35) 


# Plot Initial Curves
DURIN_TT %>% 
  # Filters out issues with hot plants with temps below lowest bath of 20
  filter(TempExp >= 20) %>% 
  ggplot(aes(x = TempExp, y= as.numeric(as.character(FvFm_Final)),
             group = PlantSampleID, color = as.factor(TempExpTime))) +
  geom_point()+ 
  geom_smooth(method = "loess",
              se = FALSE, aes(group = PlantSampleID,
                              color = as.factor(as.numeric(TempExpTime)))) +
    labs(x = "Temperature (Celsius)", y = "FvFm", color = "Exposure Time") +
  facet_grid(species ~ siteID)

```

# 

# Using Feeley and Perez Function For Calculating T50 

```{r}

# Adapted from Feeley and Perez 2020 Paper.
#Thanks for checking out this code for estimating heat tolerances using variable fluorescence!
#Be sure to the read the Read.me.txt file.
#Below, the "psiiht" function can calculate heat tolerances of PSII heat tolerance. 
#For this function, define the the columns with the 1) Temperature & 2) FvFm variables, then the 3) control temperature, the 4) id/unique factor that you 
#want to estimate the heat tolerances for, 5) indicate if you want to make plots, 6) the number of bootsrap iterations for estimating heat tolerances

# Function to estimate T50 of FvFm
# Note 'HTdf' is the template dataset used in the function by the original authors.

psiiht <- function(Temperature, FvFm, control.temp, id, TempExpTime, plot.est, boots) {
  l1 <- list(control.temp = control.temp, plot.est = plot.est, boots = boots)
  HTdf <- data.frame(Temperature = Temperature, FvFm = FvFm, id = id, TempExpTime = TempExpTime)
  
  return(do.call("rbind", by(HTdf, list(HTdf$id, HTdf$TempExpTime), function(df) {
    Temperature <- df$Temperature
    FvFm <- df$FvFm
    id <- df$id
    TempExpTime <- df$TempExpTime
    
    # Get parameters for the logistics model
    custom_logit <- function(p) {
      p[p < 0.025] <- 0.025
      p[p > 0.975] <- 0.975
      return(log(p / (1 - p)))
    }
    logit_FvFm <- custom_logit(FvFm)
    cof <- coef(lm(logit_FvFm ~ Temperature))
    
    # Example: Adjusting initial values
    start_params <- list(
        theta1 = max(FvFm), 
        theta2 = cof[1], 
        theta3 = cof[2]
    )
    
    model_fit <- tryCatch({
      nls(FvFm ~ theta1 / (1 + exp(-(theta2 + theta3 * Temperature))),
          start = start_params, algorithm = "port",
          trace = FALSE, 
          lower = c(theta1 = 0, theta2 = -Inf, theta3 = -Inf), 
          upper = c(theta1 = 1, theta2 = Inf, theta3 = Inf),
          control = list(maxiter = 5000, tol = 1e-6))
    }, error = function(e) {
      message("Error while fitting model : ", e$message)
      return(NULL)
    })
    
   if (is.null(model_fit)) {
      # Check if id, habitat, and plant_nr are empty before including them in the data.frame
      if(length(unique(id)) > 0 && length(unique(TempExpTime)) > 0) {
        return(data.frame(id = unique(id), TempExpTime = unique(TempExpTime), Tcrit.mn = NA, T50.mn = NA, T95.mn = NA))
      } else {
        return(NULL)  # Return NULL if any of the elements are empty
      }
   }
    
    print(coef(model_fit))
    # TODO: Make sure to check the sequential values depending on the experimental set-up.
    # We start at 20, rather than 23 as per the original. 
    y <- coef(model_fit)[1] / (1 + exp(-(coef(model_fit)[2] + coef(model_fit)[3] * seq(20, 66))))
    
    # Calculation of T50 and T95 values
    half <- mean(na.omit(FvFm[which(Temperature == control.temp)])) / 2  
    nine5 <- mean(na.omit(FvFm[which(Temperature == control.temp)])) * 0.05  
    
    predict.boot <- matrix(NA, 47, boots)
    T95 <- T50 <- Tcrit <- c()
    
    for(k in 1:boots) {
    srows <- sample(1:length(Temperature), length(Temperature), TRUE)
      boot_fit <- tryCatch({
        nls(FvFm[srows] ~ theta1 / (1 + exp(-(theta2 + theta3 * Temperature[srows]))),  
            start = start_params, algorithm = "port",
            trace = FALSE, 
            lower = c(theta1 = 0, theta2 = -Inf, theta3 = -Inf), 
            upper = c(theta1 = 1, theta2 = Inf, theta3 = Inf),
            control = list(maxiter = 1000, tol = 1e-3))
      }, error = function(e) {
        return(NULL)
      })
      
      if (!is.null(boot_fit)) {
        predict.boot[, k] <- coef(boot_fit)[1] / (1 + exp(-(coef(boot_fit)[2] + coef(boot_fit)[3] * seq(20, 66))))
        
        T95[k] <- (-log((coef(boot_fit)[1] / nine5) - 1) - coef(boot_fit)[2]) / coef(boot_fit)[3]
        T50[k] <- (-log((coef(boot_fit)[1] / half) - 1) - coef(boot_fit)[2]) / coef(boot_fit)[3]
        
        predict <- data.frame(x = seq(20, 66), 
                              y = coef(boot_fit)[1] / (1 + exp(-(coef(boot_fit)[2] + coef(boot_fit)[3] * seq(20, 66)))))
        df1 <- cbind(predict[-1, ], predict[-nrow(predict), ])[, c(3, 1, 4, 2)]
        df1$slp <- as.vector(apply(df1, 1, function(x) summary(lm((x[3:4]) ~ x[1:2]))[[4]][[2]]))
        slp.at.tcrit <- round(min(df1$slp), 3) * .15 
        fvfv.at.tcrit <- df1[which(abs(df1[which(df1[,1] < T50[k]),]$slp - slp.at.tcrit) == min(abs(df1[which(df1[,1] < T50[k]),]$slp - slp.at.tcrit))),][1, 3]
        Tcrit[k] <- (-log((coef(boot_fit)[1] / fvfv.at.tcrit) - 1) - coef(boot_fit)[2]) / coef(boot_fit)[3]
      } else {
        predict.boot[, k] <- NA
        T95[k] <- NA
        T50[k] <- NA
        Tcrit[k] <- NA
      }
    }
    
          FvFm.boot <- t(apply(predict.boot, 1, function(x) {quantile(x, c(0.025, 0.975), na.rm = TRUE)}))
    T50.ci <- quantile(T50, c(0.025, 0.975), na.rm = TRUE)
          FvFm.boot <- t(apply(predict.boot, 1, function(x) {quantile(x, c(0.025, 0.975), na.rm = TRUE)}))
    T50.ci <- quantile(T50, c(0.025, 0.975), na.rm = TRUE)
    
if(plot.est == TRUE) {
  plot(NULL, NULL, xlab = "Temperature", ylab = "Fv/Fm", xlim = c(20, 66), ylim = c(0, 0.9), bty = "l", lty = 2)
  
  # Add text to the top of the chart
  mtext(paste(id, TempExpTime, sep = " "), side = 3, line = 1, cex = 1, font = 4)
  
  # Plot the points
  points(Temperature, FvFm, xlab = "Temperature", ylab = "Fv/Fm", xlim = c(20, 66), ylim = c(0, 0.85), bty = "l", lty = 2, pch = 5, col = "black")
  
  # Draw the curve if the data is available
  if(!is.null(y)) {
    lines(seq(20, 66), y, lwd = 1, col = "black")
  }
  
  # Add the curves of the bootstraps if they are available
  if(boots > 1) {
    if(!is.null(FvFm.boot)) {
      lines(seq(20, 66), FvFm.boot[, 1], lty = 3, col = "black")
      lines(seq(20, 66), FvFm.boot[, 2], lty = 3, col = "black")
    }
  }
  
  # Show the values of Tcrit, T50, and T95 if they are not NA
  if(!is.na(mean(na.omit(Tcrit)))) {
    text(30, 0, paste(round(mean(na.omit(Tcrit)), 1)), pos = 4, col = "light gray")
    abline(v = round(mean(na.omit(Tcrit)), 1), lty = 2, lwd = 1.5, col = "light gray", cex = 0.8)
  } else {
    text(30, 0, "NA", pos = 4, col = "light gray")
  }

  if(!is.na(mean(na.omit(T50)))) {
    text(30, .1, paste(round(mean(na.omit(T50)), 1)), pos = 4, col = "gray")
    abline(v = round(mean(na.omit(T50)), 1), lty = 2, lwd = 1.5, col = "gray", cex = 0.8)
  } else {
    text(30, .1, "NA", pos = 4, col = "gray")
  }

  if(!is.na(mean(na.omit(T95)))) {
    text(30, .2, paste(round(mean(na.omit(T95)), 1)), pos = 4, col = "black")
    abline(v = round(mean(na.omit(T95)), 1), lty = 2, lwd = 1.5, col = "black", cex = 0.8)
  } else {
    text(30, .2, "NA", pos = 4, col = "black")
  }
}
    
        
    if(length(unique(id)) > 0 && length(unique(TempExpTime)) > 0) {
      return(data.frame(id = unique(id), 
                        TempExpTime = unique(TempExpTime),
                        Tcrit.mn = round(na.omit(Tcrit), 1),  
                        T50.mn = round(na.omit(T50), 1),  
                        T95.mn = round(na.omit(T95), 1)))
    } else {
      return(NULL)
    }
  })))
}


# Run Function on dataset(s)
# Run on original combined dataset
Temp_Comb <- psiiht(Temperature=DURIN_Methods$TempExp,
       FvFm=DURIN_Methods$FvFm_Final_num,
       control.temp=30,
       id=DURIN_Methods$species,
       TempExpTime=DURIN_Methods$TempExpTime,
       plot.est=T,
       boots=100)

# Create seperate dataframes per species if needed
DURIN_Methods_VV <- DURIN_Methods %>% filter(species == "Vaccinium vitis-idaea")
DURIN_Methods_VM <- DURIN_Methods %>% filter(species == "Vaccinium myrtillus")

psiiht(Temperature=DURIN_Methods_VV$TempExp,
       FvFm=DURIN_Methods_VV$FvFm_Final_num,
       control.temp=30,
       id=DURIN_Methods_VV$species,
       TempExpTime=DURIN_Methods_VV$TempExpTime,
       plot.est=T,
       boots=5)

psiiht(Temperature=DURIN_Methods_VM$TempExp,
       FvFm=DURIN_Methods_VM$FvFm_Final_num,
       control.temp=30,
       id=DURIN_Methods_VM$species,
       TempExpTime=DURIN_Methods_VM$TempExpTime,
       plot.est=T,
       boots=5)

```

# Plot T50 outcomes

```{r}
# Using the combined outputs for all individuals and exposure times that were able to have an estimate generated. Use custom scale colours to match across different methods experiments. 

# Make Temperature Exposure Time a factor for plotting
Temp_Comb$TempExpTime <- as.factor(as.character(Temp_Comb$TempExpTime))   


## Reordering a factor's levels
levels(Temp_Comb$TempExpTime) 
Temp_Comb$TempExpTime <- factor(Temp_Comb$TempExpTime,
                                 levels = c('1', '5', '10', '15','20'),   
                                 labels = c('1', '5', '10', '15','20'))   


## Overview with a boxplot

(time_boxplot_Methods <- ggplot(Temp_Comb, aes(x = TempExpTime, y = T50.mn,
                                            fill = TempExpTime)) +
                 geom_boxplot() +
                 facet_grid(.~id) +
                 scale_fill_manual(values = c("#ea9999", "#e06666",
                                              "#cc0000","#990000", "#660000")) +  
                 labs(x = "Exposure Time (minutes)",
                      y = "T50 of Fv/Fm (ºC)",
                      fill = "Exposure Time (mins)") + 
                 theme_bw())

# Version with the legend at the bottom or the plot
(time_boxplot_Methods_base <- ggplot(Temp_Comb, aes(x = TempExpTime, y = T50.mn,
                                            fill = TempExpTime)) +
                 geom_boxplot() +
                 facet_grid(.~id) +
                 scale_fill_manual(values = c("#ea9999", "#e06666",
                                              "#cc0000","#990000", "#660000")) +  
                 labs(x = "Exposure Time (minutes)",
                      y = "T50 of Fv/Fm (ºC)",
                      fill = "Exposure Time (mins)") + 
                 theme_bw() +
                 theme(legend.position = "bottom"))

```

Analysis of Differences

```{r}
### ONE-WAY ANOVA -----

T50_anova <- aov(T50.mn ~ TempExpTime, data = Temp_Comb)
summary(T50_anova)

# Checking normality
## What to look for: The histogram of residuals should follow a normal (gaussian) distribution and the points in the Q-Q plot should lie mostly on the straight line. --> more or less accomplish
par(mfrow = c(1,2))  # This code put two plots in the same window
hist(T50_anova$residuals)   # Makes histogram of residuals  
plot(T50_anova, which = 2)   # Makes Q-Q plot

# Checking homoscedasticity (Homogeneity of variances)
## What to look for: We want to see a straight red line centered around zero! This means residuals do NOT systematically differ across different groups. --> quite perfectly centered
plot(T50_anova, which = 1)  # Makes residuals VS fitted plot

# Calculate Summary Statistics
summary_stats <- Temp_Comb %>%
                 group_by(TempExpTime) %>%
                 dplyr::summarise(n = n(),  # Calculating sample size n
                           average_T50 = mean(T50.mn),  
                           # Calculating mean hatching time
                           SD = sd(T50.mn))%>%  # Calculating standard deviation
                 mutate(SE = SD / sqrt(n))  # Calculating standard error


## Making a boxplot


(time_boxplot2 <- ggplot(summary_stats) +
     geom_errorbar(aes(x = TempExpTime, ymin = average_T50 - SE,
                                     ymax = average_T50 + SE),
                   width = 0.2,
                   colour = "black",
                   alpha = 0.9,
                   size = 0.5) + 
                     geom_point(aes(x = TempExpTime,
                                    y = average_T50, fill = TempExpTime),
                                size = 3) +
                 scale_fill_manual(values = c("#ea9999", "#e06666", "#cc0000","#990000", "#660000")) +  
                 labs(x = "Exposure Time (Minutes)",
                      y = "Average T50 value (ºC)",
                      fill = "Exposure Time (Minutes)") + 
                 theme_bw())



## post hoc table

## The means comparison by Tukey’s test can be run on the object resulting from the analysis of variance (anova). The result (below) is an extensive table with all pairwise comparisons and the p-value for each one of them. This data can be tricky to interpret and it is usual to use letters to indicate significant differences among the means.

# Tukey's test
tukey <- TukeyHSD(T50_anova)
print(tukey)
```
